### 2020-01-14 manuscript/manuscript.Rmd
Add parameters

---
title: "parameters"
author: "Gerbrich Ferdinands"
date: "1/14/2020"
output: pdf_document
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)
library(xtable)
library(kableExtra)
# load model parameters 
source("asr-parameters.R")
```

The ASReview software knows the following parameters: 
```{r}
# configurations table 
tab <- lapply(params, FUN = function(x) paste(x, collapse = ", "))
tab <- do.call(rbind, tab)
colnames(tab) <- c("configurations")

# combinations table
combins <- expand.grid(params)

kable(tab, format  = "pandoc")
```

\
This leads to  `r nrow(combins)` combinations of configurations:

```{r}
kable(combins, row.names = TRUE, format = "pandoc")
```