### 2020-01-16 manuscript/manuscript.Rmd
meeting 16 jan

---
title: "parameters"
author: "Gerbrich Ferdinands"
date: "1/14/2020"
output: pdf_document
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(knitr)
library(xtable)
library(kableExtra)
# load model parameters 
source("asr-parameters.R")
```

```{r}
# configurations table 
tab <- lapply(params, FUN = function(x) paste(x, collapse = ", "))
tab <- do.call(rbind, tab)
colnames(tab) <- c("configurations")
```

```{r}

# combinations table
combins <- expand.grid(params)

#combins$model %in% c("lstm_base", "lstm_pool") & combins$feature_extraction != "embeddingLSTM"

combins <- combins %>%
  #filter(combins$model %in% c("lstm_base", "lstm_pool") & feature_extraction != "embeddingLSTM") %>%
  arrange(model, feature_extraction )

# rows to drop
droprows <- combins$model == "nb" & combins$feature_extraction != "tfidf"
combins <- combins[droprows, ]
```

ASReview takes the following parameters/arguments:

- number of training data (included/excluded)
- a model
- a query strategy
- a balance strategy (fixed)
- feature extraction

```{r}
kable(tab, format  = "pandoc")
```

This leads to  `r nrow(combins)` combinations of configurations:

```{r}
kable(combins, row.names = TRUE, format = "pandoc")
```