### 2020-01-20 definitions.Rmd
Fixed table combinations, put into latex format

---
title: "Definition list"
author: "Gerbrich Ferdinands"
date: "1/20/2020"
output: pdf_document
---
Term 1

~   Definition 1a
~   definition 1b
~ definition

Term 2 with *inline markup*

:   Definition 2

        { some code, part of Definition 2 }

    Third paragraph of definition 2.