### 2020-01-20 manuscript/manuscript.Rmd

---
title: "parameters"
author: "Gerbrich Ferdinands"
date: "1/14/2020"
output: pdf_document
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(knitr)
library(xtable)
library(kableExtra)
# load model parameters 
source("asr-parameters.R")
```

```{r}
# configurations table 
tab <- lapply(params, FUN = function(x) paste(x, collapse = ", "))
tab <- do.call(rbind, tab)
colnames(tab) <- c("configurations")
```

```{r}

# combinations table
combins <- expand.grid(params) %>%
  arrange(model, feature_extraction)

# rows to drop
droprows <- combins$model == "nb" & combins$feature_extraction != "tfidf"
combins <- combins[!droprows, ]
```

ASReview takes the following parameters/arguments:

- number of training data (included/excluded)
- a model
- a query strategy
- a balance strategy (fixed)
- feature extraction

```{r}
kable(tab, "latex", booktabs = TRUE) %>%
  kable_styling(full_width = FALSE) %>%
  column_spec(2, width = "30em") 
```

This leads to  `r nrow(combins)` combinations of configurations.

- Naive bayes only goes with tfidf feature extraction.
- For the feature extraction strategies we will focus on doc2vec and tfidf. (but will compute all 4)
- This leads to 3 * 7  * 4 * 3 + 1 * 7 * 1 * 3 =  273 combinations.

```{r}
combins %>%
  kable(row.names = FALSE, format = "latex", longtable = TRUE, booktabs = TRUE) %>%
  kable_styling(latex_options = c("repeat_header")) %>%
  footnote(general = "nb model only works with tfidf feature extraction strategy. " )
```