### 2020-01-28 manuscript/index/index.Rmd

---
author: 'Gerbrich Ferdinands'
date: 'May 2020'
institution: 'Utrecht University'
division: 'Methodolgy and Statistics'
advisor: 'prof. dr. Rens van de Schoot'
# If you have more two advisors, un-silence line 7
#altadvisor: 'Your Other Advisor'
department: 'Faculty of Social Sciences'
degree: 'Master of Science'
title: 'My Master thesis'
knit: "bookdown::render_book"
site: bookdown::bookdown_site
output: 
# thesisdown::thesis_pdf: default
thesisdown::thesis_gitbook: default
#  thesisdown::thesis_word: default
# thesisdown::thesis_epub: default
# If you are creating a PDF you'll need to write your preliminary content (e.g., abstract, acknowledgements) here or
# use code similar to line 22-23 for the .RMD files. If you are NOT producing a PDF, you can delete or silence lines 21-32 in this YAML header.
abstract: |
  `r if(knitr:::is_latex_output()) paste(readLines("00-abstract.Rmd"), collapse = '\n  ')`
# If you'd rather include the preliminary content in files instead of inline
# like below, use a command like that for the abstract above.  Note that a tab is 
# needed on the line after the `|`.
acknowledgements: |
  I want to thank a few people.
#dedication: |
#  You can have a dedication here if you wish.
preface: |
  This is an example of a thesis setup to use the reed thesis document class
  (for LaTeX) and the R bookdown package, in general.
# Specify the location of the bibliography below
bibliography: bib/thesis.bib
# Download your specific csl file and refer to it in the line below.
csl: csl/apa.csl
lot: true
lof: true
# If you prefer blank lines between paragraphs, un-silence lines  40-41 (this requires package tikz)
#header-includes:
#- \usepackage{tikz}
---

<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of metadata used to produce the document.  Be careful with spacing in this header!

If you'd prefer to not include a Dedication, for example, simply delete the section entirely, or silence (add #) them.

If you have other LaTeX packages you would like to include, delete the # before header-includes and list the packages after hyphens on new lines.

If you'd like to include a comment that won't be produced in your resulting file enclose it in a block like this.

If you receive a duplicate label error after knitting, make sure to delete the index.Rmd file and then knit again.
-->

```{r include_packages, include = FALSE}
# This chunk ensures that the thesisdown package is
# installed and loaded. This thesisdown package includes
# the template files for the thesis.
if(!require(devtools))
  install.packages("devtools", repos = "http://cran.rstudio.com")
if(!require(thesisdown))
  devtools::install_github("ismayc/thesisdown")
library(thesisdown)
```

<!-- On ordering the chapter files:
There are two options:
1. Name your chapter files in the order in which you want them to appear (e.g., 01-Inro, 02-Data, 03-Conclusions). 
2. Otherwise, you can specify the order in which they appear in the _bookdown.yml (for PDF only).

Do not include 00(two-hyphens)prelim.Rmd and 00-abstract.Rmd in the YAML file--they are handled in the YAML above differently for the PDF version.
-->

<!-- The {.unnumbered} option here means that the introduction will be "Chapter 0." You can also use {-} for no numbers
on chapters.
-->

# Introduction {.unnumbered}

<!-- 
If you're still on the fence about using _R Markdown_, check out the resource for newbies available at <https://ismayc.github.io/rbasics-book/> or email us at <data@reed.edu>. 
-->

**Why use it?**

_R Markdown_ creates a simple and straightforward way to interface with the beauty of LaTeX.  Packages have been written in **R** to work directly with LaTeX to produce nicely formatting tables and paragraphs. In addition to creating a user friendly interface to LaTeX, _R Markdown_ also allows you to read in your data, to analyze it and to visualize it using **R** functions, and also to provide the documentation and commentary on the results of your project.  Further, it allows for **R** results to be passed inline to the commentary of your results.  You'll see more on this later.

<!-- 
Having your code and commentary all together in one place has a plethora of benefits!
-->

**Who should use it?**

Anyone who needs to use data analysis, math, tables, a lot of figures, complex cross-references, or who just cares about the final appearance of their document should use _R Markdown_. Of particular use should be anyone in the sciences, but the user-friendly nature of _Markdown_ and its ability to keep track of and easily include figures, automatically generate a table of contents, index, references, table of figures, etc. should make it of great benefit to nearly anyone writing a thesis project.

**For additional help with bookdown** 
Please visit [the free online bookdown reference guide](https://bookdown.org/yihui/bookdown/).