### 2020-02-04 manuscript/index/05-appendix.Rmd
a lot of stuff: fix bugs add bookdown

`r if(knitr:::is_latex_output()) '\\appendix'`

`r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'`

<!--
If you feel it necessary to include an appendix, it goes here.
-->

# List of Definitions
Model.

:   The prediction model for Active learning

        some code, part of Definition 2

    Third paragraph of definition 2.

Query Strategy 
:   The way papers are queried to the researcher

        query_strategy

    Uncertainty sampling
    :  Selects the least sure instances for labelling.

    Maximum sampling
    :  Selects the samples with the highest prediction probability.

    Random and maximum sampling. 
    :  Combination of random and maximum sampling. By default samples the 95% of the instances with max sampling, and 5% of the samples with random sampling.

Feature Extraction Strategy
:   Feature Extraction Strategy

        feature_extraction

    Third paragraph of definition x.

    Doc2Vec
    :

    Term frequency-inverse document frequency (tf-idf)
    :

    sbert

<!-- # The Last Appendix -->

<!-- This first appendix includes all of the R chunks of code that were hidden throughout the document (using the `include = FALSE` chunk tag) to help with readibility and/or setup. -->

<!-- **In the main Rmd file** -->

<!-- ```{r ref.label='include_packages', results='hide', echo = TRUE} -->
<!-- ``` -->

<!-- **In Chapter \@ref(ref-labels):** -->

<!-- ```{r ref.label='include_packages_2', results='hide', echo = TRUE} -->
<!-- ``` -->