### 2020-02-04 manuscript/drafts/definitions.Rmd

---
title: "Definition list"
author: "Gerbrich Ferdinands"
date: "1/20/2020"
output: pdf_document
---

# List of definitions

Model.

:   The prediction model for Active learning

        some code, part of Definition 2

    Third paragraph of definition 2.

Query Strategy 
:   The way papers are queried to the researcher

        query_strategy

    Uncertainty sampling
    :  Selects the least sure instances for labelling.

    Maximum sampling
    :  Selects the samples with the highest prediction probability.

    Random and maximum sampling. 
    :  Combination of random and maximum sampling. By default samples the 95% of the instances with max sampling, and 5% of the samples with random sampling.

Feature Extraction Strategy
:   Feature Extraction Strategy

        feature_extraction

    Third paragraph of definition x.

    Doc2Vec
    :

    Term frequency-inverse document frequency (tf-idf)
    :

    sbert