### 2020-02-04 manuscript/index/_book/thesis.tex
ignored files

% This is the Reed College LaTeX thesis template. Most of the work
% for the document class was done by Sam Noble (SN), as well as this
% template. Later comments etc. by Ben Salzberg (BTS). Additional
% restructuring and APA support by Jess Youngberg (JY).
% Your comments and suggestions are more than welcome; please email
% them to cus@reed.edu
%
% See http://web.reed.edu/cis/help/latex.html for help. There are a
% great bunch of help pages there, with notes on
% getting started, bibtex, etc. Go there and read it if you're not
% already familiar with LaTeX.
%
% Any line that starts with a percent symbol is a comment.
% They won't show up in the document, and are useful for notes
% to yourself and explaining commands.
% Commenting also removes a line from the document;
% very handy for troubleshooting problems. -BTS

% As far as I know, this follows the requirements laid out in
% the 2002-2003 Senior Handbook. Ask a librarian to check the
% document before binding. -SN

%%
%% Preamble
%%
% \documentclass{<something>} must begin each LaTeX document
\documentclass[12pt,twoside]{reedthesis}
% Packages are extensions to the basic LaTeX functions. Whatever you
% want to typeset, there is probably a package out there for it.
% Chemistry (chemtex), screenplays, you name it.
% Check out CTAN to see: http://www.ctan.org/
%%
\usepackage{graphicx,latexsym}
\usepackage{amsmath}
\usepackage{amssymb,amsthm}
\usepackage{longtable,booktabs,setspace}
\usepackage{chemarr} %% Useful for one reaction arrow, useless if you're not a chem major
\usepackage[hyphens]{url}
% Added by CII
\usepackage{hyperref}
\usepackage{lmodern}
\usepackage{float}
\floatplacement{figure}{H}
% End of CII addition
\usepackage{rotating}

% Next line commented out by CII
%%% \usepackage{natbib}
% Comment out the natbib line above and uncomment the following two lines to use the new
% biblatex-chicago style, for Chicago A. Also make some changes at the end where the
% bibliography is included.
%\usepackage{biblatex-chicago}
%\bibliography{thesis}

% Gerbrich: try tro solve clsreferences.

% Added by CII (Thanks, Hadley!)
% Use ref for internal links
\renewcommand{\hyperref}[2][???]{\autoref{#1}}
\def\chapterautorefname{Chapter}
\def\sectionautorefname{Section}
\def\subsectionautorefname{Subsection}
% End of CII addition

% Added by CII
\usepackage{caption}
\captionsetup{width=5in}
% End of CII addition

% \usepackage{times} % other fonts are available like times, bookman, charter, palatino

% Syntax highlighting #22

% To pass between YAML and LaTeX the dollar signs are added by CII
\title{My thesis}
\author{Gerbrich Ferdinands}
% The month and year that you submit your FINAL draft TO THE LIBRARY (May or December)
\date{May 2020}
\division{Faculty of Social Sciences}
\advisor{prof. dr. Rens van de Schoot}
\institution{Utrecht University}
\degree{Master of Science}
%If you have two advisors for some reason, you can use the following
% Uncommented out by CII
\altadvisor{Jonathan de Bruin, Raoul Schram}
% End of CII addition

%%% Remember to use the correct department!
\department{Methodology and Statistics}
% if you're writing a thesis in an interdisciplinary major,
% uncomment the line below and change the text as appropriate.
% check the Senior Handbook if unsure.
%\thedivisionof{The Established Interdisciplinary Committee for}
% if you want the approval page to say "Approved for the Committee",
% uncomment the next line
%\approvedforthe{Committee}

% Added by CII
%%% Copied from knitr
%% maxwidth is the original width if it's less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\renewcommand{\contentsname}{Table of Contents}
% End of CII addition

\setlength{\parskip}{0pt}

% Added by CII

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\Acknowledgements{

}

\Dedication{

}

\Preface{

}

\Abstract{
The preface pretty much says it all.

\par

Second paragraph of abstract starts here.
}

	\usepackage{booktabs}
 \usepackage{longtable}
 \usepackage{array}
 \usepackage{multirow}
 \usepackage{wrapfig}
 \usepackage{float}
 \usepackage{colortbl}
 \usepackage{pdflscape}
 \usepackage{tabu}
 \usepackage{threeparttable}
 \usepackage{threeparttablex}
 \usepackage[normalem]{ulem}
 \usepackage{makecell}
 \usepackage{xcolor}
% End of CII addition
%%
%% End Preamble
%%
%
\begin{document}

% Everything below added by CII
  \maketitle

\frontmatter % this stuff will be roman-numbered
\pagestyle{empty} % this removes page numbers from the frontmatter

  \hypersetup{linkcolor=black}
  \setcounter{tocdepth}{2}
  \tableofcontents

  \listoftables

  \listoffigures
  \begin{abstract}
    The preface pretty much says it all.

    \par

    Second paragraph of abstract starts here.
  \end{abstract}

\mainmatter % here the regular arabic numbering starts
\pagestyle{fancyplain} % turns page numbering back on

\hypertarget{introduction}{%
\chapter*{Introduction}\label{introduction}}
\addcontentsline{toc}{chapter}{Introduction}
\begin{itemize}
\item
  Importance of sound systematic reviews (meta)
\item
  Why it was built
\item
  Advantages over other types
\item
  Whereas first results look promising, there hasn't been conducted a full proof of concept
\end{itemize}
Goals:
- Save time
- But also more systematic..?

? Misses

\appendix

\hypertarget{parameter-configurations}{%
\chapter{Parameter Configurations}\label{parameter-configurations}}

ASReview takes the following parameters/arguments:
\begin{itemize}
\tightlist
\item
  number of training data (included/excluded)
\item
  a model
\item
  a query strategy
\item
  a balance strategy (fixed)
\item
  feature extraction
\end{itemize}
\begin{table}[H]
\centering
\begin{tabular}{l>{\raggedright\arraybackslash}p{30em}}
\toprule
  & Configurations\\
\midrule
Models & Naive Bayes, Random Forest, Support Vector Machine, Logistic Regression\\
Query Strategies & Cluster Sampling, Maximum Sampling, Cluster * Maximum  Sampling, Maximum * Uncertainty Sampling, Maximum * Random Sampling, Cluster * Uncertainty Sampling, Cluster * Random Sampling\\
Feature extraction strategies & Doc2Vec, tf-idf, sbert, embeddingIdf\\
Training data [included/excluded] & 10/10, 5/5, 5/10\\
\bottomrule
\end{tabular}
\end{table}
This leads to 273 combinations of configurations.
\begin{itemize}
\tightlist
\item
  Naive bayes only goes with tfidf feature extraction.
\item
  For the feature extraction strategies we will focus on doc2vec and tfidf. (but will compute all 4)
\item
  This leads to 3 * 7 * 4 * 3 + 1 * 7 * 1 * 3 = 273 combinations.
\end{itemize}
\begin{longtable}{llll}
\toprule
Model & Query Strategy & Feature extraction strategy & Training data [included/excluded]\\
\midrule
\endfirsthead
\multicolumn{4}{@{}l}{\textit{(continued)}}\\
\toprule
Model & Query Strategy & Feature extraction strategy & Training data [included/excluded]\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
nb & cluster & tfidf & 10/10\\
nb & max & tfidf & 10/10\\
nb & max * cluster & tfidf & 10/10\\
nb & max * uncertainty & tfidf & 10/10\\
nb & max * random & tfidf & 10/10\\
\addlinespace
nb & cluster * uncertainty & tfidf & 10/10\\
nb & cluster * random & tfidf & 10/10\\
nb & cluster & tfidf & 5/5\\
nb & max & tfidf & 5/5\\
nb & max * cluster & tfidf & 5/5\\
\addlinespace
nb & max * uncertainty & tfidf & 5/5\\
nb & max * random & tfidf & 5/5\\
nb & cluster * uncertainty & tfidf & 5/5\\
nb & cluster * random & tfidf & 5/5\\
nb & cluster & tfidf & 5/10\\
\addlinespace
nb & max & tfidf & 5/10\\
nb & max * cluster & tfidf & 5/10\\
nb & max * uncertainty & tfidf & 5/10\\
nb & max * random & tfidf & 5/10\\
nb & cluster * uncertainty & tfidf & 5/10\\
\addlinespace
nb & cluster * random & tfidf & 5/10\\
rf & cluster & doc2vec & 10/10\\
rf & max & doc2vec & 10/10\\
rf & max * cluster & doc2vec & 10/10\\
rf & max * uncertainty & doc2vec & 10/10\\
\addlinespace
rf & max * random & doc2vec & 10/10\\
rf & cluster * uncertainty & doc2vec & 10/10\\
rf & cluster * random & doc2vec & 10/10\\
rf & cluster & doc2vec & 5/5\\
rf & max & doc2vec & 5/5\\
\addlinespace
rf & max * cluster & doc2vec & 5/5\\
rf & max * uncertainty & doc2vec & 5/5\\
rf & max * random & doc2vec & 5/5\\
rf & cluster * uncertainty & doc2vec & 5/5\\
rf & cluster * random & doc2vec & 5/5\\
\addlinespace
rf & cluster & doc2vec & 5/10\\
rf & max & doc2vec & 5/10\\
rf & max * cluster & doc2vec & 5/10\\
rf & max * uncertainty & doc2vec & 5/10\\
rf & max * random & doc2vec & 5/10\\
\addlinespace
rf & cluster * uncertainty & doc2vec & 5/10\\
rf & cluster * random & doc2vec & 5/10\\
rf & cluster & tfidf & 10/10\\
rf & max & tfidf & 10/10\\
rf & max * cluster & tfidf & 10/10\\
\addlinespace
rf & max * uncertainty & tfidf & 10/10\\
rf & max * random & tfidf & 10/10\\
rf & cluster * uncertainty & tfidf & 10/10\\
rf & cluster * random & tfidf & 10/10\\
rf & cluster & tfidf & 5/5\\
\addlinespace
rf & max & tfidf & 5/5\\
rf & max * cluster & tfidf & 5/5\\
rf & max * uncertainty & tfidf & 5/5\\
rf & max * random & tfidf & 5/5\\
rf & cluster * uncertainty & tfidf & 5/5\\
\addlinespace
rf & cluster * random & tfidf & 5/5\\
rf & cluster & tfidf & 5/10\\
rf & max & tfidf & 5/10\\
rf & max * cluster & tfidf & 5/10\\
rf & max * uncertainty & tfidf & 5/10\\
\addlinespace
rf & max * random & tfidf & 5/10\\
rf & cluster * uncertainty & tfidf & 5/10\\
rf & cluster * random & tfidf & 5/10\\
rf & cluster & sbert & 10/10\\
rf & max & sbert & 10/10\\
\addlinespace
rf & max * cluster & sbert & 10/10\\
rf & max * uncertainty & sbert & 10/10\\
rf & max * random & sbert & 10/10\\
rf & cluster * uncertainty & sbert & 10/10\\
rf & cluster * random & sbert & 10/10\\
\addlinespace
rf & cluster & sbert & 5/5\\
rf & max & sbert & 5/5\\
rf & max * cluster & sbert & 5/5\\
rf & max * uncertainty & sbert & 5/5\\
rf & max * random & sbert & 5/5\\
\addlinespace
rf & cluster * uncertainty & sbert & 5/5\\
rf & cluster * random & sbert & 5/5\\
rf & cluster & sbert & 5/10\\
rf & max & sbert & 5/10\\
rf & max * cluster & sbert & 5/10\\
\addlinespace
rf & max * uncertainty & sbert & 5/10\\
rf & max * random & sbert & 5/10\\
rf & cluster * uncertainty & sbert & 5/10\\
rf & cluster * random & sbert & 5/10\\
rf & cluster & embeddingIdf & 10/10\\
\addlinespace
rf & max & embeddingIdf & 10/10\\
rf & max * cluster & embeddingIdf & 10/10\\
rf & max * uncertainty & embeddingIdf & 10/10\\
rf & max * random & embeddingIdf & 10/10\\
rf & cluster * uncertainty & embeddingIdf & 10/10\\
\addlinespace
rf & cluster * random & embeddingIdf & 10/10\\
rf & cluster & embeddingIdf & 5/5\\
rf & max & embeddingIdf & 5/5\\
rf & max * cluster & embeddingIdf & 5/5\\
rf & max * uncertainty & embeddingIdf & 5/5\\
\addlinespace
rf & max * random & embeddingIdf & 5/5\\
rf & cluster * uncertainty & embeddingIdf & 5/5\\
rf & cluster * random & embeddingIdf & 5/5\\
rf & cluster & embeddingIdf & 5/10\\
rf & max & embeddingIdf & 5/10\\
\addlinespace
rf & max * cluster & embeddingIdf & 5/10\\
rf & max * uncertainty & embeddingIdf & 5/10\\
rf & max * random & embeddingIdf & 5/10\\
rf & cluster * uncertainty & embeddingIdf & 5/10\\
rf & cluster * random & embeddingIdf & 5/10\\
\addlinespace
svm & cluster & doc2vec & 10/10\\
svm & max & doc2vec & 10/10\\
svm & max * cluster & doc2vec & 10/10\\
svm & max * uncertainty & doc2vec & 10/10\\
svm & max * random & doc2vec & 10/10\\
\addlinespace
svm & cluster * uncertainty & doc2vec & 10/10\\
svm & cluster * random & doc2vec & 10/10\\
svm & cluster & doc2vec & 5/5\\
svm & max & doc2vec & 5/5\\
svm & max * cluster & doc2vec & 5/5\\
\addlinespace
svm & max * uncertainty & doc2vec & 5/5\\
svm & max * random & doc2vec & 5/5\\
svm & cluster * uncertainty & doc2vec & 5/5\\
svm & cluster * random & doc2vec & 5/5\\
svm & cluster & doc2vec & 5/10\\
\addlinespace
svm & max & doc2vec & 5/10\\
svm & max * cluster & doc2vec & 5/10\\
svm & max * uncertainty & doc2vec & 5/10\\
svm & max * random & doc2vec & 5/10\\
svm & cluster * uncertainty & doc2vec & 5/10\\
\addlinespace
svm & cluster * random & doc2vec & 5/10\\
svm & cluster & tfidf & 10/10\\
svm & max & tfidf & 10/10\\
svm & max * cluster & tfidf & 10/10\\
svm & max * uncertainty & tfidf & 10/10\\
\addlinespace
svm & max * random & tfidf & 10/10\\
svm & cluster * uncertainty & tfidf & 10/10\\
svm & cluster * random & tfidf & 10/10\\
svm & cluster & tfidf & 5/5\\
svm & max & tfidf & 5/5\\
\addlinespace
svm & max * cluster & tfidf & 5/5\\
svm & max * uncertainty & tfidf & 5/5\\
svm & max * random & tfidf & 5/5\\
svm & cluster * uncertainty & tfidf & 5/5\\
svm & cluster * random & tfidf & 5/5\\
\addlinespace
svm & cluster & tfidf & 5/10\\
svm & max & tfidf & 5/10\\
svm & max * cluster & tfidf & 5/10\\
svm & max * uncertainty & tfidf & 5/10\\
svm & max * random & tfidf & 5/10\\
\addlinespace
svm & cluster * uncertainty & tfidf & 5/10\\
svm & cluster * random & tfidf & 5/10\\
svm & cluster & sbert & 10/10\\
svm & max & sbert & 10/10\\
svm & max * cluster & sbert & 10/10\\
\addlinespace
svm & max * uncertainty & sbert & 10/10\\
svm & max * random & sbert & 10/10\\
svm & cluster * uncertainty & sbert & 10/10\\
svm & cluster * random & sbert & 10/10\\
svm & cluster & sbert & 5/5\\
\addlinespace
svm & max & sbert & 5/5\\
svm & max * cluster & sbert & 5/5\\
svm & max * uncertainty & sbert & 5/5\\
svm & max * random & sbert & 5/5\\
svm & cluster * uncertainty & sbert & 5/5\\
\addlinespace
svm & cluster * random & sbert & 5/5\\
svm & cluster & sbert & 5/10\\
svm & max & sbert & 5/10\\
svm & max * cluster & sbert & 5/10\\
svm & max * uncertainty & sbert & 5/10\\
\addlinespace
svm & max * random & sbert & 5/10\\
svm & cluster * uncertainty & sbert & 5/10\\
svm & cluster * random & sbert & 5/10\\
svm & cluster & embeddingIdf & 10/10\\
svm & max & embeddingIdf & 10/10\\
\addlinespace
svm & max * cluster & embeddingIdf & 10/10\\
svm & max * uncertainty & embeddingIdf & 10/10\\
svm & max * random & embeddingIdf & 10/10\\
svm & cluster * uncertainty & embeddingIdf & 10/10\\
svm & cluster * random & embeddingIdf & 10/10\\
\addlinespace
svm & cluster & embeddingIdf & 5/5\\
svm & max & embeddingIdf & 5/5\\
svm & max * cluster & embeddingIdf & 5/5\\
svm & max * uncertainty & embeddingIdf & 5/5\\
svm & max * random & embeddingIdf & 5/5\\
\addlinespace
svm & cluster * uncertainty & embeddingIdf & 5/5\\
svm & cluster * random & embeddingIdf & 5/5\\
svm & cluster & embeddingIdf & 5/10\\
svm & max & embeddingIdf & 5/10\\
svm & max * cluster & embeddingIdf & 5/10\\
\addlinespace
svm & max * uncertainty & embeddingIdf & 5/10\\
svm & max * random & embeddingIdf & 5/10\\
svm & cluster * uncertainty & embeddingIdf & 5/10\\
svm & cluster * random & embeddingIdf & 5/10\\
lr & cluster & doc2vec & 10/10\\
\addlinespace
lr & max & doc2vec & 10/10\\
lr & max * cluster & doc2vec & 10/10\\
lr & max * uncertainty & doc2vec & 10/10\\
lr & max * random & doc2vec & 10/10\\
lr & cluster * uncertainty & doc2vec & 10/10\\
\addlinespace
lr & cluster * random & doc2vec & 10/10\\
lr & cluster & doc2vec & 5/5\\
lr & max & doc2vec & 5/5\\
lr & max * cluster & doc2vec & 5/5\\
lr & max * uncertainty & doc2vec & 5/5\\
\addlinespace
lr & max * random & doc2vec & 5/5\\
lr & cluster * uncertainty & doc2vec & 5/5\\
lr & cluster * random & doc2vec & 5/5\\
lr & cluster & doc2vec & 5/10\\
lr & max & doc2vec & 5/10\\
\addlinespace
lr & max * cluster & doc2vec & 5/10\\
lr & max * uncertainty & doc2vec & 5/10\\
lr & max * random & doc2vec & 5/10\\
lr & cluster * uncertainty & doc2vec & 5/10\\
lr & cluster * random & doc2vec & 5/10\\
\addlinespace
lr & cluster & tfidf & 10/10\\
lr & max & tfidf & 10/10\\
lr & max * cluster & tfidf & 10/10\\
lr & max * uncertainty & tfidf & 10/10\\
lr & max * random & tfidf & 10/10\\
\addlinespace
lr & cluster * uncertainty & tfidf & 10/10\\
lr & cluster * random & tfidf & 10/10\\
lr & cluster & tfidf & 5/5\\
lr & max & tfidf & 5/5\\
lr & max * cluster & tfidf & 5/5\\
\addlinespace
lr & max * uncertainty & tfidf & 5/5\\
lr & max * random & tfidf & 5/5\\
lr & cluster * uncertainty & tfidf & 5/5\\
lr & cluster * random & tfidf & 5/5\\
lr & cluster & tfidf & 5/10\\
\addlinespace
lr & max & tfidf & 5/10\\
lr & max * cluster & tfidf & 5/10\\
lr & max * uncertainty & tfidf & 5/10\\
lr & max * random & tfidf & 5/10\\
lr & cluster * uncertainty & tfidf & 5/10\\
\addlinespace
lr & cluster * random & tfidf & 5/10\\
lr & cluster & sbert & 10/10\\
lr & max & sbert & 10/10\\
lr & max * cluster & sbert & 10/10\\
lr & max * uncertainty & sbert & 10/10\\
\addlinespace
lr & max * random & sbert & 10/10\\
lr & cluster * uncertainty & sbert & 10/10\\
lr & cluster * random & sbert & 10/10\\
lr & cluster & sbert & 5/5\\
lr & max & sbert & 5/5\\
\addlinespace
lr & max * cluster & sbert & 5/5\\
lr & max * uncertainty & sbert & 5/5\\
lr & max * random & sbert & 5/5\\
lr & cluster * uncertainty & sbert & 5/5\\
lr & cluster * random & sbert & 5/5\\
\addlinespace
lr & cluster & sbert & 5/10\\
lr & max & sbert & 5/10\\
lr & max * cluster & sbert & 5/10\\
lr & max * uncertainty & sbert & 5/10\\
lr & max * random & sbert & 5/10\\
\addlinespace
lr & cluster * uncertainty & sbert & 5/10\\
lr & cluster * random & sbert & 5/10\\
lr & cluster & embeddingIdf & 10/10\\
lr & max & embeddingIdf & 10/10\\
lr & max * cluster & embeddingIdf & 10/10\\
\addlinespace
lr & max * uncertainty & embeddingIdf & 10/10\\
lr & max * random & embeddingIdf & 10/10\\
lr & cluster * uncertainty & embeddingIdf & 10/10\\
lr & cluster * random & embeddingIdf & 10/10\\
lr & cluster & embeddingIdf & 5/5\\
\addlinespace
lr & max & embeddingIdf & 5/5\\
lr & max * cluster & embeddingIdf & 5/5\\
lr & max * uncertainty & embeddingIdf & 5/5\\
lr & max * random & embeddingIdf & 5/5\\
lr & cluster * uncertainty & embeddingIdf & 5/5\\
\addlinespace
lr & cluster * random & embeddingIdf & 5/5\\
lr & cluster & embeddingIdf & 5/10\\
lr & max & embeddingIdf & 5/10\\
lr & max * cluster & embeddingIdf & 5/10\\
lr & max * uncertainty & embeddingIdf & 5/10\\
\addlinespace
lr & max * random & embeddingIdf & 5/10\\
lr & cluster * uncertainty & embeddingIdf & 5/10\\
lr & cluster * random & embeddingIdf & 5/10\\*
\end{longtable}
\hypertarget{list-of-definitions}{%
\chapter{List of Definitions}\label{list-of-definitions}}

Balance Strategy
\begin{description}
\item[Model.]
The prediction model for Active learning
\begin{verbatim}
some code, part of Definition 2
\end{verbatim}
Third paragraph of definition 2.
\item[Query Strategy]
The way papers are queried to the researcher
\begin{verbatim}
query_strategy
\end{verbatim}
\begin{description}
\tightlist
\item[Uncertainty sampling]
Selects the least sure instances for labelling.
\item[Maximum sampling]
Selects the samples with the highest prediction probability.
\item[Random and maximum sampling.]
Combination of random and maximum sampling. By default samples the 95\% of the instances with max sampling, and 5\% of the samples with random sampling.
\end{description}
\item[Feature Extraction Strategy]
Feature Extraction Strategy
\begin{verbatim}
feature_extraction
\end{verbatim}
Third paragraph of definition x.
\begin{description}
\tightlist
\item[Doc2Vec]
\item[Term frequency-inverse document frequency (tf-idf)]
\end{description}
sbert
\end{description}
\backmatter

\hypertarget{references}{%
\chapter*{References}\label{references}}
\addcontentsline{toc}{chapter}{References}

\markboth{References}{References}

\noindent

\setlength{\parindent}{-0.20in}
\setlength{\leftskip}{0.20in}
\setlength{\parskip}{8pt}

% Index?

\end{document}