### 2020-02-04 manuscript/index/05-appendix.Rmd
add model definitions

`r if(knitr:::is_latex_output()) '\\appendix'`

`r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'`

<!--
If you feel it necessary to include an appendix, it goes here.
-->

# Parameter Configurations
```{r}
# load model parameters 
source("R/asr-parameters.R")
source("R/packages.R")
```

```{r}
# configurations table 
tab <- lapply(words, FUN = function(x) paste(x, collapse = ", "))
tab <- do.call(rbind, tab)
colnames(tab) <- c("Configurations")

```

```{r}
# combinations table
combins <- expand.grid(params) %>%
  arrange(model, feature_extraction)

# rows to drop
droprows <- combins$model == "nb" & combins$feature_extraction != "tfidf"
combins <- combins[!droprows, ]
```

model: nb
query_strategy: max
balance_strategy: double
feature_extraction: tfidf
n_instances: 1
n_queries: 10
n_prior_included: 5
n_prior_excluded: 5
mode: simulate
model_param: {}
query_param: {}
feature_param: {}
balance_param: {}
abstract_only: False

```{r}
kable(tab, "latex", booktabs = TRUE) %>%
  kable_styling(full_width = FALSE) %>%
  column_spec(2, width = "30em") 
```

This leads to  `r nrow(combins)` combinations of configurations.

- Naive bayes only goes with tfidf feature extraction.
- For the feature extraction strategies we will focus on doc2vec and tfidf. (but will compute all 4)
- This leads to 3 * 7  * 4 * 3 + 1 * 7 * 1 * 3 =  273 combinations.

```{r}
names(combins) <- c("Model", "Query Strategy", "Feature extraction strategy", "Training data [included/excluded]")
combins %>%
  kable(row.names = FALSE, format = "latex", longtable = TRUE, booktabs = TRUE) %>%
  kable_styling(latex_options = c("repeat_header")) 
```

# List of Definitions

Model.

:   The prediction model for Active learning

        some code, part of Definition 2

    Third paragraph of definition 2.

Query Strategy 
:   Type of active learning, determining the way papers are queried to the researcher.

        query_strategy

    Uncertainty sampling
    :  Selects the least sure instances for labelling.

    Maximum sampling
    :  Selects the samples with the highest prediction probability. The references who are most likely to be included are presented first.

    Random and maximum sampling. 
    :  Combination of random and maximum sampling. By default samples the 95% of the instances with max sampling, and 5% of the samples with random sampling.

Feature Extraction Strategy
:   The way the words in the abstracts are represented in the statistical model.

        feature_extraction

    Doc2Vec
    :

        doc2vec

    Term frequency-inverse document frequency (tf-idf)
    :

        sbert

<!-- # The Last Appendix -->

<!-- This first appendix includes all of the R chunks of code that were hidden throughout the document (using the `include = FALSE` chunk tag) to help with readibility and/or setup. -->

<!-- **In the main Rmd file** -->

<!-- ```{r ref.label='include_packages', results='hide', echo = TRUE} -->
<!-- ``` -->

<!-- **In Chapter \@ref(ref-labels):** -->

<!-- ```{r ref.label='include_packages_2', results='hide', echo = TRUE} -->
<!-- ``` -->