### 2020-02-04 manuscript/index/index.Rmd

---
author: 'Gerbrich Ferdinands'
date: 'May 2020'
institution: 'Utrecht University'
division: 'Faculty of Social Sciences'
advisor: 'prof. dr. Rens van de Schoot'
# If you have more two advisors, un-silence line 7
altadvisor: 'Jonathan de Bruin, Raoul Schram'
department: 'Methodology and Statistics'
degree: 'Master of Science'
title: 'My thesis'
knit: "bookdown::render_book"
site: bookdown::bookdown_site
output: 
#  thesisdown::thesis_pdf: default  
  thesisdown::thesis_gitbook: default # creates html book in _book folder.  
#  thesisdown::thesis_word: default
#  thesisdown::thesis_epub: default
# If you are creating a PDF you'll need to write your preliminary content (e.g., abstract, acknowledgements) here or
# use code similar to line 22-23 for the .RMD files. If you are NOT producing a PDF, you can delete or silence lines 21-32 in this YAML header.
abstract: |
  `r if(knitr:::is_latex_output()) paste(readLines("00-abstract.Rmd"), collapse = '\n  ')`
# If you'd rather include the preliminary content in files instead of inline
# like below, use a command like that for the abstract above.  Note that a tab is 
# needed on the line after the `|`.
#acknowledgements: |
#  I want to thank a few people.
#dedication: |
#  You can have a dedication here if you wish. 
#preface: |
#  This is an example of a thesis setup to use the reed thesis document class
#  (for LaTeX) and the R bookdown package, in general.
# Specify the location of the bibliography below
bibliography: bib/asreview.bib
# Download your specific csl file and refer to it in the line below.
csl: csl/apa.csl
lot: true
lof: true
# If you prefer blank lines between paragraphs, un-silence lines  40-41 (this requires package tikz)
#header-includes:
#- \usepackage{tikz}
---

<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of metadata used to produce the document.  Be careful with spacing in this header!

If you'd prefer to not include a Dedication, for example, simply delete the section entirely, or silence (add #) them.

If you have other LaTeX packages you would like to include, delete the # before header-includes and list the packages after hyphens on new lines.

If you'd like to include a comment that won't be produced in your resulting file enclose it in a block like this.

If you receive a duplicate label error after knitting, make sure to delete the index.Rmd file and then knit again.
-->

```{r include_packages, include = FALSE}
# This chunk ensures that the thesisdown package is
# installed and loaded. This thesisdown package includes
# the template files for the thesis.
if(!require(devtools))
  install.packages("devtools", repos = "http://cran.rstudio.com")
if(!require(thesisdown))
  devtools::install_github("ismayc/thesisdown")
library(thesisdown)

knitr::opts_chunk$set(echo = FALSE)

source("R/packages.R")
```

<!-- On ordering the chapter files:
There are two options:
1. Name your chapter files in the order in which you want them to appear (e.g., 01-Inro, 02-Data, 03-Conclusions). 
2. Otherwise, you can specify the order in which they appear in the _bookdown.yml (for PDF only).

Do not include 00(two-hyphens)prelim.Rmd and 00-abstract.Rmd in the YAML file--they are handled in the YAML above differently for the PDF version.
-->

<!-- The {.unnumbered} option here means that the introduction will be "Chapter 0." You can also use {-} for no numbers
on chapters.
-->

# Introduction {.unnumbered}
ASReview takes the following parameters/arguments:

- a model
- a query strategy
- a balance strategy (fixed)
- a feature extraction strategy
- number of training data

The model is typically a learning algorithm used to predict the relevance of text. The following models will be compared:

* Naive Bayes (*nb*) [@Zhang2004]. 
*