### 2020-02-18 manuscript/drafts/analyseplan.Rmd
updated analysis plan

---
title: "analyseplan"
author: "Gerbrich Ferdinands"
date: "2/13/2020"
output: pdf_document
---
Goal: evaluate performance of different models of the ASReview tool.

### Stage 1: hyperparameter optimization
We are testing 5 models on 5 different datasets.

Every model has its own set of hyperparameters.
The hyperparameters are optimized on the 5 datasets in three different ways:

- 1 on 1: maximum performance 
- 4 on 1: cross-validation 
- 5 on 1: more data = more better?

This results $(5+5+1)*5$ sets of hyperparameters.

### Stage 2: simulation
for every for every model (5), for every dataset (5) and for every set of optimized hyperparameters (3), a simulation study is performed. From these $5*5*3=75$ simulation studies, performance of the different models is evaluated.

### Datasets

#### ptsd
#### ace
#### hall

#### nagtegaal - van PhD van Lars
#### medische van Jan

### Models
(input table here?)

- Naive Bayes 
- Random Forests
- Support Vecor Machine (doc2vec)
- Logistic Regression
- Dense Neural Network (doc2vec)

The other parameters remain fixed over the 5 models:

- Feature extraction = tf-idf
- Query Strategy = max
- Balance Strategy = triple
- n_instances=10 (number of papers each query)
- n_prior_included = 5
- n_prior_excluded = 5
- mix_ratio = 0.95 (95% max, 5% random)

## Hyperparameters

## Scripts 
  - simulatie:
- [ ] script die optimalisatie genereert (config files.)
- [ ] script die `.ini`'s voor simulatie genereert.

```{bash}
output/
├── active_learning
│   ├── nb_max_double_tfidf
│   │   └── depression_hall_ace_ptsd_nagtegaal
│   │       ├── best
│   │       │   ├── ace
│   │       │   ├── depression
│   │       │   ├── hall
│   │       │   ├── nagtegaal
│   │       │   └── ptsd
│   │       ├── current
│   │       │   ├── ace
│   │       │   ├── depression
│   │       │   ├── hall
│   │       │   ├── nagtegaal
│   │       │   └── ptsd
│   │       └── trials.pkl
```