### 2020-02-25 manuscript/test/index.Rmd
update manuscript   - update notes   - update bookdown files

--- 
title: "A Minimal Book Example"
author: "Yihui Xie"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
---

# thesis Ferdinands
Goal: evaluate performance of different models of the ASReview tool.

### Stage 1: hyperparameter optimization
We are testing 5 models on 5 different datasets.

Every model has its own set of hyperparameters.
The hyperparameters are optimized on the 5 datasets in three different ways:

- 1 on 1: maximum performance 
- 4 on 1: cross-validation 
- 5 on 1: more data = more better?

This results $(5+5+1)*5$ sets of hyperparameters.

### Stage 2: simulation
for every for every model (5), for every dataset (5) and for every set of optimized hyperparameters (3), a simulation study is performed. From these $5*5*3=75$ simulation studies, performance of the different models is evaluated.

## Scripts 
  - simulatie:
- [ ] script die optimalisatie genereert (config files.)
- [ ] script die `.ini`'s voor simulatie genereert.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')

# load all necessary packages
source("R/packages.R")

# path to all R folder
pathRdata <- "../../R/data/"

# set chunk default options knitr
knitr::opts_chunk$set(echo = FALSE)

```