### 2020-03-13 manuscript/drafts/manuscript.tex
introduce bug

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Manuscript drafts},
  pdfauthor={Gerbrich Ferdinands},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage[]{biblatex}
\addbibresource{asreview.bib}

\title{Manuscript drafts}
\author{Gerbrich Ferdinands}
\date{1/14/2020}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\paragraph{Introduction}\label{introduction}}

\hypertarget{methods}{%
\paragraph{Methods}\label{methods}}

A convenience sample of 5 existing systematic reviews on varying topics
was collected.

\hypertarget{results}{%
\paragraph{Results}\label{results}}

\hypertarget{discussion}{%
\paragraph{Discussion}\label{discussion}}

\newpage

\hypertarget{introduction-1}{%
\section{Introduction}\label{introduction-1}}

Systematic Reviews (SR's) are booming - but they are a lot of work
Various machine learning tools have been proposed to reduce workload in
abstract screening.

\begin{itemize}
\tightlist
\item
  objectives - to demonstrate effectiveness of ml algorithm in reducing
  abstract classification for systematic reviews
\item
  justification -
\item
  background
\item
  guidance to reader
\item
  summary/conclusion
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

This study is about how machine learning algorithms can increase
efficiency in systematic reviews. I will write about what SRs are, and
how workload can be reduced.

Systematic reviews are top of the bill in research. As more and more
papers are published and reproducibility crisis has emerged, science
calls for more meta. It is important reflect on research by giving an
overview of research areas which is typically done by a systematic
review {[}\ldots{]}. Performing a systematic review is a tedious and
time-consuming task. To review a specific research area, one starts out
with an initial search of thousands of academic papers. All these papers
abstracts need to be screened to find an initial batch of possibly
relevant papers. With now hopefully only a couple of hundred papers
left, the researcher needs to read these papers full-text to arrive at a
final selection of papers that are relevant for the final systematic
review {[}this is prisma process?{]}. This whole processes costs this
and this much time {[}shelmilt{]}.

The stage of abstract screening where abstracts are systematically
screened is where a lot is to be gained. This stage is the target of
possible learning algorithms that can assist the reviewer in selecting
the relevant papers. Together with the reviewer /human machine
interaction. The algorithm aims to compute which papers in the pool need
to be excluded and which need to be included, based on the reviewers
decisions. It learns from the reviewers decisions and asks the reviewer
to provide more labels, incrementally improving its class predictions.

The goal of the algorithm defined in the current study is to reduce to
number of abstracts needed to screen (maybe not right term, bit
biomedical). To be more specific, the algorithm aims to present the
reader with the primary studies as soon as possible. This means that at
some point you probably have seen all relevant abstracts and are only
viewing excluded papers, which means you can stop reviewing much earlier
(theoretically spoken). Also reviewing is now much more fun. As compared
to when you have to review all abstracts and you perhaps see only one
relevant abstract every other week/day.

So you might wonder, how does such an algorithm actually work? Active
learning strategy, starting with a pool of unlabeled abstracts (U). The
reviewer starts labeling some instances in U, creating L. The algorithm
utilizes L to predict labels for all abstracts (classifier), by using a
set of features from the papers called X, for example the text in the
abstract (feature extraction method). Now, it made an initial
classification. The algorithm now aims to improve its classification by
which paper from U will be presented to the reviewer next. By labeling
the next paper, the reviewer provides the algorithm with new information
which the algorithm uses to update its prediction.

We approach asr as a classification problem: All papers obtained in the
systematic search form a pool of instances x. All instances x need to be
classified, e.g.~we want to give them a label y. all x are now part of
U,. Binary classification, either inclusion or exclusion. We want to
classify based on some features from the instances \(x\), feature vector
\(\mathbf{X}\).

By starting of with L, the algorithm utilizes characteristics of X to
predict labels in U.

We want to classify the papers in Where the whole collection of papers
in the systematic search form a pool of instances x, with unknown label
y. \textless x,y\textgreater.

\begin{verbatim}
input: a pool of unlabeled abstracts $\mathcal{U}$ 
oracle labels a few initial papers, from $\mathcal{U}$, $\mathcal{L}$,

  M = train($\mathcal{L}$)
  query x \isin

\end{verbatim}

\begin{itemize}
\item
  pool unlabeled abstracts \(\mathcal{U}\)
\item
  labeled data set \(\mathcal{L}\),
\item
  instance \(x\), label \(y\)
\item
  utility measure \(\phi_A(\cdot)\)
\item
  \(x^*_A\) best query instance according to \(\phi_A(\cdot)\)
\end{itemize}

Now there are a few technical details. Many different versions of such
algorithms exist. Many of such algorithms have been described in the
active learning literature and have been applied in the systematic
reviewing process. Not exhaustive, but the algorithm can apply many
different strategies to arrive at its predictions, which can be divided
in following parameters: classifier, feature extraction strategy,
balancing, query strategy.

Most often the SVM classifier is used, popular and very good results.
Also lots of other configurations. However, other classifiers have not
been tested a lot (polygon thing by cohe, naïve bayes and random forest
by \ldots), but mostly SVM still. Also, most research in the medical
sciences (well there are some exceptions of course {[}conversation
between cohen and matwill{]}

In the current study, we aim at exploring the performance of several
classifiers on reducing workload while maintaining performance in
abstract screening process. This is done by performing simulations on
existing systematic reviews. Performance is evaluated by how much time
can be saved \ldots{} while maintaining accuracy.

Present research questions. RQ1 -- which classifiers perform best? -
RQ1a -- does classifier performance vary over different research areas?
(in what terms does it perform best)

RQ2 different hyperparameter optimizations? Classifiers come with
hyperparameters. We have to make a choice on how to set these
hyperparameters. What we do is create 3 sets, optimizing in three ways,
aggregate results obtained.

The goal is to gain insight in classifiers other than the widely applied
SVM, overall various research areas. So not only medical sciences.

To perform all these computations the research was carried out using the
ASReview software by Utrecht University, which has a simulation mode
that you can just input your labelled review file into and perform a
simulation study with it. To be found on GitHub. It has many adjustable
components/is very versatile.

Then if we still have time left we explore the effect of another feature
extraction method, namely doc2vec. This might possibly increase
performance as it performs better at grasping structure/hidden
relations/hierarchy between words in the texts. So, it could be
interesting in more `fuzzy' research areas. A downside is that it takes
more computing time.

Then if there's even more time we might want to compare a different
balance strategy.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

A SR can be divided into phases. Everything starts with a
\textbf{systematic search}, leading to then citation screening is
performed, then full-text screening \autocite{PRISMA-PGroup2015}

What must be the objective of our tool?: It is the tedious task citation
screening part where loads of time can be saved.

models are designed in a `realistic' way (you have some inclusions)

Selecting papers is a two-step process: abstract \& fulltext screening

Binary classification problem. predict whether a paper in the pool is an
inclusion or exclusion, based on labeled instances from \(\mathcal{L}\)
and use training data to understand how input variables are related to
class.

We're building active learning model who takes X as an input an predicts
class using labels from training set. Model improves by deciding asking
more information from the reviewer (oracle), accounting for imbalance.

\hypertarget{assumptions}{%
\paragraph{Assumptions}\label{assumptions}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  decisions of the original SR are \textbf{ground truth} (benchmark)
  (oracle)
\end{enumerate}

The inclusion rate is \ldots{} data is imbalanced. what is the
philosophy False negatives must be avoided \ldots{} The cost of a false
negative outweighs the cost of a false positive. Note that we assume the
oracle/original user to hold the truth. This is of course not always the
case.

There are two classes in the data: exlusions and inclusions. The
inclusions are clearly the minority class.

Datasets from the medical and social sciences, software engineering and
public administration. Medical sciences SRs are viewed as more
`strict'/`structured' and social sciences more messy.

\hypertarget{active-learning-for-systematic-reviews}{%
\subsection{active learning for systematic
reviews}\label{active-learning-for-systematic-reviews}}

corpus = all the text:

Active learning = increasing classification performance with every
query. The query strategy determines the way unlabeled papers are
queried to the researcher.

\autocite{modAL2018}

RQ1 - what are good classifiers RQ2 - what are good optimization
strategies

\hypertarget{background}{%
\subsection{Background}\label{background}}

\autocite{OMara-Eves2015} literature review

\autocite{Yu2018a}, \autocite{Yu2019} simulated 32 svm classifiers, on
software engineering. A popular classifier is SVM. succes with HUTM
(fastread), uncertainty, mix of weighting and agressive undersampling,
In terms of Yu et al, we adopt .CT.

SVM - tf-idf on medical data, uncertainty sampling, agressive
undersampling. \autocite{Wallace2010}

abstrackr

SVM + Weighting + uncertainty (bow) produced good methods
\autocite{Miwa2014} Also include social sciences data besides medical
data.

\autocite{Cohen2006} perceptron-based classifier (neural network)

SVM on legal documents (no balancing, certainty ) \autocite{Cormack2014}
in limitations section mentions that LR yields about same results, nb
inferior results.

\autocite{Kilicoglu2009} - SVM, naive bayes, boosting and combinations.
future work should optimize parameters. ``Regarding the base classifiers
used in identifying method- ologically rigorous studies, boosting
consistently strikes the best balance between precision and recall,
whereas naive Bayes in general performs well on recall (demonstrating a
tradeoff between recall and precision), as does polynomial SVM on
precision. The AUC results are mixed, although boosting has a slight
edge overall. These results demonstrate that different classifiers can
be used to satisfy different information needs (SVM for specificity,
naive Bayes for sensitivity, and boosting for balance between the two,
for example).''

Our extensions is that we try different classifiers, on more datasets.

This study was approved by the Ethics Committee of the Faculty of Social
and Behavioural Sciences of Utrecht University, filed as an amendement
under study 20-104.

All simulations were run using through cartesius EINF-156

\newpage

\hypertarget{methods-1}{%
\section{Methods}\label{methods-1}}

Goal: evaluate performance of different models of the ASReview tool. The
screening process is simulated using ASReview, seeing if the original
inclusions replicate. What would happen if the citation screening would
have been performed using asreview? All datasets accompanying the
systematic reviews are openly published.

\hypertarget{datasets}{%
\subsection{Datasets}\label{datasets}}

The algorithm will be tested on five systematic reviews from various
research areas. test datasets serve as systematic search results, then
perform active learning to detect inclusions.

Cohen et al.~collected systematic review datasets from the medical
sciences \autocite{Cohen2006}. All systematic reviews in this database
are on drug efficacy. The \emph{ace} dataset used in the current study
comes from a systematic review on the efficacy of Angiotensin-converting
enzyme (ACE) inhibitors.

a machine learning-based citation classification tool to reduce workload
in systematic reviews of drug class efficacy. Using a perceptron
classifier, \href{mailto:WSS@95}{\nolinkurl{WSS@95}}\% = 56.61 in
\autocite{Cohen2006}. (5x2 crossvalidation). Can we beat this? The data

The \emph{software} dataset is retrieved from \autocite{Yu2018a}, who
collected datasets on literature reviews from the software engineering
field. This dataset is on fault prediction in software engineering by
\autocite{Hall2012}.

The \emph{nudging} dataset comes from a review from the behavioural
public administration area. The SR includes studies on nudging
healthcare professionals \autocite{Nagtegaal2019}. The data was stored
on the Harvard Dataverse \autocite{Nagtegaal2019a}.

A literature review from field of psychology, \emph{ptsd}. The SR is on
studies applying latent trajectory analyses on posttraumatic stress
after exposure to trauma \autocite{vandeSchoot2017}. The corresponding
data can be found on the Open Science Framework {[}\ldots{]}.

\emph{wilson}, a dataset from the medical sciences
\autocite{Appenzeller-Herzog2020}. TA review on effectiveness treatments
of Wilson disease \autocite{Appenzeller-Herzog2019}.

All datasets started with an initial pool of thousands of papers. A
fraction of these papers where deemed relevant for the SR, with
inclusion rates around 1-2 percent with one outlier of about 5 percent
(Table 1).

The raw datafiles published with the SRs were preprocessed into a test
dataset. The test datasets contain title information on all citations
obtained in the search strategy. The instances in the data consist of a
title, an abstract and were labeled to indicate which citations were
included in the systematic review. Instances with missing abstracts were
removed. Duplicate instances were removed. Preprocessing scripts can be
found on the GitHub\footnote{\url{https://github.com/GerbrichFerdinands/asreview-thesis}}

Table 1: Statistics on datasets from original systematic reviews.

\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
\multicolumn{1}{c}{} & \multicolumn{3}{c}{Original study} & \multicolumn{3}{c}{Test collection} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
Dataset & Candidate studies & Final inclusions & Inclusion rate (\%) & Candidate studies & Final inclusions & Inclusion rate (\%)\\
\midrule
ace & 2544 & 41 & 1.61 & 2235 & 41 & 1.83\\
nudging & 2006 & 100 & 4.99 & 1848 & 100 & 5.41\\
ptsd & 6185 & 38 & 0.61 & 5031 & 38 & 0.76\\
software & 8911 & 104 & 1.17 & 8896 & 104 & 1.17\\
wilson & 3453 & 26 & 0.75 & 2334 & 24 & 1.03\\
\bottomrule
\end{tabu}

\hypertarget{models}{%
\subsection{Models}\label{models}}

Five different active learning models were build. Every model \(m\) was
used to perform an automated systematic review on SR dataset \(d\). The
models all apply a different classifier \(c\) with its own set of
(hyper)parameters \(h\).

\hypertarget{classifiers}{%
\paragraph{Classifiers}\label{classifiers}}

The classifier predicts the class of all papers, given the training
dataset/labeled data set \(\mathcal{L}\).

Given the labels and some features from all abstracts in the pool, the
classifier To predict whether a paper should be an exclusion or an
inclusion, different classifiers

Logistic Regression (L) -

Naive Bayes (B) - predicts the class of an instance given input features
\(\mathbf{X}\). Naive Bayes assumes all features are independent given
the class value. This is obviously not the case but still the algorithm
performs impressively \autocite{Zhang2004}. Especially at \ldots{}
tasks.

Random Forests (R) is where a large number of decision trees are fit on
bootstrapped samples of the original data. All trees cast a vote on the
class, which are aggregated into a class prediciton for each input
\(\mathbf{X}\) \autocite{Breiman2001}.

Support Vecor Machine (S) - finds a multidimensional hyperplane to
separate classes. \autocite{Tong2001}

Dense Neural Network (N) -

Besides \(c\), components of \(m\) are fea

To summarize, every model \(m\) consists of the following key
components: classifier, feature extraction strategy, query strategy,
balance strategy.

For example \(M_1\) BTMD (naive bayes, tfidf, certainty sampling, double
balance) STMD

\hypertarget{word-representation}{%
\paragraph{Word representation}\label{word-representation}}

To be able to predict whether a paper needs to be included or excluded
(e.g.~to predict class), the classifier needs some features from the
papers.\\
As features we use the title and abstract from every paper - this is
what is prescribed by prisma, (check!) The classifier cannot predict the
paper class from the raw titles and abstracts as they are. Therefore,
the content of the texts needs to be transformed into numerical
representations called feature vectors.

A classical example is a `bag of words' representation. For each each
text, the number of occurrences of each word is stored. This leads to n
features, where n is the number of distinct words in the texts
\autocite{scikit-learn}. The bag-of-words method is simplistic and will
highly value often occuring but otherwise meaningless words such as
``and''. Term-frequency Inverse Document Frequency (TF-IDF)
\autocite{Ramos2003} circumvents this problem by adjusting a term
frequency in a text with the inverse docuement frequency, the frequency
of a given word in the entire corpus.

\begin{enumerate}
\def\labelenumi{(\Alph{enumi})}
\setcounter{enumi}{19}
\item
  \begin{enumerate}
  \def\labelenumii{(\Alph{enumii})}
  \setcounter{enumii}{3}
  \tightlist
  \item
  \end{enumerate}
\end{enumerate}

\hypertarget{the-next-abstract-to-be-queried}{%
\paragraph{The next abstract to be
queried}\label{the-next-abstract-to-be-queried}}

The model has various ways of deciding which instance should be queried
next.

\hypertarget{rebalancing-the-training-set}{%
\paragraph{Rebalancing the training
set}\label{rebalancing-the-training-set}}

To account for class imbalance in the data, the model can apply several
strategies to rebalance the training set.

A reweighting strategy is applied where inclusions (the minority class)
are weighted more heavily than the exclusions.

When no balancing is applied, the training data set = labeled data set
\(\mathcal{L}\)

\hypertarget{starting-point}{%
\paragraph{Starting point}\label{starting-point}}

The initial training set consists of 5 inclusions and 5 exclusions,
randomly sampled from the dataset. We use 5 inclusions and exclusions as
we assume the researcher has some prior knowledge on this. The
researcher has some prior knowledge about the pool, some papers ought to
be included in the SR.

\begin{verbatim}
c = naivebayes
f = tfidf
q = max
n_prior_included = 5
n_prior_excluded = 5
\end{verbatim}

\hypertarget{retraining}{%
\paragraph{Retraining}\label{retraining}}

We can choose to retrain the model after labeling \(n\) instances. -
n\_instances=10 (number of papers each query)

\hypertarget{simulations}{%
\subsection{Simulations}\label{simulations}}

To evaluate performance of the five models described above, the model is
used to simulate five existing systematic reviews.

\hypertarget{optimizing-hyperparameters}{%
\subsubsection{Optimizing
hyperparameters}\label{optimizing-hyperparameters}}

For every model*data combination, 3 sets of hyperparameters are
generated to \ldots{} parallel Every model has its own hyperparameters.
For every model, the hyperparameters are optimized three times, arriving
at three versions of the model:

We now have 75 combinations. for every for every model (5), for every
dataset (5) and for every set of optimized hyperparameters (3), a
simulation study consisting trials is performed. From these \(5*5*3=75\)
simulation studies, performance of the different models is evaluated.

A simulation is of one model on one dataset. The simulation is repeated
for 10 trials \(t\).

Every simulation study consists of 10 trials, to account for the
randomness of prior inclusions and exclusions. So every trial the prior
inclusions and exclusions are randomly selected. Results are aggregated
(?)

\hypertarget{statistical-analysis}{%
\subsubsection{Statistical analysis}\label{statistical-analysis}}

\hypertarget{measures}{%
\paragraph{measures}\label{measures}}

\hypertarget{comparison-over-conditions}{%
\paragraph{comparison over
conditions}\label{comparison-over-conditions}}

\hypertarget{the-software}{%
\section{The software}\label{the-software}}

ASReview takes the following parameters/arguments:

\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X}
\toprule
  & Configurations\\
\midrule
Models & 2-Layer Neural Network, Naive Bayes, Random Forest, Support Vector Machine, Logistic Regression\\
Query Strategies & Cluster Sampling, Maximum Sampling, Cluster * Maximum  Sampling, Maximum * Uncertainty Sampling, Maximum * Random Sampling, Cluster * Uncertainty Sampling, Cluster * Random Sampling\\
Feature extraction strategies & Doc2Vec, TF-IDF, sbert, embeddingIdf\\
\bottomrule
\end{tabu}

Use these inputs to predict relevance of papers.

\hypertarget{stage-1-hyperparameter-optimization}{%
\subsubsection{Stage 1: hyperparameter
optimization}\label{stage-1-hyperparameter-optimization}}

Or, more specific:

\begin{tabular}{ll}
\toprule
Models & Feature extraction strategies\\
\midrule
dense\_nn & doc2vec\\
nb & tfidf\\
rf & tfidf\\
svm & doc2vec\\
lr & tfidf\\
\bottomrule
\end{tabular}

\hypertarget{hyperparameters}{%
\subsubsection{Hyperparameters}\label{hyperparameters}}

Every model has its own set of hyperparameters:

\hypertarget{optimization}{%
\subsubsection{Optimization}\label{optimization}}

The hyperparameters are optimized on the 5 datasets in three different
ways:

\begin{itemize}
\tightlist
\item
  1 on 1: maximum performance
\end{itemize}

\[ d = D \] - 4 on 1: cross-validation \[ d \notin D\]

\[ D = {1, 2, 3, 4}\]

\begin{itemize}
\tightlist
\item
  5 on 1: more data = more better?
\end{itemize}

\[ d \in D \]

This results \((5+5+1)*5\) sets of hyperparameters.

\hypertarget{performance-metrics}{%
\paragraph{Performance metrics}\label{performance-metrics}}

For each model, Several metrics are used to compare performance of
different models over datasets,

\begin{longtable}[]{@{}lllllll@{}}
\toprule
\begin{minipage}[b]{0.29\columnwidth}\raggedright
Dataset\strut
\end{minipage} & \begin{minipage}[b]{0.03\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright
Naive Bayes\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright
Random Forests\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright
Support Vector Machine\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright
Logistic Regression\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright
Dense Neural Network\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.29\columnwidth}\raggedright
ptsd\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
?\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.29\columnwidth}\raggedright
ace\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
?\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.29\columnwidth}\raggedright
hall\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
?\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.29\columnwidth}\raggedright
nagtegaal\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
?\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.29\columnwidth}\raggedright
\ldots.\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
?\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

The goal is twofold: we want to identify all relevant papers, as fast as
we can. Tradeoff: identifying all relevant papers and reducing workload.
A good metric to evaluate this is..

What is more important: recall or precision?

Recall more highly valued than precision.

What about class imbalance?

\hypertarget{rrf}{%
\paragraph{RRF}\label{rrf}}

Amount of relevant references found after having screened a certain
percentage of the total number of abstracts.

\hypertarget{work-saved-over-sampling-wss}{%
\paragraph{Work saved over sampling
(WSS)}\label{work-saved-over-sampling-wss}}

Indicates how much time can be saved, at a given level of recall. WSS is
in terms of the percentage of abstracts that don't have to be screened
by the researcher. Typically, WSS is measured at a recall of 0.95.
Reasonable because..

\[\texttt{WSS} = \frac{TN + FN}{N} - (1- recall) \]

\hypertarget{raoul}{%
\paragraph{Raoul}\label{raoul}}

\hypertarget{utility}{%
\paragraph{Utility?}\label{utility}}

\hypertarget{f-measure}{%
\paragraph{F-measure}\label{f-measure}}

\hypertarget{rocauc}{%
\paragraph{ROC/AUC}\label{rocauc}}

Is performance related to some characteristic (n, inclusion rate,
\ldots)

? How to compare outcomes of 3 different optimization strategies?

\newpage

\hypertarget{results-1}{%
\section{Results}\label{results-1}}

\newpage

\hypertarget{discussion-1}{%
\section{Discussion}\label{discussion-1}}

\newpage

\hypertarget{appendix-a---list-of-definitions}{%
\section{Appendix A - list of
definitions}\label{appendix-a---list-of-definitions}}

\hypertarget{feature-extraction-strategies}{%
\subsubsection{Feature Extraction
Strategies}\label{feature-extraction-strategies}}

split\_ta = overall hyperparameter

\hypertarget{tf-idf}{%
\paragraph{TF-IDF}\label{tf-idf}}

\hypertarget{hyperparameters-1}{%
\subparagraph{hyperparameters}\label{hyperparameters-1}}

\begin{verbatim}
ngram_max: int
        Can use up to ngrams up to ngram_max. For example in the case of
        ngram_max=2, monograms and bigrams could be used.
\end{verbatim}

\hypertarget{doc2vec}{%
\paragraph{Doc2Vec}\label{doc2vec}}

Predicts words from context. Aims at capturing the relations between
word (man-woman, king-queen). \autocite{Le2014}. Using a neural network.

using Continuous Bag-of-Words (CBOW), Skip-Gram model, \ldots. Word
vector \emph{W} and extra: document vector \emph{D}, trained to predict
words in the text.

From gensim \autocite{Rehurek2010}.

\begin{verbatim}
    Arguments
    ---------
    vector_size: int
        Output size of the vector.
    epochs: int
        Number of epochs to train the doc2vec model.
    min_count: int
        Minimum number of occurences for a word in the corpus for it to
        be included in the model.
    workers: int
        Number of threads to train the model with.
    window: int
        Maximum distance over which word vectors influence each other.
    dm_concat: int
        Whether to concatenate word vectors or not.
        See paper for more detail.
    dm: int
        Model to use.
        0: Use distribute bag of words (DBOW).
        1: Use distributed memory (DM).
        2: Use both of the above with half the vector size and concatenate
        them.
    dbow_words: int
        Whether to train the word vectors using the skipgram metho

\end{verbatim}

\hypertarget{sbert}{%
\paragraph{SBERT}\label{sbert}}

BERT-base model with mean-tokens pooling \autocite{Reimers2019}

\hypertarget{embeddingidf}{%
\paragraph{embeddingIdf}\label{embeddingidf}}

This model averages the weighted word vectors of all the words in the
text, in order to get a single feature vector for each text. The weights
are provided by the inverse document frequencies

\hypertarget{models-1}{%
\subsubsection{Models}\label{models-1}}

\hypertarget{naive-bayes}{%
\paragraph{Naive Bayes}\label{naive-bayes}}

Naive Bayes assumes all features are independent given the class value.
\autocite{Zhang2004}

ASReview uses the \texttt{MultinomialNB} from the scikit-learn package
\autocite{scikit-learn}, that implements the naive Bayes algorithm for
multinomially distributed data. \texttt{nb}

Hyperparameters

\begin{itemize}
\tightlist
\item
  alpha - accounts for features not present in learning samples and
  prevents zero probabilities in further computations.
\end{itemize}

\hypertarget{random-forests}{%
\paragraph{Random Forests}\label{random-forests}}

A number of decision trees are fit on bootstrapped samples of the
original data, \autocite{Breiman2001} RandomForestClassifier from
sklearn

Arguments --------- n\_estimators: int Number of estimators.
max\_features: int Number of features in the model. class\_weight: float
Class weight of the inclusions. random\_state: int, RandomState Set the
random state of the RNG. """

\hypertarget{support-vector-machine}{%
\paragraph{Support Vector Machine}\label{support-vector-machine}}

Arguments --------- gamma: str Gamma parameter of the SVM model.
class\_weight: class\_weight of the inclusions. C: C parameter of the
SVM model. kernel: SVM kernel type. random\_state: State of the RNG.

\hypertarget{logistic-regression}{%
\paragraph{Logistic Regression}\label{logistic-regression}}

\hypertarget{dense-neural-network}{%
\paragraph{\texorpdfstring{\textbf{Dense Neural
Network}}{Dense Neural Network}}\label{dense-neural-network}}

\hypertarget{query-strategies}{%
\subsubsection{Query Strategies}\label{query-strategies}}

\begin{itemize}
\tightlist
\item
  Max - Choose the most likely samples to be included according to the
  model
\item
  Uncertainty - choose the most uncertain samples according to the model
  (i.e.~closest to 0.5 probability) \autocite{Lewis1994}
\item
  Random - randomly selects abstracts with no regard to model assigned
  probabilities.
\item
  Cluster - Use clustering after feature extraction on the dataset. Then
  the highest probabilities within random clusters are sampled
\end{itemize}

The following combinations are simulated:

\begin{itemize}
\tightlist
\item
  cluster
\item
  max
\item
  cluster * random
\item
  cluster * uncertainty
\item
  max * cluster
\item
  max * random
\item
  max * uncertainty
\end{itemize}

\hypertarget{balance-strategies}{%
\subsubsection{Balance Strategies}\label{balance-strategies}}

\hypertarget{amount-of-training-data}{%
\subsubsection{amount of training data}\label{amount-of-training-data}}

\begin{itemize}
\tightlist
\item
  n\_instances = number of papers queried each query
\item
  n\_queries = number of queries
\item
  n\_prior\_included: 5
\item
  n\_prior\_excluded:
\end{itemize}

\hypertarget{combinations}{%
\section{Combinations}\label{combinations}}

This leads to 119 combinations of configurations.

\begin{itemize}
\tightlist
\item
  Naive bayes only goes with tfidf feature extraction.
\item
  For the feature extraction strategies we will focus on doc2vec and
  tfidf. (but will compute all 4)
\item
  This leads to 3 * 7 * 4 * 3 + 1 * 7 * 1 * 3 = 273 combinations.
\end{itemize}

See appendix A for a table containing all 273 combinations.

\hypertarget{cross-validation}{%
\subsection{Cross-validation}\label{cross-validation}}

Should give an accurate estimate of maximum performance / future
systematic reviews to be performed.

\hypertarget{appendix-b---combinations}{%
\section{Appendix B - combinations}\label{appendix-b---combinations}}

\begin{longtable}{lll}
\toprule
Model & Query Strategy & Feature extraction strategy\\
\midrule
\endfirsthead
\multicolumn{3}{@{}l}{\textit{(continued)}}\\
\toprule
Model & Query Strategy & Feature extraction strategy\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
dense\_nn & cluster & doc2vec\\
dense\_nn & max & doc2vec\\
dense\_nn & max * cluster & doc2vec\\
dense\_nn & max * uncertainty & doc2vec\\
dense\_nn & max * random & doc2vec\\
\addlinespace
dense\_nn & cluster * uncertainty & doc2vec\\
dense\_nn & cluster * random & doc2vec\\
dense\_nn & cluster & tfidf\\
dense\_nn & max & tfidf\\
dense\_nn & max * cluster & tfidf\\
\addlinespace
dense\_nn & max * uncertainty & tfidf\\
dense\_nn & max * random & tfidf\\
dense\_nn & cluster * uncertainty & tfidf\\
dense\_nn & cluster * random & tfidf\\
dense\_nn & cluster & sbert\\
\addlinespace
dense\_nn & max & sbert\\
dense\_nn & max * cluster & sbert\\
dense\_nn & max * uncertainty & sbert\\
dense\_nn & max * random & sbert\\
dense\_nn & cluster * uncertainty & sbert\\
\addlinespace
dense\_nn & cluster * random & sbert\\
dense\_nn & cluster & embeddingIdf\\
dense\_nn & max & embeddingIdf\\
dense\_nn & max * cluster & embeddingIdf\\
dense\_nn & max * uncertainty & embeddingIdf\\
\addlinespace
dense\_nn & max * random & embeddingIdf\\
dense\_nn & cluster * uncertainty & embeddingIdf\\
dense\_nn & cluster * random & embeddingIdf\\
nb & cluster & tfidf\\
nb & max & tfidf\\
\addlinespace
nb & max * cluster & tfidf\\
nb & max * uncertainty & tfidf\\
nb & max * random & tfidf\\
nb & cluster * uncertainty & tfidf\\
nb & cluster * random & tfidf\\
\addlinespace
rf & cluster & doc2vec\\
rf & max & doc2vec\\
rf & max * cluster & doc2vec\\
rf & max * uncertainty & doc2vec\\
rf & max * random & doc2vec\\
\addlinespace
rf & cluster * uncertainty & doc2vec\\
rf & cluster * random & doc2vec\\
rf & cluster & tfidf\\
rf & max & tfidf\\
rf & max * cluster & tfidf\\
\addlinespace
rf & max * uncertainty & tfidf\\
rf & max * random & tfidf\\
rf & cluster * uncertainty & tfidf\\
rf & cluster * random & tfidf\\
rf & cluster & sbert\\
\addlinespace
rf & max & sbert\\
rf & max * cluster & sbert\\
rf & max * uncertainty & sbert\\
rf & max * random & sbert\\
rf & cluster * uncertainty & sbert\\
\addlinespace
rf & cluster * random & sbert\\
rf & cluster & embeddingIdf\\
rf & max & embeddingIdf\\
rf & max * cluster & embeddingIdf\\
rf & max * uncertainty & embeddingIdf\\
\addlinespace
rf & max * random & embeddingIdf\\
rf & cluster * uncertainty & embeddingIdf\\
rf & cluster * random & embeddingIdf\\
svm & cluster & doc2vec\\
svm & max & doc2vec\\
\addlinespace
svm & max * cluster & doc2vec\\
svm & max * uncertainty & doc2vec\\
svm & max * random & doc2vec\\
svm & cluster * uncertainty & doc2vec\\
svm & cluster * random & doc2vec\\
\addlinespace
svm & cluster & tfidf\\
svm & max & tfidf\\
svm & max * cluster & tfidf\\
svm & max * uncertainty & tfidf\\
svm & max * random & tfidf\\
\addlinespace
svm & cluster * uncertainty & tfidf\\
svm & cluster * random & tfidf\\
svm & cluster & sbert\\
svm & max & sbert\\
svm & max * cluster & sbert\\
\addlinespace
svm & max * uncertainty & sbert\\
svm & max * random & sbert\\
svm & cluster * uncertainty & sbert\\
svm & cluster * random & sbert\\
svm & cluster & embeddingIdf\\
\addlinespace
svm & max & embeddingIdf\\
svm & max * cluster & embeddingIdf\\
svm & max * uncertainty & embeddingIdf\\
svm & max * random & embeddingIdf\\
svm & cluster * uncertainty & embeddingIdf\\
\addlinespace
svm & cluster * random & embeddingIdf\\
lr & cluster & doc2vec\\
lr & max & doc2vec\\
lr & max * cluster & doc2vec\\
lr & max * uncertainty & doc2vec\\
\addlinespace
lr & max * random & doc2vec\\
lr & cluster * uncertainty & doc2vec\\
lr & cluster * random & doc2vec\\
lr & cluster & tfidf\\
lr & max & tfidf\\
\addlinespace
lr & max * cluster & tfidf\\
lr & max * uncertainty & tfidf\\
lr & max * random & tfidf\\
lr & cluster * uncertainty & tfidf\\
lr & cluster * random & tfidf\\
\addlinespace
lr & cluster & sbert\\
lr & max & sbert\\
lr & max * cluster & sbert\\
lr & max * uncertainty & sbert\\
lr & max * random & sbert\\
\addlinespace
lr & cluster * uncertainty & sbert\\
lr & cluster * random & sbert\\
lr & cluster & embeddingIdf\\
lr & max & embeddingIdf\\
lr & max * cluster & embeddingIdf\\
\addlinespace
lr & max * uncertainty & embeddingIdf\\
lr & max * random & embeddingIdf\\
lr & cluster * uncertainty & embeddingIdf\\
lr & cluster * random & embeddingIdf\\*
\end{longtable}

\hypertarget{appendix-c---supercomputer-cartesius}{%
\section{Appendix C - supercomputer
Cartesius}\label{appendix-c---supercomputer-cartesius}}

500,000 SBU

Running on Cartesius is charged in System Billing Units (SBUs), and
charging is based on the wall clock time of a job. On fat and thin
nodes, an SBU is equal to using 1 core for 1 hour (a core hour), or 1
core for 20 minutes on a GPU node. Since compute nodes are allocated
exclusively to a single job at a time, you will be charged for all cores
on that node - even if you are using less.

\printbibliography[title=References]

\end{document}