### 2020-03-17 simulation/README.Rmd

---
title: "logbook simulations"
output:   
  md_document:
    variant: markdown_github
---

This is the log of simulations. 
There are three stages, we start with the first where we are comparing different classifiers.

```{r include = FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r, echo = FALSE}
conditiongrid <- readRDS("00_conditiongrid.RDS")

conditiongrid$condition %>%
  kable(format = "markdown") %>%
  pack_rows("Classifier", 1, 5) %>%
  pack_rows("Feature Extraction", 6, 10) %>%
  pack_rows("Balance Strategy", 11, 20)
```

### Version information
Personal device

Cartesius supercomputer

-----

# Stage 1 - Classifiers 
Comparing five different classifiers
```{r}
hyperparametersets <- readRDS("01_hyperparameter_sets.RDS")

hyperparametersets %>%
  kable(format = "markdown")
```

R - 275 secs 119% - 5 datasets 
```{r}
275 / 5 * 1000 / 60 # 916 minutes
```

## 1.1 Optimizing (hyper)parameters per model, per dataset (1-1)

### 1.1.1 Ace
```{r}
lubridate::date("2020-03-17")
```

Optimization run on own device 
```{bash, eval = FALSE}
# RCTW
mpirun -n 2 asreview hyper-active -m nb -b double -e tfidf -q max -d ace -n 10000 -r 8 --mpi 
```

Optimization run on Cartesius
```{bash, eval = FALSE}
# RCTW
./cart_hyper_run.py hyper-active -t 10 -m rf -b double -e tfidf -q max -d ace

# SCTW

# LCTW

# NCTW
```

### 1.1.2 Nudging
### 1.1.3 PTSD
### 1.1.4 Software
### 1.1.5 Wilson

# Stage 2 - Feature exraction 
Comparing tfidf with doc2vec

# Stage 3 - Balance strategy
Redo all the analyses with a different balance strategy (aggressive undersampling?)
Probably won't have time for this.