### 2020-03-17 simulation/README.Rmd
udpate simulation

---
title: "logbook simulations"
output:   
  md_document:
    variant: markdown_github
---

This is the log of simulations. 
There are three stages, we start with the first where we are comparing different classifiers.

```{r include = FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)

knitr::opts_chunk$set(echo = FALSE)
```

```{r}
conditiongrid <- readRDS("00_conditiongrid.RDS")

conditiongrid %>%
  kable(format = "markdown")
```

### Version information
Personal device

Cartesius supercomputer

asreview

-----

# Stage 1 - Classifiers 
Comparing five different classifiers
```{r}
hyperparametersets <- readRDS("01_hyperparameter_sets.RDS")
```

R - 275 secs 119% - 5 datasets 
```{r}
275 / 5 * 1000 / 60 # 916 minutes
```

## 1.1 Optimizing (hyper)parameters per model, per dataset (1-1)

### 1.1.1 Ace

```{r}
lubridate::date("2020-03-17")
```

Optimization run on own device 
```{bash, eval = FALSE}
# RCTW
mpirun -n 2 asreview hyper-active -m nb -b double -e tfidf -q max -d ace -n 10000 -r 8 --mpi 
```

Optimization run on Cartesius
```{bash, eval = FALSE}
# RCTW
./cart_hyper_run.py hyper-active -t 5 -m rf -b double -e tfidf -q max -d ace -r 23

# SCTW

# LCTW

# NCTW
```

### 1.1.2 Nudging
### 1.1.3 PTSD
### 1.1.4 Software
### 1.1.5 Wilson

```{r}
# to-do grid 
todo <- hyperparametersets

# set emojis for grid 
c <- emo::ji("check")

todo$ace[1] <- c
```

```{r}
# save the to-do grid in markdown style
todo <- todo %>%
  kable(format = "markdown")
saveRDS(todo, file = "02_todo_hp.RDS")

todo 
```

# Stage 2 - Feature exraction 
Comparing tfidf with doc2vec

# Stage 3 - Balance strategy
Redo all the analyses with a different balance strategy (aggressive undersampling?)
Probably won't have time for this.