### 2020-03-17 datasets/data_preprocessing.nb.Rmd
- add nagtegaal dataset

---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
This notebook processes all five systematic review datasets into datasets that can be used for testing asreview.

```{r, results = "hide"}
library(tidyverse)
library(knitr)
library(kableExtra)
```

### Set-up
```{r}
# create storage for statistics
template <- data.frame(search = rep(0,4), # papers obtained in systematic search
                       ftext = rep(0, 4), # number of abstract inclusions
                       incl = rep(0,4), # papers included in final systematic review

                       row.names = c("paper", # numbers reported in the publication
                                     "raw", # numbers published on raw dataset (by author,  mostly on OSF)
                                     "asreview", # numbers on asreview repository 
                                     "test")) # numbers in test dataset

# statistics on duplicates and dropping in ASReview dataset
drops <- data.frame(n = rep(0,5), # candidates
                    na = rep(0,5),# number of missing abstracts
                    narate = rep(0,5), # percenta of missing abstracts
                    dup = rep(0,5),  # number of duplicate abstracts 
                    duprate = rep(0,5), # percentage of missing abstracts
                    row.names = c("ace", "nudging", "ptsd", "software", "wilson"))

# some functions to extract statistics
data_descr <- function(data){
  return(c(nrow(data), # all papers obtained in the systematic search
           ifelse(sum(data$inclusion_code > 0), sum(data$inclusion_code > 0), NA), # number of abstract inclusions
           sum(data$included))) # papers included in systematic review
}

nostudies <- function(all, set, stage){ # number of studies
  sapply(all, function(x) x[set, stage])
}

inclrate <- function(all, set){ # inclusion rate
  sapply(all, function(x) round(x[set,"incl"]/x[set,"search"]*100,2))
}
```

## Ace dataset 
```{r}
ace <- template
# paper ------------------------------------------------------------------------
ace["paper", ] <- c(2544, NA, round(2544*0.0160))

# raw --------------------------------------------------------------------------
# original paper not to linked directly. cannot be retrieved? 
# checked dataset online by cohen et al. 
#https://dmice.ohsu.edu/cohenaa/systematic-drug-class-review-data.html
ace["raw", ] <- c(2544, NA, 41)

# asreview ---------------------------------------------------------------------
ace_asr <- read.csv("https://raw.github.com/asreview/systematic-review-datasets/master/datasets/Cohen_EBM/output/ACEInhibitors.csv",
                     header=T)

ace["asreview", ] <- c(nrow(ace_asr),
                       NA,
                       sum(ace_asr$label_included))
```

### Create test dataset
```{r}
# replace empty abstracts by NA values. 
ace_asr[ace_asr$abstract == "", "abstract"] <- NA

# drop missings and duplicates. 
ace_test <- ace_asr %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  filter(abstract != "") %>%
  distinct(abstract, .keep_all = TRUE) # remove entries with duplicate abstracts

# select columns
ace_test <- ace_test %>%
       select(authors, title, abstract, label_included) # select relevant columns

# data on 
ace["test", ] <- c(nrow(ace_test),
                   NA,
                   sum(ace_test$label_included))
```

### Some statistics 
```{r}
# missingness and duplicates in raw data on ASReview
drops["ace",] <- ace_asr %>%
          summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    na_rate = sum(is.na(abstract))/length(abstract),
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    dup_rate = sum(duplicated(abstract, incomparables = NA))/length(abstract))
drops["ace",]
ace
```

## nudging dataset 
```{r}
# by Rosanna Nagtegaal
nudging <- template
nudging["paper", ] <- c(2006, 377, 100) 
```

```{r eval = FALSE}
# data is not public yet

# raw --------------------------------------------------------------------------
n_raw <- read.csv("../../scripting/nagtegaal_preprocessing/data/output/nagtegaal.csv")
# asreview ---------------------------------------------------------------------

# remove missing abstracts 
nudging_test <- n_raw %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  filter(abstract != "", abstract != "NA") %>%
  distinct(abstract, .keep_all = TRUE) # remove entries with duplicate abstracts
# missing statistics

```

### Create test dataset

```{r}
nudging["test", ] <- data_descr(nudging_test)
```

## PTSD dataset
```{r}
ptsd <- template 
ptsd["paper", ] <- c(6185, 363, 38) # from flowchart

# raw --------------------------------------------------------------------------
ptsd["raw", ]

# asreview ---------------------------------------------------------------------
ptsd_asr <- read.csv("https://raw.github.com/asreview/systematic-review-datasets/master/datasets/Van_de_Schoot_PTSD/output/PTSD_VandeSchoot_18.csv", 
                     header=T)

ptsd["asreview", ] <- data_descr(ptsd_asr)

```

### Create test dataset
Now, remove duplicate entries and empty abstracts to arrive at test set. 
```{r}
ptsd_test <- ptsd_asr %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  filter(abstract != "") %>%
  distinct(abstract, .keep_all = TRUE)# remove entries with duplicate abstracts
```

There are 38 inclusions in this dataset, should be 34. 
```{r}
# get rid of 4 extra inclusions (see log book)
excl <- 
  ptsd_test %>%
  filter(included == 1, 
         str_detect(authors, "Sterling, M., Hendrikz, J., Kenardy, J.") |
         str_detect(authors, "Hou") |
         str_detect(authors, "Mason") |
         str_detect(authors, "Pérez")) %>%
  select(id)

indices <- ptsd_test$id %in% excl

# get rid of 4 papers
ptsd_test[indices, "included"] <- 0
```

```{r}
# finalize dataset  
ptsd_test <- ptsd_test %>%
  select(authors, title, abstract, included, inclusion_code) # select relevant columns
```

Derive statistics in final test set.
```{r}
ptsd["test", ] <- data_descr(ptsd_test)

# statistics on duplicates from asr data to test
drops["ptsd", ] <- ptsd_asr %>%
  summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    narate = sum(is.na(abstract))/length(abstract),
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract))

ptsd
```

## software dataset
```{r}
hall <- template

# paper ------------------------------------------------------------------------
hall["paper", ] <- c(8911, NA, 104)

# raw --------------------------------------------------------------------------
h_raw <- read.csv("https://zenodo.org/record/1162952/files/Hall.csv",
                  header=T)

hall["raw", ] <- c(length(h_raw$Document.Title),
                   NA,
                   sum(h_raw$label == "yes"))

# asreview ---------------------------------------------------------------------
hall_asr <- read.csv("https://raw.github.com/asreview/systematic-review-datasets/master/datasets/Four%20Software%20Engineer%20Data%20Sets/Software%20Engineering%201%20Hall.csv",
                     header=T)

hall["asreview", ] <- data_descr(hall_asr)
```

### Creating test dataset
```{r}
hall_test <- hall_asr %>%
   drop_na(abstract) %>% # drop entries with missing abstracts
   filter(abstract != "") %>%
   distinct(abstract, .keep_all = TRUE)# remove entries with duplicate abstracts

hall["test", ] <- data_descr(hall_test)
```

### Statistics 
```{r}
drops["software", ] <- 
  hall_asr %>%
   summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    narate = sum(is.na(abstract))/length(abstract),
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract))

hall
```

## Wilson dataset

```{r}
wilson <- template
# paper ------------------------------------------------------------------------
wilson["paper", ] <- c(3453, 174, 26)

# raw --------------------------------------------------------------------------
# w_ftext <- read.delim("../datasets/raw/DOKU_All FT-Screening_20200116_cap.txt")
# w_incl <- read.csv("../datasets/raw/DOKU_All Included_20200116_cap.csv")
# w_all <- read.csv("../datasets/raw/DOKU_All TiAb-Screening_20200116_cap.csv")
# wilson["raw", ] <- c(nrow(w_all), nrow(w_ftext), nrow(w_incl))
wilson["raw", ] <- c(3453, 174, 26) # looked it up

# asreview ---------------------------------------------------------------------
w_asr <- read.csv("https://raw.github.com/asreview/systematic-review-datasets/master/datasets/Appenzeller-Herzog_Wilson/output/output_csv_wilson.csv", header=T)

# add inclusion_code label 
w_asr$inclusion_code <- w_asr$label_abstract_screening + w_asr$label_included
w_asr$included <- w_asr$label_included
wilson["asreview", ] <- data_descr(w_asr)
```

```{r}
# create test dataset
w_test <- w_asr %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  filter(abstract != "",
         abstract!= " ") %>%
  distinct(abstract, .keep_all = TRUE) # remove entries with duplicate abstracts

# select relevant columns
w_test <- w_test %>%
  select(authors, title, abstract, included, inclusion_code)

wilson["test",] <- data_descr(w_test)
```

Statistics
```{r}
drops["wilson", ] <- w_asr %>%
   summarise(n = length(abstract),
                    na = sum(is.na(abstract)), # todo correct this, "NA", or ""
                    narate = sum(is.na(abstract))/length(abstract),
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract))

wilson
```

# Save descriptive statistics on all datasets

```{r eval = TRUE}
# put everything together in list. 
all <- list(ace = ace,
            nudging = nudging,
            ptsd = ptsd,
            software = hall,
            wilson = wilson)

# save datafile, to serve as data for descriptive table in manuscript
save(all, file = "../manuscript/drafts/data/datasets_table.Rdata")
```

# Statistics
All candidate papers, paper selected for full text screening, papers included in final review and inclusion rate, over various versions of the dataset.  
```{r, echo = FALSE}
tibble(Dataset = names(all), 
       # paper 
       `candidates_paper` = nostudies(all, "paper", "search"), 
       `fulltext_paper` = nostudies(all, "paper", "ftext"),
       `incl_paper` = nostudies(all, "paper", "incl"),
       `inclrate_paper` = inclrate(all, "paper"),

       # raw data 
       `candidates_raw` = nostudies(all, "raw", "search"), 
       `fulltext_raw` = nostudies(all, "raw", "ftext"),
       `incl_raw` = nostudies(all, "raw", "incl"),
       `inclrate_raw` = inclrate(all, "raw"),

       # asreview repository 
       `candidates_asr` = nostudies(all, "asreview", "search"), 
       `fulltext_asr` = nostudies(all, "asreview", "ftext"),
       `incl_asr` = nostudies(all, "asreview", "incl"),
       `inclrate_asr` = inclrate(all, "asreview"),

       # test set 
       `candidates_test` =  nostudies(all, "test", "search"), 
       `fulltext_test` =  nostudies(all, "test", "ftext"), 
       `incl_test` =  nostudies(all, "test", "incl"), 
       `inclrate_test` = inclrate(all, "test")) %>% 
  t() %>% 
  kable("html") %>%
  kable_styling(full_width = TRUE) %>%
  #scroll_box(width = "100%", height = "200px") %>%
  pack_rows("Paper", 2, 5) %>%
  pack_rows("Raw data", 6, 9) %>%
  pack_rows("ASReview data", 10, 13) %>%
  pack_rows("Test data set", 14, 17)  
```

Descriptives on missingness and duplicate abstracts in the ASReview test data set. 
```{r, echo = FALSE}
drops %>%
  kable("html",
        col.names = c("n", "NA", "NA rate (%)", "duplicates", "duplicate rate (%)"),
        digits = 2) %>%
  kable_styling(full_width = TRUE)
```

# Write all test dataset files 
```{r}
# ace
write.csv(ace_test, "test_datasets/ace.csv")

# nudging 
write.csv(nudging_test, "test_datasets/nudging.csv")

# ptsd
write.csv(ptsd_test, "test_datasets/ptsd.csv")

# hall
write.csv(hall_test, "test_datasets/software.csv")
# wilson
write.csv(w_test, "test_datasets/wilson.csv")

```