### 2020-04-01 simulation_study/hpoptimization/config/configfiles.Rmd
test ini file

---
title: "create_config_files"
author: "Gerbrich Ferdinands"
date: "4/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
hpsets <- readRDS("../R/01_hyperparameter_sets.RDS")

```

## R Markdown

```{bash, eval = FALSE}
# get config files from surfsara to hard drive 
scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/config /Volumes/Gerbrich/asreview/config/optimization_output
```

Manually update config files to simulation 
Push finished ones to GH repository
Fetch github repo on surfsara

make link in surfsara 
ln -s /home/gerbrich/asreview/manuscript/asreview-thesis/simulation_study/simulation/config simconfig

```{r}
scp -r 
gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/siminput
```