### 2020-04-06 meetings/11_notes_7apr.Rmd
update hyperopt, update meeting notes

---
title: "11_notes_7apr"
output: github_document
---

# Notes - core team meeting - core team meeting

Questions:

How to make simulations faster?

- plot making in lc

- logistic regression 
  - [use saga as solver instead of liblinear?](https://stackoverflow.com/questions/38640109/logistic-regression-python-solvers-defintions/52388406#52388406)
  - l2 regularization --> why? why nog l1? 
  - [](https://towardsdatascience.com/dont-sweat-the-solver-stuff-aea7cddc3451)

# recap/decisions last week

# Meeting

# To do