### 2020-04-09 simulation_study/hpoptimization/config/README.Rmd
add hyperparaemter optimization

---
title: "Results hyperparameter optimization"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(tidyverse)
library(ggplot2)
library(knitr)

php <- function(data){
  data %>%
    ggplot(aes(x = value, y = "", color = group)) +
    geom_jitter(height = 0.2) +
    facet_wrap(vars(param), nrow = 8, scales = "free_x") +
    geom_vline(aes(xintercept = lb), color = "grey") +
    geom_vline(aes(xintercept = ub), color = "grey") 
}

```

```{r, figure.heigth = 12}
hpsets <- readRDS("hpresults.RDS")
```

```{r}
php(hpsets$lr_tfidf) + 
  labs(title = "Hyperparameters Logistic Regression + TF-IDF")
```

```{r, eval = FALSE}
php(hpsets$lr_doc2vec) + 
  labs(title = "Hyperparameters Logistic Regression + Doc2Vec")
```

```{r}
php(hpsets$rf_tfidf) + 
  labs(title = "Hyperparameters Random Forest + TF-IDF")
```

statistics for random forest + tf-idf
```{r}

hpsets$rf_tfidf %>%
   group_by(param, group) %>%
    summarise(variance = var(value), 
              sd = sd(value),
              min = min(value), 
              mean = mean(value),
              median = median(value),
              max = max(value)) %>%
  kable("markdown")

```

```{r}
php(hpsets$rf_d2v) + 
  labs(title = "Hyperparameters Random Forest + Doc2Vec")
```