### 2020-04-10 simulation_study/simulation/README.Rmd
Fix bug C-parameter SCTD config files

---
title: "Simulations log"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

createLines <- function(cond, d, hp){
  comm <- "./cart_simulate.py simdata/%s.csv --config_file simconfig/%s/%s/%s-%s.ini --state_file simoutput/%s/%s/%s/results.h5 -r 15"
  files <- data.frame(c = condition, 
                    f = c("nb_max_double_tfidf",
                    "rf_max_double_tfidf",
                    "svm_max_double_tfidf",
                    "logistic_max_double_tfidf",
                    "nn-2-layer_max_double_tfidf",
                    "nb_max_double_doc2vec",
                    "rf_max_double_doc2vec",
                    "svm_max_double_doc2vec",
                    "logistic_max_double_doc2vec",
                    "nn-2-layer_max_double_doc2vec"))

  # create condition grid
  cons <- expand.grid(d = d, hp = hp, c = cond, stringsAsFactors = FALSE)
  cons <- left_join(cons, files)

  # write 
  script <- sprintf(comm, 
        # input data
        cons$d, 
        # config file details
        cons$hp, cons$c, cons$f, cons$d,
        # output file details
        cons$hp, cons$c, cons$d)
return(script)
}
transfer <- function(hp, c){
    cons <- expand.grid(hp = hp, c = c, stringsAsFactors = FALSE)
    txt <- "scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/simoutput/%s/%s /Volumes/Gerbrich/asreview/simoutput/%s/%s"

    sprintf(txt, hp, c, hp, c)
}

```

```{r, include=FALSE}
sims <- readRDS("../R/01_hyperparameter_sets.RDS")
d <- colnames(sims)
condition <- rownames(sims)
classifier <- c("nb", "rf", "svm", "logistic", "nn-2-layer")
names(classifier) <- c("B", "R", "S", "L", "N")
```

# create config files 
#### 01.04.2020
```{bash, eval = FALSE}
# get config files from surfsara to hard drive 
scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/config /Volumes/Gerbrich/asreview/config/optimization_output
```

Manually update config files to simulation 
Push finished ones to GH repository
Fetch github repo on surfsara

```{r}
sims[condition[c(1,2,4)], d, "one"] <- 1 # simulation configs are up
```

# Run simulations

## hp = one

#### BCTD
```{r}
bctd <- createLines(condition[1], d=d, hp = "one")
write.table(bctd, file = "scripts/one/bctd.txt", sep = "\t", quote = FALSE,
            row.names = FALSE)
names(bctd) <- c("7883184", "7883202", "7883207", "7883209", "7883213", "7883214")
```

#### RCTD, LCTD
```{r}
rctd_lctd <- createLines(condition[c(2,4)], d=d, hp = "one")
write.table(rctd_lctd, file = "scripts/one/rctd_lctd.txt", sep = "\t", quote = FALSE,
            row.names = FALSE)
#one_rctd_virus etc 
names(rctd_lctd) <- c("7949940", "7949952", "7949953", "7949954", "7949956", "7949959",
                     "7883228", "7883229", "7883231", "7883233", "7883234", "7883236")
```

```{r}
sims[condition[c(1,2,4)], d, "one"] <- 2 # simulations are running 
```

# transfer simulation files to harddrive

```{bash, eval = FALSE}
scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/config /Volumes/Gerbrich/asreview/config/optimization_output

./cart_simulate.py simdata/wilson.csv --config_file simconfig/one/LCTD/logistic_max_double_tfidf-wilson.ini --state_file simoutput/one/LCTD/wilson/results.h5 -r 15
```

```{r}
transfer("one", condition[c(1,2,4)])
```

```{r}
# output is on harddrive
sims[condition[1], , "one"] <- 3
```

#### 02.04.2020
get config files 'all' BCTD, LCTD.

scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/config /Volumes/Gerbrich/asreview/config/optimization_output

```{r}
# output is on harddrive
sims[condition[2], , "one"] <- 3
```

#### 03.04.2020
SCDD model

```{r}
# get config files
# manually transform
# push to github 
sims[condition[8], d, "one"] <- 1 # simulation configs are up

# run simulations 
scdd <- createLines(condition[8], d, "one")
write.table(scdd, file = "scripts/one/scdd.txt", sep = "\t", quote = FALSE,
            row.names = FALSE)
names(scdd) <- c("7893237", "7893241", "7893243", "7893244", "7893246", "7893247")
sims[condition[8], d, "one"] <- 2 # simulations are running

```

```{r}
sctd <- createLines(condition[3], d, "one")
write.table(sctd, file = "scripts/one/sctd.txt", sep = "\t", quote = FALSE,
            row.names = FALSE)
names(sctd) <- c("7945897", "7945899", "7945902", "7945903", "7945910", "7945912")
sims[condition[3], d, "one"] <- 2 # simulations are running

rcdd <- createLines(condition[7], d, "one")
write.table(rcdd, file = "scripts/one/rcdd.txt", sep = "\t", quote = FALSE,
            row.names = FALSE)
names(rctd) <- c("")
sims[condition[7], d, "one"] <- 2 # simulations are running

lcdd <- createLines(condition[8], d, "one")

```

10.04.2020

```{r}
# transfer to harddrive 
transfer("one", condition[8])
#sims[condition[8], , "one"] <- 3 # output is on harddrive
transfer("one", condition[4])

# simulations both failed? 
```

### all 
RCTD all
BCTD all 
LCTD all

```{r}
# manually transform
# push to github 
sims[condition[4], , "all"] <- 1 # configs are on github

```

```{r}
# run simulations 
lctd_n <- createLines(condition[4], d, "all")
write.table(lctd_n, file = "scripts/n/lctd_n.txt", sep = "\t", quote = FALSE,
            row.names = FALSE)
names(lctd_n) <- c()
#sims[condition[4], d, "all"] <- 2 # simulations are running

# transfer to harddrive 
transfer("n", condition[4])
#sims[condition[8], , "one"] <- 3 # output is on harddrive
```

---

```{r template, eval = FALSE}
# get config files
# manually transform
# push to github 
sims[condition[1], d, "one"] <- 1 # simulation configs are up
# run simulations 
createLines()
sims[condition[1], d, "one"] <- 2 # simulations are running 
# transfer to harddrive 
transfer()
sims[condition[1], , "one"] <- 3 # output is on harddrive
```