### 2020-04-13 simulation_study/simulation/README.Rmd

---
title: "Simulations log"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

createLines <- function(cond, d, hp){
  comm <- "./cart_simulate.py simdata/%s.csv --config_file simconfig/%s/%s/%s-%s.ini --state_file simoutput/%s/%s/%s/results.h5 -r 15"
  files <- data.frame(c = condition, 
                    f = c("nb_max_double_tfidf",
                    "rf_max_double_tfidf",
                    "svm_max_double_tfidf",
                    "logistic_max_double_tfidf",
                    "nn-2-layer_max_double_tfidf",
                    "nb_max_double_doc2vec",
                    "rf_max_double_doc2vec",
                    "svm_max_double_doc2vec",
                    "logistic_max_double_doc2vec",
                    "nn-2-layer_max_double_doc2vec"))

  # create condition grid
  cons <- expand.grid(d = d, hp = hp, c = cond, stringsAsFactors = FALSE)
  cons <- left_join(cons, files)

  # write 
  script <- sprintf(comm, 
        # input data
        cons$d, 
        # config file details
        cons$hp, cons$c, cons$f, cons$d,
        # output file details
        cons$hp, cons$c, cons$d)
return(script)
}
transfer <- function(hp, c){
    cons <- expand.grid(hp = hp, c = c, stringsAsFactors = FALSE)
    txt <- "scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/simoutput/%s/%s /Volumes/Gerbrich/asreview/simoutput/%s/%s"

    sprintf(txt, hp, c, hp, c)
}

```

```{r, include=FALSE}
sims <- readRDS("../R/01_hyperparameter_sets.RDS")
d <- colnames(sims)
condition <- rownames(sims)
classifier <- c("nb", "rf", "svm", "logistic", "nn-2-layer")
names(classifier) <- c("B", "R", "S", "L", "N")
```

# create config files 
#### 01.04.2020
```{bash, eval = FALSE}
# get config files from surfsara to hard drive 
scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/config /Volumes/Gerbrich/asreview/config/optimization_output
```

Manually update config files to simulation 
Push finished ones to GH repository
Fetch github repo on surfsara

```{r}
sims[condition[c(1,2,4)], d, "one"] <- 1 # simulation configs are up
```

# Run simulations

## hp = one

#### BCTD
```{r}
bctd <- createLines(condition[1], d=d, hp = "one")
write.table(bctd, file = "scripts/one/bctd.txt", sep = "\t", quote = FALSE,
            row.names = FALSE)
names(bctd) <- c("7883184", "7883202", "7883207", "7883209", "7883213", "7883214")
```

#### RCTD, LCTD
```{r}
rctd_lctd <- createLines(condition[c(2,4)], d=d, hp = "one")
write.table(rctd_lctd, file = "scripts/one/rctd_lctd.txt", sep = "\t", quote = FALSE,
            row.names = FALSE)
#one_rctd_virus etc 
names(rctd_lctd) <- c(rep("0", 6), "7949940", "7949952", "7949953", "7949954", "7949956", "7949959")
```

```{r}
sims[condition[c(1,2,4)], d, "one"] <- 2 # simulations are running 
```

# transfer simulation files to harddrive

```{bash, eval = FALSE}
scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/config /Volumes/Gerbrich/asreview/config/optimization_output

./cart_simulate.py simdata/wilson.csv --config_file simconfig/one/LCTD/logistic_max_double_tfidf-wilson.ini --state_file simoutput/one/LCTD/wilson/results.h5 -r 15
```

```{r}
transfer("one", condition[c(1,2,4)])
```

```{r}
# output is on harddrive
sims[condition[1], , "one"] <- 3
```

#### 02.04.2020
get config files 'all' BCTD, LCTD.

scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/config /Volumes/Gerbrich/asreview/config/optimization_output

```{r}
# output is on harddrive
sims[condition[2], , "one"] <- 3
```

#### 03.04.2020
SCDD model

```{r}
# get config files
# manually transform
# push to github 
sims[condition[8], d, "one"] <- 1 # simulation configs are up

# run simulations 
scdd <- createLines(condition[8], d, "one")
write.table(scdd, file = "scripts/one/scdd.txt", sep = "\t", quote = FALSE,
           row.names = FALSE)
 names(scdd) <- c("7950186", "7950187", "7950188", "7950189", "7950190", "7950191")
 sims[condition[8], d, "one"] <- 2 # simulations are running 
```

```{r}
 sctd <- createLines(condition[3], d, "one")
 write.table(sctd, file = "scripts/one/sctd.txt", sep = "\t", quote = FALSE,
            row.names = FALSE)
names(sctd) <- c("7950483", "7950484", "7950485", "7950487", "7950489", "7950490")
sims[condition[3], d, "one"] <- 2 # simulations are running
```

```{r}
rcdd <- createLines(condition[7], d, "one")
write.table(rcdd, file = "scripts/one/rcdd.txt", sep = "\t", quote = FALSE,
            row.names = FALSE)
names(rcdd) <- c("7950491", "7950493", "7950495", "7950496", "7950497", "7950499")
sims[condition[7], d, "one"] <- 2 # simulations are running
```

```{r}
lcdd <- createLines(condition[9], d, "one")
write.table(lcdd, file = "scripts/one/lcdd.txt", sep = "\t", quote = FALSE, row.names = FALSE)
names(lcdd) <- c("7965136", "7965141", "7965143", "7965145", "7965146", "7965147")
sims[condition[9], d, "one"] <- 2 # simulations are running
```

10.04.2020

```{r}
# LCTD and SCDD both failedsimulations both failed? 
# ran again 
# transfer to harddrive
transfer("one", condition[c(4,7:9)])
# sctd is not finished yet! 
sims[condition[4],,"one"] <- 3 # output is on harddrive

# sims[condition[7],,"one"] <- 3 # output is on harddrive 
# sims[condition[8],,"one"] <- 3 # output is on harddrive 
# too little space for 9, so run again

sims[condition[9],,"one"] <- 3 # output is on harddrive 
# but ace wasn't finished yet!

```

Repeat all failed simulations

```{r}
# run again 
# transfer to harddrive 
transfer("one", condition[8])

#sims[condition[8], , "one"] <- 3 # output is on harddrive
```

### all 
```{r}
# manually transform
# push to github 
sims[condition[4], , "all"] <- 1 # configs are on github

```

```{r}
# run simulations 
# lctd_n <- createLines(condition[4], d, "all")
# write.table(lctd_n, file = "scripts/n/lctd_n.txt", sep = "\t", quote = FALSE,
#             row.names = FALSE)
# names(lctd_n) <- c()
# #sims[condition[4], d, "all"] <- 2 # simulations are running 
# 
# # transfer to harddrive 
# transfer("n", condition[4])
#sims[condition[8], , "one"] <- 3 # output is on harddrive
```

---

```{r template, eval = FALSE}
# get config files
# manually transform
# push to github 
sims[condition[1], d, "one"] <- 1 # simulation configs are up
# run simulations 
createLines()
sims[condition[1], d, "one"] <- 2 # simulations are running 
# transfer to harddrive 
transfer()
sims[condition[1], , "one"] <- 3 # output is on harddrive
```