### 2020-04-24 simulation_study/hpoptimization/README.Rmd

---
title: "Hyperparamter Optimization"
output:   
  md_document:
    variant: markdown_github
---
# Optimizing Hyperparameters
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r packages}
library(tidyverse)
library(ggplot2)
library(knitr)
library(directlabels)
```

To every model belongs a unique set of hyperparameters. To maximize model performance, we need to find optimal values for the hyperparameters. For every model the optimal hyperparameter values are determined by optimizing on data $d$. The hyperparameters are optimized by running several hundreds of optimization trials, in which hyperparameter values are sampled from their possible parameter space.

Maximum model performance is defined as the average time it takes to find an inclusion in the data, or more specific: the loss function minimizes the average number of papers needed to screen to find an inclusion (eg the area above the curve in the inclusion plot).

The optimization data d consists of (a subset from) the six systematic review datasets D mentioned above. Three different approaches in composing d are explored:

- __one__, where hyperparameters are optimized on only one of the six datasets. Such hyperparameters are expected to lead to maximum performance in the same dataset d.
- __n__, where hyperparameters are optimized on all six data sets. This optimization approach intends to serve in producing the most optimal hyperparameters overall. 
- __n-1__, where hyperparameters are optimized on all six datasets but one. Serving as a sensitivity analysis for the former condition, eg how sensitive are the hyperparamters to different data sets.

This results in 6+6+1=13 sets of hyperparameters for every model. Optimization trials were visually inspected to check if an optimum (minimal loss) has been reached. More trials were run if the loss still seemed to go down at a quick pace. The hyperparameter sets that were found to lead to a minimum loss value were visually inspected.

```{r functions}
php <- function(data){
  data %>%
    filter(param %in% numericalvars) %>%
    mutate(value = as.numeric(value)) %>% 
    ggplot(aes(x = value, y = "", color = group)) +
    geom_jitter(height = 0.2) +
    facet_wrap(vars(param), ncol = 3, scales = "free_x") +
    geom_vline(aes(xintercept = lb), color = "grey") +
    geom_vline(aes(xintercept = ub), color = "grey") +
    theme_minimal()
}

php_c <- function(data){
  data %>%
    filter(param %in% categoricalvars) %>%
    mutate(value = as.factor(value)) %>% 
    ggplot(aes(x = value, color = group, fill = group, group = group)) +
    geom_bar(position="stack") +
    facet_wrap(vars(param), nrow = 2, scales = "free_x") +
    labs(title = "(optimal) categorical hyperparameters", y = "", x = "Parameter values")+
    theme_minimal()
}
```

```{r}
hpsets <- readRDS("config/hpresults.RDS")
numericalvars <- c("balance_param.a", "balance_param.alpha", "balance_param.b", "model_param.alpha", "model_param.c", "model_param.class_weight", "feature_param.vector_size", "feature_param.epochs", "model_param.n_estimators")
categoricalvars <-c("feature_param.ngram_max", "feature_param.split_ta", "feature_param.use_keywords", "model_param.gamma", "feature_param.dbow_words", "feature_param.dm", "feature_param.dm_concat", "feature_param.min_count", "feature_param.window", "model_param.max_features") 
```

```{r loss_function}
# loss i where x is index, a.k.a. loss plot sorted from high to low loss.
loss_i <- function(trials){
  trials %>%
  ggplot(aes(x = index, y = loss)) +
  geom_line(aes(group = datasets, colour = group), lineend = "round") + 
  geom_dl(aes(label = datasets, colour = group), method = list(dl.trans(x = x + 0.2), dl.combine("last.points"), cex = 0.8)) + 
  labs(title = "Loss plot", x = "Trials sorted from high to low loss") +
    theme_minimal()
}

# loss t where x is trials, loss plot is sorted over time
loss_t <- function(trials){
  trials %>%
    ggplot(aes(x = trial, y = loss)) +
    geom_line(aes(group = datasets, colour = group), lineend = "round") + 
    geom_dl(aes(label = datasets, colour = group), method = list(dl.trans(x = x + 0.2), dl.combine("last.points"), cex = 0.8)) +
    labs(title = "Loss plot", x = "Over trials") +
    theme_minimal()
}
```

## Results
Loss plots: Two times the same data, first plot is over trials, second sorted from highest to lowest loss.

Hyperparameters on a continuous scale are plotted as dots in a scatterplot. 
Discrete hyperparameters are displayed in a barplot

Conclusions: 
- loss is going down over trials 
- some datasets go down faster
- some datasets go down more

# Naive Bayes + TF-IDF

## Loss plots

```{r}
nbtfidftrials <- readRDS("trials/nb_tfidf-trials.RDS")

nbtfidftrials <- nbtfidftrials %>%
  group_by(datasets) %>%
  mutate(index = rev(index))

# loss plot by index 
nbtfidftrials %>% 
  loss_i()

# loss plot sorted by trial numbers 
nbtfidftrials %>%
  loss_t()
```

## Hyperparameter values
```{r NB_TFIDF}
php(hpsets$nb_tfidf) + 
  labs(title = "Hyperparameters Naive Bayes + TF-IDF")

# plot feature_param.ngram_max
hpsets$nb_tfidf %>%
  php_c() 
```

Inspect outliers: 
```{r}
# inspect outliers
hpsets$nb_tfidf %>% filter(param %in% c("balance_param.a", "model_param.alpha")) %>% group_by(param) %>% filter(value > 1000) %>% select(-ub, -lb) %>%
  kable(format = "markdown")
```

# SVM + TF-IDF

```{r SVM_TFIDF}
svmtfidftrials <- readRDS("trials/svm_tfidf-trials.RDS")

svmtfidftrials <- svmtfidftrials %>%
  group_by(datasets) %>%
  mutate(index = rev(index))

# loss plot by index 
svmtfidftrials %>% 
  loss_i()

# loss plot sorted by trial numbers 
svmtfidftrials %>%
  loss_t()

# parameter values numerical 
hpsets$svm_tfidf %>%
  php() +
  labs(title = "Hyperparameters SVM + TF-IDF")

hpsets$svm_tfidf %>%
  php_c() 
```
# SVM + Doc2Vec

```{r SVM_D2V}
svmd2vtrials <- readRDS("trials/svm_d2v-trials.RDS")

svmd2vtrials <- svmd2vtrials %>%
  group_by(datasets) %>%
  mutate(index = rev(index))

# loss plot by index 
svmd2vtrials %>% 
  loss_i()

# loss plot sorted by trial numbers 
svmd2vtrials %>%
  loss_t()

# parameter values numerical 
hpsets$svm_tfidf %>%
  php() +
  labs(title = "Hyperparameters SVM + TF-IDF")

hpsets$svm_tfidf %>%
  php_c() 
```

# Logistic Regression + TF-IDF 
```{r LR_TFIDF}
lrtfidftrials <- readRDS("trials/lr_tfidf-trials.RDS")

lrtfidftrials <- lrtfidftrials %>%
  group_by(datasets) %>%
  mutate(index = rev(index))

# loss plot by index 
lrtfidftrials %>% 
  loss_i()

# loss plot sorted by trial numbers 
lrtfidftrials %>%
  loss_t()

php(hpsets$lr_tfidf) + 
  labs(title = "Hyperparameters Logistic Regression + TF-IDF")

hpsets$lr_tfidf %>%
  php_c()
```

Inspect outliers
```{r}
hpsets$lr_tfidf %>% filter(param %in% c("balance_param.a", "model_param.c")) %>% group_by(param) %>% filter(value > 40) %>% select(-ub, -lb) %>%
  kable(format = "markdown")
hpsets$lr_tfidf %>% filter(param %in% c("model_param.class_weight")) %>% group_by(param) %>% filter(value > 1000) %>% select(-ub, -lb) %>%
  kable(format = "markdown")
```

# Logistic Regression + Doc2Vec
```{r LR_D2V}
lrd2vtrials <- readRDS("trials/lr_d2v-trials.RDS")

lrd2vtrials <- lrd2vtrials %>%
  group_by(datasets) %>%
  mutate(index = rev(index))

# loss plot by index 
lrd2vtrials %>% 
  loss_i()

# loss plot sorted by trial numbers 
lrd2vtrials %>%
  loss_t()

php(hpsets$lr_d2v) + 
  labs(title = "Hyperparameters Logistic Regression + Doc2Vec")

hpsets$lr_d2v %>%
  php_c()
```

# Random Forest + TF-IDF
```{r rf_tfidf}
rftfidftrials <- readRDS("trials/rf_tfidf-trials.RDS")

rftfidftrials <- rftfidftrials %>%
  group_by(datasets) %>%
  mutate(index = rev(index))

# loss plot by index 
rftfidftrials %>% 
  loss_i()

# loss plot sorted by trial numbers 
rftfidftrials %>%
  loss_t()

php(hpsets$rf_tfidf) + 
  labs(title = "Hyperparameters Random Forest + TF-IDF")

hpsets$rf_tfidf %>%
  php_c()
```

# Random Forest + Doc2Vec
## Loss plot
```{r rf_d2v_loss}
rfd2vtrials <- readRDS("trials/rf_d2v-trials.RDS")

rfd2vtrials <- rfd2vtrials %>%
  group_by(datasets) %>%
  mutate(index = rev(index))

# loss plot by index 
rfd2vtrials %>% 
  loss_i()

# loss plot sorted by trial numbers 
rfd2vtrials %>%
  loss_t()

```

## Hyperparameter Values
```{r rf_d2v}
php(hpsets$rf_d2v) + 
  labs(title = "Hyperparameters Random Forest + Doc2Vec")

hpsets$rf_d2v %>%
  php_c()
```