### 2020-04-25 manuscript/drafts/manuscript.tex
First draft introduction section

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Manuscript drafts},
  pdfauthor={Gerbrich Ferdinands},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage[]{biblatex}
\addbibresource{asreview.bib}

\title{Manuscript drafts}
\author{Gerbrich Ferdinands}
\date{22/04/2020}

\begin{document}
\maketitle

Keywords: Human-Machine interaction, active learning, systematic
reviews, text classification

\hypertarget{introduction}{%
\paragraph{Introduction}\label{introduction}}

\hypertarget{methods}{%
\paragraph{Methods}\label{methods}}

\hypertarget{results}{%
\paragraph{Results}\label{results}}

\hypertarget{discussion}{%
\paragraph{Discussion}\label{discussion}}

\newpage

\hypertarget{introduction-1}{%
\section{Introduction}\label{introduction-1}}

Systematic reviews are top of the bill in building evidence in research.
A systematic review attempts to bring together all studies relevant to
answer a specific research question \autocite{PRISMA-PGroup2015}.
Systematic reviews inform practice and policy \autocite{Gough2002} and
are key in developing clinical guidelines \autocite{Chalmers2007}.

Systematic reviews are costly. Conducting a systematic review involves
the manual screening of thousands of titles and abstracts, identifying
publications relevant to answering the research question. An experienced
reviewer takes on average 30 seconds to screen one title and abstract,
whereas an inexperienced reviewer takes even longer
\autocite{Wallace2010}. Conducting a systematic review typically
requires over a year of work from a team of researchers
\autocite{Borah2017}.

Systematic reviewers face competing demands. The production of a
systematic review is a time-consuming process, but is often bound to a
limited budget and timeframe. Currently, the demand for systematic
reviews exceeds the available time and resources by far
\autocite{Lau2019}. Especially when the need for guidelines is urgent -
such as in the context of the current COVID-19 crisis - it is almost
impossible to provide a review that is both timely and comprehensive. To
ensure a timely review, reducing workload in systematic reviews is
imperative.

With advances in Artificial Intelligence (AI), there has been wide
interest in tools to reduce workload in systematic reviews
\autocite{Harrison2020}. More specifically, the field of AI-aided title
and abstract screening is rapidly evolving. Various learning models have
been proposed, aiming to detect whether a given publication is relevant
or irrelevant to the systematic review. Findings suggest that such
models potentially reduce workload with 30-70\% at the cost of losing
5\% of relevant publications (95\% recall) \autocite{OMara-Eves2015}.

Screening prioritization is a well-established approach in increasing
efficiency in title and abstract screening
\autocite{Cohen2009,Shemilt2014}. In screening prioritization, the
learning model reorders publications to be screened by their likeliness
to be relevant, aiming to present the reviewer with the most relevant
publications first. Such an approach allows for substantial time-savings
in the screening process: Reviewing relevant publications early
facilitates a faster transition of those publications to the next steps
in the review process \autocite{Cohen2009}.\\
Moreover, when combined with a stopping criterion, screening
prioritization can be used to reduce the number of publications needed
to screen. Additionally, several studies report increasing efficiency
beyond reducing workload \autocite{OMara-Eves2015}.

Recent studies have demonstrated the effectiveness of screening
prioritization by means of active learning models
\autocite{Yu2019,Yu2018,Miwa2014}. Active learning is when the model can
iteratively improve its predictions by allowing the model to choose the
data from which it can learn \autocite{Settles2012}. Active learning has
proven to be an efficient strategy in large datasets where labels are
scarce, which makes identifying relevant publications an ideal candidate
for such models.\\
When applied in the screening process, an active learning model can
iteratively improve its relevancy predictions by interacting with the
reviewer. The model decides which publication it wants the reviewer to
screen next, who then provides the model with a label (`relevant',
`irrelevant'). In this active learning cycle, the model can
incrementally improve its predictions on the relevancy of the remaining
unlabeled publications. Based on previous labelling decisions by the
reviewer, the model constantly reorders the remaining publications in
the dataset.

Although the application of (active) learning models in reducing
workload of systematic reviews has been extensively studied, the complex
nature of the field is making it difficult to draw overarching
conclusions about best practice. Moreover, the lack of replication on
data outside the biomedical sciences makes it impossible to draw
conclusions about the general effectiveness of such technologies
\autocite{OMara-Eves2015}. The question remains how different active
learning models for screening prioritization perform across different
review contexts. Hence, additional evaluations of active learning models
are required.

The purpose of this paper is to increase the evidence base of active
learning models for reducing workload in title and abstract screening in
systematic reviews. Combining latest insights from this area, we present
a pipeline of active learning for prioritization screening, called
\texttt{ASReview}. \texttt{ASReview} is an open source and generic tool
such that users can adapt and add modules as they like, encouraging
fellow researchers to replicate findings from previous studies. Working
towards a general consensus in this emerging field, this study
implements various active learning models (in \texttt{ASReview}). Models
are evaluated by performing a simulation on data from six existing
systematic reviews from various research areas. All data presented in
this study are openly published, hoping to facilitate usability and
acceptability of AI-assisted title and abstract screening in the field
of systematic review.

The remaining part of this paper is organized as follows. The Technical
Details section will cover the workings of active learning models for
study selection in systematic reviews on a conceptual level. The method
section describes \ldots{} The results section reports \ldots{} The
discusssion section summarises the findings, comments on them and
summarises the main findings, discusses discuss limitations, draw
conclusion and

\autocite{Marshall2020}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{summarise-models-used}{%
\paragraph{summarise models used}\label{summarise-models-used}}

\hypertarget{propose-pipeline}{%
\paragraph{propose pipeline}\label{propose-pipeline}}

\hypertarget{and-specific-components}{%
\subparagraph{and specific components?}\label{and-specific-components}}

\hypertarget{summarise-datasets-tested}{%
\paragraph{summarise datasets tested}\label{summarise-datasets-tested}}

\hfill\break
\hfill\break

\hypertarget{background}{%
\subsection{Background}\label{background}}

\autocite{OMara-Eves2015} literature review

\autocite{Yu2018a}, \autocite{Yu2019} simulated 32 svm classifiers, on
software engineering. A popular classifier is SVM. succes with HUTM
(fastread), uncertainty, mix of weighting and agressive undersampling,
In terms of Yu et al, we adopt .CT.

SVM - tf-idf on medical data, uncertainty sampling, agressive
undersampling. \autocite{Wallace2010}

abstrackr

SVM + Weighting + uncertainty (bow) produced good methods
\autocite{Miwa2014} Also include social sciences data besides medical
data.

\autocite{Cohen2006} perceptron-based classifier (neural network)

SVM on legal documents (no balancing, certainty ) \autocite{Cormack2014}
in limitations section mentions that LR yields about same results, nb
inferior results.

\autocite{Kilicoglu2009} - SVM, naive bayes, boosting and combinations.
future work should optimize parameters. ``Regarding the base classifiers
used in identifying method- ologically rigorous studies, boosting
consistently strikes the best balance between precision and recall,
whereas naive Bayes in general performs well on recall (demonstrating a
tradeoff between recall and precision), as does polynomial SVM on
precision. The AUC results are mixed, although boosting has a slight
edge overall. These results demonstrate that different classifiers can
be used to satisfy different information needs (SVM for specificity,
naive Bayes for sensitivity, and boosting for balance between the two,
for example).''

Our extensions is that we try different classifiers, on more datasets.

Goal: evaluate performance of different models of the ASReview tool. The
screening process is simulated using ASReview, seeing if the original
inclusions replicate. What would happen if the citation screening would
have been performed using asreview? All datasets accompanying the
systematic reviews are openly published. We built several machine
learning models to perform automated systematic reviews, who are then
applied on existing systematic reviews.

When no balancing is applied, the training data set = labeled data set
\(\mathcal{L}\) The model can query the labels, who serve as the
reviewer, active learning then perform active learning to detect
inclusions.

a machine learning-based citation classification tool to reduce workload
in systematic reviews of drug class efficacy. Using a perceptron
classifier, \href{mailto:WSS@95}{\nolinkurl{WSS@95}}\% = 56.61 in
\autocite{Cohen2006}. (5x2 crossvalidation). Can we beat this? The data

Openness, reproducible,

Benefits:

Adopting some sort of stopping criterion (outside scope of the current
thesis) the reviewer can quit reviewing after having read only a
fraction of candidate studies. meaning the screening process can be
finished after reading a fraction of all candidate studies. Saving hours
of time and resources.

paper organization:

The goal is to gain insight in classifiers other than the widely applied
SVM, overall various research areas. So not only medical sciences.

Although considerable research has been devoted to \ldots, less
attention has been paid to the comparison of different classifiers. Few
studies have evaluated different classifiers in any systematic way
different models to use haven't been investigated in a systematic way
(only in software engineering)

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  the lack of replication of methods is making it impossible to draw any
  overall conclusions about best practice/ best approaches of the
  problem of reducing workload in screening.
\item
  screening prioritization is appealing to systematic reviewers because
  \ldots{}
\item
\item
  implemantation: usabliity and acceptability of such tools amongst
  researchers conducting a systematic review.
\end{enumerate}

lack of studies investigating the effect of different methods over
different research areas. (yu but only software engineering, miwa
perhaps?) There is also a significant lack of examples outside of
healthcare with the exception of one example in software engingeering

However, there is a need for consensus () and transparancy?

So far, however, there has been little discussion about the different
classifiers to be used. Support Vector Machine has been the default in
almost all studies.

Several algorithms to assist the reviewer in the abstract screening
process have been proposed. (they are available in many shapes/sorts).

\hypertarget{indicating-the-structure-of-the-writing}{%
\subparagraph{indicating the structure of the
writing)}\label{indicating-the-structure-of-the-writing}}

principle findings:

For truly evaluate the effectiveness of such methods,

automatically (by machine learning techniques) In the current study we
focus on the effectiveness of different machine learning models in
reducing workload in the citation screening process. Assisting the
reviewer. Combining various machine learning techniques. Combining
recent insights on, we propose ASReview. Approaching the citation
screening phase as a classfication task, The benefit is that ASReview is
dynamic, user can decide open source, multiple classifiers and other
settings,

The current study focus is on reducing workload by prioritizing the most
relevant abstracts. Presenting the abstracts orders the abstracts by
likelihood of relevance, e.g.~th The reviewer sees the most relevant
abstracts first, If the model performs well all relevent abstracts will
be seen much earlier than when abstracts are read at a random order.

Such a solution can save time and resources. Time saving is not the only
benefit: and help to minimize bias..

\hypertarget{proposed-solution}{%
\paragraph{Proposed solution}\label{proposed-solution}}

This study is about how machine learning models can increase efficiency
in systematic reviews. Their main objective is to identify ..

The strategy to reduce workload proposed in the current study is by
prioritizing publications that are deemed most relevant to the
systematic review. As the relevant publications are screened first, the
reviewing process can be quit earlier.

The stage of abstract screening where abstracts are systematically
screened is where a lot is to be gained. This stage is the target of
possible learning algorithms that can assist the reviewer in selecting
the relevant papers. Together with the reviewer /human machine
interaction. The algorithm aims to compute which papers in the pool need
to be excluded and which need to be included, based on the reviewers
decisions. It learns from the reviewers decisions and asks the reviewer
to provide more labels, incrementally improving its class predictions.

The goal of the algorithm defined in the current study is to reduce to
number of abstracts needed to screen (maybe not right term, bit
biomedical). To be more specific, the algorithm aims to present the
reader with the primary studies as soon as possible. This means that at
some point you probably have seen all relevant abstracts and are only
viewing excluded papers, which means you can stop reviewing much earlier
(theoretically spoken). Also reviewing is now much more fun. As compared
to when you have to review all abstracts and you perhaps see only one
relevant abstract every other week/day.

\hypertarget{background-1}{%
\subsubsection{Background}\label{background-1}}

\ldots{} starts with a search for potentially relevant publications.
This initial set of candidate studies need to be manually screened to
identify the publications relevant for answering the research question.
Because the initial set often consists of thousands of papers and the

To gather the findings relevant to answering the research question, a
systematic search is performed. A systematic search starts with
collecting all publications that meet pre-specified eligibility
criteria. From this collection of candidate studies the researcher has
to identify the publications relevant for answering the research
question. Of all candidate studies only a fraction is relevant
{[}\ldots{]}.

As more and more papers are published and reproducibility crisis has
emerged, A systematic search often results in thousands of candidate
studies. Relevant publications are then identified by screening title
and abstract of all candidate studies. This screening process is a
manual task often executed by multiple reviewers to ensure reliability.
This is a time consuming process that weighs heavily on resources.

Most often the SVM classifier is used, popular and very good results.
Also lots of other configurations. However, other classifiers have not
been tested a lot (polygon thing by cohe, naïve bayes and random forest
by \ldots), but mostly SVM still. Also, most research in the medical
sciences (well there are some exceptions of course {[}conversation
between cohen and matwill{]}

A solution is \ldots{}

It is important reflect on research by giving an overview of research
areas which is typically done by a systematic review {[}\ldots{]}.

To review a specific research area, one starts out with an initial
search of thousands of academic papers. All these papers abstracts need
to be screened to find an initial batch of possibly relevant papers.
With now hopefully only a couple of hundred papers left, the researcher
needs to read these papers full-text to arrive at a final selection of
papers that are relevant for the final systematic review {[}this is
prisma process?{]}. This whole processes costs this and this much time
{[}shelmilt{]}.

\hypertarget{technical-details}{%
\paragraph{Technical details}\label{technical-details}}

So you might wonder, how does such an algorithm actually work? Active
learning strategy, starting with a pool of unlabeled abstracts (U). The
reviewer starts labeling some instances in U, creating L. The algorithm
utilizes L to predict labels for all abstracts (classifier), by using a
set of features from the papers called X, for example the text in the
abstract (feature extraction method). Now, it made an initial
classification. The algorithm now aims to improve its classification by
which paper from U will be presented to the reviewer next. By labeling
the next paper, the reviewer provides the algorithm with new information
which the algorithm uses to update its prediction.

We approach asr as a classification problem: All papers obtained in the
systematic search form a pool of instances x. All instances x need to be
classified, e.g.~we want to give them a label y. all x are now part of
U,. Binary classification, either inclusion or exclusion. We want to
classify based on some features from the instances \(x\), feature vector
\(\mathbf{X}\).

By starting of with L, the algorithm utilizes characteristics of X to
predict labels in U.

We want to classify the papers in Where the whole collection of papers
in the systematic search form a pool of instances x, with unknown label
y. \textless x,y\textgreater.

\begin{verbatim}
input:  
    a pool of unlabeled abstracts $\mathcal{U}$ 
    oracle labels a few initial papers, from $\mathcal{U}$, $\mathcal{L}$,

  M = train($\mathcal{L}$)
  Query x isin mathcal{U} (<x, y=?)
  oracle <x, y>

\end{verbatim}

\begin{itemize}
\item
  pool unlabeled abstracts \(\mathcal{U}\)
\item
  labeled data set \(\mathcal{L}\),
\item
  instance \(x\), label \(y\)
\item
  utility measure \(\phi_A(\cdot)\)
\item
  \(x^*_A\) best query instance according to \(\phi_A(\cdot)\)
\end{itemize}

Now there are a few technical details. Many different versions of such
algorithms exist. Many of such algorithms have been described in the
active learning literature and have been applied in the systematic
reviewing process. Not exhaustive, but the algorithm can apply many
different strategies to arrive at its predictions, which can be divided
in following parameters: classifier, feature extraction strategy,
balancing, query strateg y.

To perform all these computations the research was carried out using the
ASReview software by Utrecht University, which has a simulation mode
that you can just input your labelled review file into and perform a
simulation study with it. To be found on GitHub. It has many adjustable
components/is very versatile.

Then if we still have time left we explore the effect of another feature
extraction method, namely doc2vec. This might possibly increase
performance as it performs better at grasping structure/hidden
relations/hierarchy between words in the texts. So, it could be
interesting in more `fuzzy' research areas. A downside is that it takes
more computing time.

Then if there's even more time we might want to compare a different
balance strategy.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

A SR can be divided into phases. Everything starts with a
\textbf{systematic search}, leading to then citation screening is
performed, then full-text screening \autocite{PRISMA-PGroup2015}

What must be the objective of our tool?: It is the tedious task citation
screening part where loads of time can be saved.

models are designed in a `realistic' way (you have some inclusions)

Selecting papers is a two-step process: abstract \& fulltext screening

Binary classification problem. predict whether a paper in the pool is an
inclusion or exclusion, based on labeled instances from \(\mathcal{L}\)
and use training data to understand how input variables are related to
class.

We're building active learning model who takes X as an input an predicts
class using labels from training set. Model improves by deciding asking
more information from the reviewer (oracle), accounting for imbalance.

The researcher has some prior knowledge about the pool, some papers
ought to be included in the SR.

\hypertarget{assumptions}{%
\paragraph{Assumptions}\label{assumptions}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  decisions of the original SR are \textbf{ground truth} (benchmark)
  (oracle)
\end{enumerate}

The inclusion rate is \ldots{} data is imbalanced. what is the
philosophy False negatives must be avoided \ldots{} The cost of a false
negative outweighs the cost of a false positive. Note that we assume the
oracle/original user to hold the truth. This is of course not always the
case.

There are two classes in the data: exlusions and inclusions. The
inclusions are clearly the minority class.

Datasets from the medical and social sciences, software engineering and
public administration. Medical sciences SRs are viewed as more
`strict'/`structured' and social sciences more messy.

\hypertarget{active-learning-for-systematic-reviews}{%
\subsection{active learning for systematic
reviews}\label{active-learning-for-systematic-reviews}}

corpus = all the text:

Active learning = increasing classification performance with every
query. The query strategy determines the way unlabeled papers are
queried to the researcher.

\autocite{modAL2018}

RQ1 - what are good classifiers RQ2 - what are good optimization
strategies

\newpage

\hypertarget{methods-1}{%
\section{Methods}\label{methods-1}}

This study was approved by the Ethics Committee of the Faculty of Social
and Behavioural Sciences of Utrecht University, filed as an amendement
under study 20-104.

All simulations were run using through cartesius EINF-156

Performance of several machine learning models was demonstrated on six
systematic review datasets.\\
study design: retrospective

\hypertarget{models}{%
\subsection{Models}\label{models}}

\emph{x} machine learning models (\(M\)) were built. The models consist
of multiple components, of which the classifier and the feature
extraction strategy varied over the models.

\hypertarget{classifiers}{%
\paragraph{Classifiers}\label{classifiers}}

Every model applies a classifier \(c\) to predict the relevance of
publications in the data. The classfiier predicts the class of a
publication in the dataset expediting a training dataset
\(\mathcal{L}\), predicts the class of an instance given input features
\(\mathbf{X}\)

Logistic Regression (LR) - L2-regularized logistic regression. Logistic
models the posterior directly, naive bayes has higher bias but lower
variance, logistic might perform better when trainign set increases.
\autocite{Ng2002} Naive Bayes (NB) - Naive Bayes assumes all features
are independent given the class value. This is obviously not the case
but still the algorithm performs impressively \autocite{Zhang2004}.
Especially at \ldots{} tasks.

Random Forests (RF) is where a large number of decision trees are fit on
bootstrapped samples of the original data. All trees cast a vote on the
class, which are aggregated into a class prediction for each input
\(\mathbf{X}\) \autocite{Breiman2001}.

Support Vecor Machine (SVM) - finds a multidimensional hyperplane to
separate classes. \autocite{Tong2001}

\hypertarget{word-representation}{%
\paragraph{Word representation}\label{word-representation}}

To predict the class of a publication class (e.g.~whether a publication
should be included or excluded), the classifier uses information from
the publications in the dataset. Examples of such information are titles
and abstracts. However, the classifier cannot predict the publication
class from titles and abstracts as they are. Their textual content to
needs to be mapped to feature vectors. This process of numerically
representing textual content is called `word embeddings'.

A classical example of word embeddings a `bag of words' representation.
For each text in the data set, the number of occurrences of each word is
stored. This leads to n features, where n is the number of distinct
words in the texts \autocite{scikit-learn}. The bag-of-words method is
simplistic and will highly value often occuring but otherwise
meaningless words such as ``and''. Term-frequency Inverse Document
Frequency (TF-IDF) \autocite{Ramos2003} circumvents this problem by
adjusting a term frequency in a text with the inverse docuement
frequency, the frequency of a given word in the entire data set.

The model expedits features of previously labeled publications and Based
on features of the previously labeled inclusions and exclusions, the
model

The features used are title and abstract from every publication. - this
is what is prescribed by prisma, (check!)

(TF-IDF) (Doc2Vec)

\hypertarget{fixed-components}{%
\paragraph{Fixed components}\label{fixed-components}}

All models \(M\) apply the same query strategy of certainty sampling, in
which the presented publication is always the one of which the model is
most certain for it to be relevant.

To decrease the class imbalance in the training data, the model
rebalances the training set by Dynamic Supersampling (DS). DS decreases
the number of irrelevant papers in the training data, whereas the number
of relevant papers are increased (by copy) such that the total number of
samples remains the same. The ratio between relevant and irrelevant
papers is not fixed, but dynamically updated and depends on the number
of training samples, the total number of publications and the ratio
between relevant and irrelevant publications.

\hypertarget{datasets}{%
\subsection{Datasets}\label{datasets}}

Six Systematic Reviews from various research areas were used to simulate
the automated systematic review process. Data were preprocessed from the
original source into a test dataset, containing all publications
obtained in the systematic search. The test datasets are labelled
indicating the relevant and irrelevant publications. Using these labels,
a computer simulation of an automated systematic review can be performed
on the systematic review data.

The test datasets contain title information on all publications obtained
in the search strategy. The instances in the data consist of a title and
an abstract and were labeled to indicate which publications were
included in the systematic review. Instances with missing abstracts and
duplicate instances were removed from the data. The data preprocessing
scripts can be found on the GitHub\footnote{\url{https://github.com/GerbrichFerdinands/asreview-thesis}}

Cohen et al.~collected systematic review datasets from the medical
sciences \autocite{Cohen2006}. All systematic reviews in this database
are on drug efficacy. The \emph{ace} dataset used in the current study
comes from a systematic review on the efficacy of Angiotensin-converting
enzyme (ACE) inhibitors.

The \emph{software} dataset is retrieved from \autocite{Yu2018a}, who
collected datasets on literature reviews from the software engineering
field. This dataset is on fault prediction in software engineering by
\autocite{Hall2012}.

The \emph{nudging} dataset comes from a review from the behavioural
public administration area. The SR includes studies on nudging
healthcare professionals \autocite{Nagtegaal2019}. The data was stored
on the Harvard Dataverse \autocite{Nagtegaal2019a}.

A literature review from field of psychology, \emph{ptsd}. The SR is on
studies applying latent trajectory analyses on posttraumatic stress
after exposure to trauma \autocite{vandeSchoot2017}. The corresponding
data can be found on the Open Science Framework {[}\ldots{]}.

\emph{wilson}, a dataset from the medical sciences
\autocite{Appenzeller-Herzog2020}. TA review on effectiveness treatments
of Wilson disease \autocite{Appenzeller-Herzog2019}.

All datasets started with an initial pool of thousands of papers. A
fraction of these papers where deemed relevant for the SR, with
inclusion rates around 1-2 percent with one outlier of about 5 percent
(Table 1).

Table 1: Statistics on datasets from original systematic reviews.

\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
\multicolumn{1}{c}{} & \multicolumn{3}{c}{Original study} & \multicolumn{3}{c}{Test collection} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
Dataset & Candidate studies & Final inclusions & Inclusion rate (\%) & Candidate studies & Final inclusions & Inclusion rate (\%)\\
\midrule
ace & 2544 & 41 & 1.61 & 2235 & 41 & 1.83\\
nudging & 2006 & 100 & 4.99 & 1847 & 100 & 5.41\\
ptsd & 6185 & 38 & 0.61 & 5031 & 38 & 0.76\\
software & 8911 & 104 & 1.17 & 8896 & 104 & 1.17\\
wilson & 3453 & 26 & 0.75 & 2333 & 23 & 0.99\\
\bottomrule
\end{tabu}

\hypertarget{optimizing-hyperparameters}{%
\subsubsection{Optimizing
hyperparameters}\label{optimizing-hyperparameters}}

Every model component contains hyperparameters, leading to a unique set
of hyperparameters for each model. To maximize model performance, we
need to find optimal values for the hyperparameters. For every model the
optimal hyperparameter values are determined by optimizing on the data
\(d\). The hyperparameters are optimized by running several hundreds of
optimization trials, in which hyperparameter values are sampled from
their possible parameter space. A description of all hyperparameters and
their sample space can be found in the appendix.

Maximum model performance is defined as the average time it takes to
find an inclusion in the data, or more specific: the loss function
minimizes the average number of papers needed to screen to find an
inclusion (e.g.~the area above the curve in the inclusion plot).

The optimization data \(d\) consists of (a subset from) the six
systematic review datasets \(D\) mentioned above. Three different
approaches in composing \(d\) are explored:

\begin{itemize}
\tightlist
\item
  \textbf{one}, where hyperparameters are optimized on only one of the
  six datasets, \(d \in D\). Such hyperparameters are expected to lead
  to maximum performance in the same dataset \(d\).
\item
  \textbf{n}, where hyperparameters are optimized on all six data sets,
  \(d = D\). This optimization approach intends to serve in producing
  the most optimal hyperparameters overall.
\item
  \textbf{n-1}, where hyperparameters are optimized on all six datasets
  but one, \(d \subset D\). Serving as a sensitivity analysis for the
  former condition, e.g.~how sensitive are the hyperparamters. also as a
  cross-validation later on: hyperparameters obtained by training data,
  test data is never seen before. where d are all datasets but the one
  where we want to simulate later on. This results in \(6+6+1=13\) sets
  of hyperparameters for every model.
\end{itemize}

Results were visually inspected to check if an optimum (minimal loss)
has been reached. More trials were run if the loss still seemed to go
down at a quick pace.

The hyperparameter values that were found to lead to a minimum loss
value were visually inspected.

\hypertarget{optimization-results}{%
\subsubsection{Optimization results}\label{optimization-results}}

For every model, 13 sets of hyperparameters were optimized.

\emph{to include: plotting the loss reduction over trials}

Optimal values of the hyperparameters were visually inspected.

As an example, the hyperparameters of the RF\_TF-IDF model are presented
in Figure 1. A panel displays the optimal values for a certain
hyperparameter, where the blue colored dots represent the \textbf{one}
condition, the green dots the \textbf{n-1} condition and the orange dot
represents the optimal hyperparameter value when optmizing over all
datasets (\textbf{n}). The x-axis represents the possible parameter
space where the vertical grey lines mark the boundaries of the
hyperparameter space (if possible). Note that the feature parameters
ngram\_max, split\_ta, use\_keywords and model parameters max\_features
and n\_estimators are categorical.

Overall, the optimal values are distributed over the parameter space.
Outlying values all belong to the \textbf{one} condition where
optimization was dependent on one dataset only. Fore example, the
class\_weight parameter has an outlying value of 11.3 that belongs to
the nudging dataset. (Why this is the case I can only speculate, but it
is worth mentioning that this dataset has a relatively high inclusion
rate of 5.41\%, compared to the other datasets).

\includegraphics{figs/rf_tfidf.eps}

\hypertarget{appendix-x---hyperparameters-and-their-sample-space}{%
\subsection{Appendix x - Hyperparameters and their sample
space}\label{appendix-x---hyperparameters-and-their-sample-space}}

Classifier hyperparameters

\begin{verbatim}
class_weight: normal(0,1) constrained to be > 0
class weight of the inclusions.

\end{verbatim}

Logistic Regression

\begin{verbatim}
c: float. normal(0,1), constrained to be > 0
\end{verbatim}

Support Vector Machine

\begin{verbatim}
gamma: ["auto", "scale"],
Gamma parameter of the SVM model.

C: 
C parameter of the SVM model.

kernel:
SVM kernel type.["linear", "rbf", "poly", "sigmoid"]
\end{verbatim}

Naive Bayes

\begin{verbatim}
"model_param.alpha", # exp(normal(0,1))

\end{verbatim}

Random Forest

\begin{verbatim}
max_features: int (between 6 and 10)
    Number of features in the model.

n_estimators: int between 10 and 100 
    Number of estimators.

model_param.n_estimators" #quniform(10,100,1)
\end{verbatim}

Balance strategy hyperparameters: Dynamic supersampling:

\begin{verbatim}
a: float
    Governs the weight of the 1's. Higher values mean linearly more 1's
    in your training sample.
alpha: float
    Governs the scaling the weight of the 1's, as a function of the
    ratio of ones to zeros. A positive value means that the lower the
    ratio of zeros to ones, the higher the weight of the ones.
b: float
    Governs how strongly we want to sample depending on the total
    number of samples. A value of 1 means no dependence on the total
    number of samples, while lower values mean increasingly stronger
    dependence on the number of samples.

\end{verbatim}

Feature extraction strategy hyperparameters

\begin{verbatim}
split_ta: 0 or 1
   whether titles and abstracts are split

use_keywords: 0 or 1
    whether keywords should be used

\end{verbatim}

TF-IDF

\begin{verbatim}
ngram_max: 1, 2 or 3
    Can use up to ngrams up to ngram_max. For example in the case of 
    ngram_max=2, monograms and bigrams could be used.
\end{verbatim}

Doc2Vec

\begin{verbatim}
vector_size: int (between 32 and 127)
    Output size of the vector.

epochs: int (between 20 and 50)
    Number of epochs to train the doc2vec model.

min_count: int (between 1 and 3)
    Minimum number of occurences for a word in the corpus for it to be 
    included in the model.

window: int (between 5 and 9)
    Maximum distance over which word vectors influence each other.

dm_concat: int 0 or 1
    Whether to concatenate word vectors or not.

dm: int 
    Model to use.
    0: Use distribute bag of words (DBOW).
    1: Use distributed memory (DM).
    2: Use both of the above with half the vector size and concatenate them.

dbow_words: int 0 or 1
    Whether to train the word vectors using the skipgram method.
\end{verbatim}

\hypertarget{simulations}{%
\subsection{Simulations}\label{simulations}}

To demonstrate the models, simulations are performed of the models on
six original systematic reviews.\\
With the labeled datasets, the systematic review can be reproduced but
with now using a machine learning model to optimize the reviewing
process.

The initial training set of the active learning model in the simulation
is one prior included publication and one prior excluded publication,
randomly sampled from the data. This demonstrates a `worst case
scenario' where the reviewer has minimal prior knowledge on publications
in the data. For this simulations there is no need for a human reviewer
since the labels in the data can be queried. The oracle is not the
reviewer but the labels in the data. The model retrains every time after
a label is provided by the oracle. The simulation ends when all
publications in the dataset have been queried.

To account for variance, every simulation is repeated for 15 trials
\(t\).

\hypertarget{statistical-analysis}{%
\subsubsection{Statistical analysis}\label{statistical-analysis}}

Results are aggregated over 15 runs of every simulation.

\hypertarget{measures}{%
\paragraph{measures}\label{measures}}

\hypertarget{comparison-over-conditions}{%
\paragraph{comparison over
conditions}\label{comparison-over-conditions}}

\hypertarget{the-software}{%
\section{The software}\label{the-software}}

ASReview takes the following parameters/arguments: We now have 75
combinations. for every for every model (5), for every dataset (5) and
for every set of optimized hyperparameters (3), a simulation study
consisting trials is performed. From these \(5*5*3=75\) simulation
studies, performance of the different models is evaluated.

\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X}
\toprule
  & Configurations\\
\midrule
Models & 2-Layer Neural Network, Naive Bayes, Random Forest, Support Vector Machine, Logistic Regression\\
Query Strategies & Cluster Sampling, Maximum Sampling, Cluster * Maximum  Sampling, Maximum * Uncertainty Sampling, Maximum * Random Sampling, Cluster * Uncertainty Sampling, Cluster * Random Sampling\\
Feature extraction strategies & Doc2Vec, TF-IDF, sbert, embeddingIdf\\
\bottomrule
\end{tabu}

Use these inputs to predict relevance of papers.

\hypertarget{stage-1-hyperparameter-optimization}{%
\subsubsection{Stage 1: hyperparameter
optimization}\label{stage-1-hyperparameter-optimization}}

Or, more specific:

\begin{tabular}{ll}
\toprule
Models & Feature extraction strategies\\
\midrule
dense\_nn & doc2vec\\
nb & tfidf\\
rf & tfidf\\
svm & doc2vec\\
lr & tfidf\\
\bottomrule
\end{tabular}

\hypertarget{performance-metrics}{%
\paragraph{Performance metrics}\label{performance-metrics}}

For each model, Several metrics are used to compare performance of
different models over datasets,

\begin{longtable}[]{@{}lllllll@{}}
\toprule
\begin{minipage}[b]{0.29\columnwidth}\raggedright
Dataset\strut
\end{minipage} & \begin{minipage}[b]{0.03\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright
Naive Bayes\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright
Random Forests\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright
Support Vector Machine\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright
Logistic Regression\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright
Dense Neural Network\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.29\columnwidth}\raggedright
ptsd\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
?\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.29\columnwidth}\raggedright
ace\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
?\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.29\columnwidth}\raggedright
hall\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
?\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.29\columnwidth}\raggedright
nagtegaal\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
?\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.29\columnwidth}\raggedright
\ldots.\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
?\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

The goal is twofold: we want to identify all relevant papers, as fast as
we can. Tradeoff: identifying all relevant papers and reducing workload.
A good metric to evaluate this is..

What is more important: recall or precision?

Recall more highly valued than precision.

What about class imbalance?

\hypertarget{rrf}{%
\paragraph{RRF}\label{rrf}}

Amount of relevant references found after having screened a certain
percentage of the total number of abstracts.

\hypertarget{work-saved-over-sampling-wss}{%
\paragraph{Work saved over sampling
(WSS)}\label{work-saved-over-sampling-wss}}

Indicates how much time can be saved, at a given level of recall. WSS is
in terms of the percentage of abstracts that don't have to be screened
by the researcher. Typically, WSS is measured at a recall of 0.95.
Reasonable because..

\[\texttt{WSS} = \frac{TN + FN}{N} - (1- recall) \]

\hypertarget{raoul}{%
\paragraph{Raoul}\label{raoul}}

\hypertarget{utility}{%
\paragraph{Utility?}\label{utility}}

\hypertarget{f-measure}{%
\paragraph{F-measure}\label{f-measure}}

\hypertarget{rocauc}{%
\paragraph{ROC/AUC}\label{rocauc}}

Is performance related to some characteristic (n, inclusion rate,
\ldots)

? How to compare outcomes of 3 different optimization strategies?

\newpage

\hypertarget{results-1}{%
\section{Results}\label{results-1}}

\begin{tabular}{l|r|r|r|r|r|r}
\hline
  & ace & nudging & ptsd & software & virus & wilson\\
\hline
BCTD & 0 & 0 & 0 & 0 & 0 & 0\\
\hline
RCTD & 0 & 0 & 0 & 0 & 0 & 0\\
\hline
SCTD & 0 & 0 & 0 & 0 & 0 & 0\\
\hline
LCTD & 0 & 0 & 0 & 0 & 0 & 0\\
\hline
NCTD & 0 & 0 & 0 & 0 & 0 & 0\\
\hline
BCDD & 0 & 0 & 0 & 0 & 0 & 0\\
\hline
RCDD & 0 & 0 & 0 & 0 & 0 & 0\\
\hline
SCDD & 0 & 0 & 0 & 0 & 0 & 0\\
\hline
LCDD & 0 & 0 & 0 & 0 & 0 & 0\\
\hline
NCDD & 0 & 0 & 0 & 0 & 0 & 0\\
\hline
\end{tabular}

\newpage

\hypertarget{discussion-1}{%
\section{Discussion}\label{discussion-1}}

\begin{itemize}
\tightlist
\item
  we look for final inclusions but we screen only the abstracts (do they
  satisfy the imformation need (blake (page 19 omara et evs)))
\end{itemize}

future research: - stopping rule is not discussed -
computation/retraining time

strengths:

\begin{itemize}
\tightlist
\item
  open data
\item
  different research areas
\item
  different models on same dataset
\item
  different datasets on same model
\end{itemize}

limitations

future research - all models save time, difficult to distinguish
performance over datasets, especially when applied on a dataset of which
no prior information is known (e.g.~inclusions isn't known in practice).
Perhaps go for other criteria like the fastest model, replicate study
with computation time?

Simulating the title and abstract screening process, models are evaluted
on their capability/speed of detecting the final inclusions. However, in
a manual SR these final inclusions are selected after reading the
fulltext. Information the text mining tool does not have. To truly
assess the added value of such a tool, models should be evaluated on
their capability of detecting the abstract inclusions. Call for
systematic reviewers to openly publish need for open data containing
abstract inclusions, not only final inclusions!

\newpage

\hypertarget{appendix-a---list-of-definitions}{%
\section{Appendix A - list of
definitions}\label{appendix-a---list-of-definitions}}

\hypertarget{feature-extraction-strategies}{%
\subsubsection{Feature Extraction
Strategies}\label{feature-extraction-strategies}}

split\_ta = overall hyperparameter

\hypertarget{tf-idf}{%
\paragraph{TF-IDF}\label{tf-idf}}

\hypertarget{hyperparameters}{%
\subparagraph{hyperparameters}\label{hyperparameters}}

\begin{verbatim}
ngram_max: int
        Can use up to ngrams up to ngram_max. For example in the case of
        ngram_max=2, monograms and bigrams could be used.
\end{verbatim}

\hypertarget{doc2vec}{%
\paragraph{Doc2Vec}\label{doc2vec}}

Predicts words from context. Aims at capturing the relations between
word (man-woman, king-queen). \autocite{Le2014}. Using a neural network.

using Continuous Bag-of-Words (CBOW), Skip-Gram model, \ldots. Word
vector \emph{W} and extra: document vector \emph{D}, trained to predict
words in the text.

From gensim \autocite{Rehurek2010}.

\begin{verbatim}
    Arguments
    ---------
    vector_size: int
        Output size of the vector.
    epochs: int
        Number of epochs to train the doc2vec model.
    min_count: int
        Minimum number of occurences for a word in the corpus for it to
        be included in the model.
    workers: int
        Number of threads to train the model with.
    window: int
        Maximum distance over which word vectors influence each other.
    dm_concat: int
        Whether to concatenate word vectors or not.
        See paper for more detail.
    dm: int
        Model to use.
        0: Use distribute bag of words (DBOW).
        1: Use distributed memory (DM).
        2: Use both of the above with half the vector size and concatenate
        them.
    dbow_words: int
        Whether to train the word vectors using the skipgram metho

\end{verbatim}

\hypertarget{sbert}{%
\paragraph{SBERT}\label{sbert}}

BERT-base model with mean-tokens pooling \autocite{Reimers2019}

\hypertarget{embeddingidf}{%
\paragraph{embeddingIdf}\label{embeddingidf}}

This model averages the weighted word vectors of all the words in the
text, in order to get a single feature vector for each text. The weights
are provided by the inverse document frequencies

\hypertarget{models-1}{%
\subsubsection{Models}\label{models-1}}

\hypertarget{naive-bayes}{%
\paragraph{Naive Bayes}\label{naive-bayes}}

Naive Bayes assumes all features are independent given the class value.
\autocite{Zhang2004}

ASReview uses the \texttt{MultinomialNB} from the scikit-learn package
\autocite{scikit-learn}, that implements the naive Bayes algorithm for
multinomially distributed data. \texttt{nb}

Hyperparameters

\begin{itemize}
\tightlist
\item
  alpha - accounts for features not present in learning samples and
  prevents zero probabilities in further computations.
\end{itemize}

\hypertarget{random-forests}{%
\paragraph{Random Forests}\label{random-forests}}

A number of decision trees are fit on bootstrapped samples of the
original data, \autocite{Breiman2001} RandomForestClassifier from
sklearn

Arguments --------- n\_estimators: int Number of estimators.
max\_features: int Number of features in the model. class\_weight: float
Class weight of the inclusions. random\_state: int, RandomState Set the
random state of the RNG. """

\hypertarget{support-vector-machine}{%
\paragraph{Support Vector Machine}\label{support-vector-machine}}

Arguments --------- gamma: str Gamma parameter of the SVM model.
class\_weight: class\_weight of the inclusions. C: C parameter of the
SVM model. kernel: SVM kernel type. random\_state: State of the RNG.

\hypertarget{logistic-regression}{%
\paragraph{Logistic Regression}\label{logistic-regression}}

\hypertarget{dense-neural-network}{%
\paragraph{\texorpdfstring{\textbf{Dense Neural
Network}}{Dense Neural Network}}\label{dense-neural-network}}

\hypertarget{query-strategies}{%
\subsubsection{Query Strategies}\label{query-strategies}}

\begin{itemize}
\tightlist
\item
  Max - Choose the most likely samples to be included according to the
  model
\item
  Uncertainty - choose the most uncertain samples according to the model
  (i.e.~closest to 0.5 probability) \autocite{Lewis1994}
\item
  Random - randomly selects abstracts with no regard to model assigned
  probabilities.
\item
  Cluster - Use clustering after feature extraction on the dataset. Then
  the highest probabilities within random clusters are sampled
\end{itemize}

The following combinations are simulated:

\begin{itemize}
\tightlist
\item
  cluster
\item
  max
\item
  cluster * random
\item
  cluster * uncertainty
\item
  max * cluster
\item
  max * random
\item
  max * uncertainty
\end{itemize}

\hypertarget{balance-strategies}{%
\subsubsection{Balance Strategies}\label{balance-strategies}}

\hypertarget{amount-of-training-data}{%
\subsubsection{amount of training data}\label{amount-of-training-data}}

\begin{itemize}
\tightlist
\item
  n\_instances = number of papers queried each query
\item
  n\_queries = number of queries
\item
  n\_prior\_included: 5
\item
  n\_prior\_excluded:
\end{itemize}

\hypertarget{combinations}{%
\section{Combinations}\label{combinations}}

This leads to 119 combinations of configurations.

\begin{itemize}
\tightlist
\item
  Naive bayes only goes with tfidf feature extraction.
\item
  For the feature extraction strategies we will focus on doc2vec and
  tfidf. (but will compute all 4)
\item
  This leads to 3 * 7 * 4 * 3 + 1 * 7 * 1 * 3 = 273 combinations.
\end{itemize}

See appendix A for a table containing all 273 combinations.

\hypertarget{cross-validation}{%
\subsection{Cross-validation}\label{cross-validation}}

Should give an accurate estimate of maximum performance / future
systematic reviews to be performed.

\hypertarget{appendix-b---combinations}{%
\section{Appendix B - combinations}\label{appendix-b---combinations}}

\begin{longtable}{lll}
\toprule
Model & Query Strategy & Feature extraction strategy\\
\midrule
\endfirsthead
\multicolumn{3}{@{}l}{\textit{(continued)}}\\
\toprule
Model & Query Strategy & Feature extraction strategy\\
\midrule
\endhead
\
\endfoot
\bottomrule
\endlastfoot
dense\_nn & cluster & doc2vec\\
dense\_nn & max & doc2vec\\
dense\_nn & max * cluster & doc2vec\\
dense\_nn & max * uncertainty & doc2vec\\
dense\_nn & max * random & doc2vec\\
\addlinespace
dense\_nn & cluster * uncertainty & doc2vec\\
dense\_nn & cluster * random & doc2vec\\
dense\_nn & cluster & tfidf\\
dense\_nn & max & tfidf\\
dense\_nn & max * cluster & tfidf\\
\addlinespace
dense\_nn & max * uncertainty & tfidf\\
dense\_nn & max * random & tfidf\\
dense\_nn & cluster * uncertainty & tfidf\\
dense\_nn & cluster * random & tfidf\\
dense\_nn & cluster & sbert\\
\addlinespace
dense\_nn & max & sbert\\
dense\_nn & max * cluster & sbert\\
dense\_nn & max * uncertainty & sbert\\
dense\_nn & max * random & sbert\\
dense\_nn & cluster * uncertainty & sbert\\
\addlinespace
dense\_nn & cluster * random & sbert\\
dense\_nn & cluster & embeddingIdf\\
dense\_nn & max & embeddingIdf\\
dense\_nn & max * cluster & embeddingIdf\\
dense\_nn & max * uncertainty & embeddingIdf\\
\addlinespace
dense\_nn & max * random & embeddingIdf\\
dense\_nn & cluster * uncertainty & embeddingIdf\\
dense\_nn & cluster * random & embeddingIdf\\
nb & cluster & tfidf\\
nb & max & tfidf\\
\addlinespace
nb & max * cluster & tfidf\\
nb & max * uncertainty & tfidf\\
nb & max * random & tfidf\\
nb & cluster * uncertainty & tfidf\\
nb & cluster * random & tfidf\\
\addlinespace
rf & cluster & doc2vec\\
rf & max & doc2vec\\
rf & max * cluster & doc2vec\\
rf & max * uncertainty & doc2vec\\
rf & max * random & doc2vec\\
\addlinespace
rf & cluster * uncertainty & doc2vec\\
rf & cluster * random & doc2vec\\
rf & cluster & tfidf\\
rf & max & tfidf\\
rf & max * cluster & tfidf\\
\addlinespace
rf & max * uncertainty & tfidf\\
rf & max * random & tfidf\\
rf & cluster * uncertainty & tfidf\\
rf & cluster * random & tfidf\\
rf & cluster & sbert\\
\addlinespace
rf & max & sbert\\
rf & max * cluster & sbert\\
rf & max * uncertainty & sbert\\
rf & max * random & sbert\\
rf & cluster * uncertainty & sbert\\
\addlinespace
rf & cluster * random & sbert\\
rf & cluster & embeddingIdf\\
rf & max & embeddingIdf\\
rf & max * cluster & embeddingIdf\\
rf & max * uncertainty & embeddingIdf\\
\addlinespace
rf & max * random & embeddingIdf\\
rf & cluster * uncertainty & embeddingIdf\\
rf & cluster * random & embeddingIdf\\
svm & cluster & doc2vec\\
svm & max & doc2vec\\
\addlinespace
svm & max * cluster & doc2vec\\
svm & max * uncertainty & doc2vec\\
svm & max * random & doc2vec\\
svm & cluster * uncertainty & doc2vec\\
svm & cluster * random & doc2vec\\
\addlinespace
svm & cluster & tfidf\\
svm & max & tfidf\\
svm & max * cluster & tfidf\\
svm & max * uncertainty & tfidf\\
svm & max * random & tfidf\\
\addlinespace
svm & cluster * uncertainty & tfidf\\
svm & cluster * random & tfidf\\
svm & cluster & sbert\\
svm & max & sbert\\
svm & max * cluster & sbert\\
\addlinespace
svm & max * uncertainty & sbert\\
svm & max * random & sbert\\
svm & cluster * uncertainty & sbert\\
svm & cluster * random & sbert\\
svm & cluster & embeddingIdf\\
\addlinespace
svm & max & embeddingIdf\\
svm & max * cluster & embeddingIdf\\
svm & max * uncertainty & embeddingIdf\\
svm & max * random & embeddingIdf\\
svm & cluster * uncertainty & embeddingIdf\\
\addlinespace
svm & cluster * random & embeddingIdf\\
lr & cluster & doc2vec\\
lr & max & doc2vec\\
lr & max * cluster & doc2vec\\
lr & max * uncertainty & doc2vec\\
\addlinespace
lr & max * random & doc2vec\\
lr & cluster * uncertainty & doc2vec\\
lr & cluster * random & doc2vec\\
lr & cluster & tfidf\\
lr & max & tfidf\\
\addlinespace
lr & max * cluster & tfidf\\
lr & max * uncertainty & tfidf\\
lr & max * random & tfidf\\
lr & cluster * uncertainty & tfidf\\
lr & cluster * random & tfidf\\
\addlinespace
lr & cluster & sbert\\
lr & max & sbert\\
lr & max * cluster & sbert\\
lr & max * uncertainty & sbert\\
lr & max * random & sbert\\
\addlinespace
lr & cluster * uncertainty & sbert\\
lr & cluster * random & sbert\\
lr & cluster & embeddingIdf\\
lr & max & embeddingIdf\\
lr & max * cluster & embeddingIdf\\
\addlinespace
lr & max * uncertainty & embeddingIdf\\
lr & max * random & embeddingIdf\\
lr & cluster * uncertainty & embeddingIdf\\
lr & cluster * random & embeddingIdf\\*
\end{longtable}

\hypertarget{appendix-c---supercomputer-cartesius}{%
\section{Appendix C - supercomputer
Cartesius}\label{appendix-c---supercomputer-cartesius}}

500,000 SBU

Running on Cartesius is charged in System Billing Units (SBUs), and
charging is based on the wall clock time of a job. On fat and thin
nodes, an SBU is equal to using 1 core for 1 hour (a core hour), or 1
core for 20 minutes on a GPU node. Since compute nodes are allocated
exclusively to a single job at a time, you will be charged for all cores
on that node - even if you are using less.

In the current study, the classifier and the feature extraction strategy
are varied, whereas the query and balance strategy remain fixed. In the
current study only a fraction of all possible configurations are tested
for the sake of brevity. There are many more options available and open
to exploration.

\printbibliography[title=References]

\end{document}