### 2020-05-01 manuscript/drafts/manuscript.Rmd
Update manuscript and dataset table

---
title: "Manuscript drafts"
author: "Gerbrich Ferdinands"
date: "22/04/2020"
output:
  word_document: default
  citation_package: biblatex
  pdf_document:
bibliography: asreview.bib
csl: systematic-reviews.csl
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

options(tinytex.verbose = TRUE)

library(tidyverse)
library(knitr)
library(xtable)
library(kableExtra)
library(captioner)
library(glue)
library(rbbt)

# load model parameters 
source("asr-parameters.R")

#pathRdata <- "data/"
pathRdata <- "../../datasets/data_statistics/"
```

```{r}
# configurations table 
tab <- lapply(words, FUN = function(x) paste(x, collapse = ", "))
tab <- do.call(rbind, tab)
# remove row training data (28.02)
colnames(tab) <- c("Configurations")
tab <- tab[-4,]
```

```{r}
# combinations table
params[[4]] <- NULL # remove training data
combins <- expand.grid(params) %>%
  arrange(model, feature_extraction)

# rows to drop
droprows <- combins$model == "nb" & combins$feature_extraction != "tfidf"
combins <- combins[!droprows, ]
```

```{r all tables, results = 'hide'}
table_nums <- captioner(prefix = "Table")
# table  on datasets statistics
table_nums(name = "datasets", caption = "Statistics on datasets from original systematic reviews.")
```

Keywords: Human-Machine interaction, active learning, systematic reviews, text classification

#### Introduction 
<!-- This study aims to contribute to this growing area of research by exploring  -->

#### Methods
<!-- A convenience sample of 5 existing systematic reviews on varying topics was collected. -->

#### Results
#### Discussion

\newpage

# Introduction
Systematic reviews are top of the bill in building evidence in research. A systematic review brings together all studies relevant to answer a specific research question [@PRISMA-PGroup2015]. Systematic reviews inform practice and policy [@Gough2002] and are key in developing clinical guidelines [@Chalmers2007]. 
<!-- PRISMA definition:  -->
<!-- > "A systematic review attempts to collate all relevant evidences that fits pre-specified eligibility criteria to answer a specific research -->
<!-- question. 
add that this is a systematic/rigorous process?-->
However, systematic reviews are costly because they involve the manual screening of thousands of titles and abstracts, identifying publications relevant to answering the research question. An experienced reviewer takes on average 30 seconds to screen one title and abstract, whereas an inexperienced reviewer takes even longer [@Wallace2010]. Conducting a systematic review typically requires over a year of work by a team of researchers [@Borah2017]. 
<!-- Besides time-consuming, the process of title and abstract screening is prone to human error [@Wang2020]. -->
<!-- (14 to 77,910, trimmed mean 1,286).  -->
<!-- Need to be screened by multiple researchers.  -->

Moreover, systematic reviewers are often bound to a limited budget and timeframe. Currently, the demand for systematic reviews exceeds the available time and resources by far [@Lau2019]. Especially when the need for guidelines is urgent - such as in the context of the current COVID-19 crisis - it is almost impossible to provide a review that is both timely and comprehensive. To ensure a timely review, reducing workload in systematic reviews is imperative.

With advances in Artificial Intelligence (AI), there has been wide interest in tools to reduce workload in systematic reviews [@Harrison2020]. Various learning models have been proposed, aiming to predict whether a given publication is relevant or irrelevant to the systematic review. Findings suggest that such models potentially reduce workload with 30-70% at the cost of losing 5% of relevant publications (95% recall) [@OMara-Eves2015].
<!-- To date, several studies have begun to examine the use of learning algorithms for screening prioritization. -->

Screening prioritization is a well-established approach in increasing efficiency in title and abstract screening [@Cohen2009; @Shemilt2014]. In screening prioritization, the learning model reorders publications to be screened by their likeliness to be relevant. The model presents the reviewer with the publications which are most likely to be relevant first, thereby expediting the process of finding all of the relevant publications. Such an approach allows for substantial time-savings in the screening process. Reviewing relevant publications early facilitates a faster transition of those publications to the next steps in the review process [@Cohen2009]. Additionally, several studies report increasing efficiency beyond saving time [@OMara-Eves2015].

<!-- assist the reviewer in the screening process -->
<!-- Several studies have proposed active learning models for screening prioritization -->
Recent studies have demonstrated the effectiveness of screening prioritization by means of active learning models [@Yu2018; @Yu2018; @Miwa2014; @Cormack2014; @Cormack2015; @Wallace2010]. Active learning is when the model can iteratively improve its predictions by allowing the model to choose the data from which it can learn [@Settles2012]. Active learning has proven to be an efficient strategy in large datasets where labels are scarce, which makes title and abstract screening an ideal candidate for such models. When applied in screening prioritization, the reviewer screens publications that are presented by an active learning model. Subsequently, the active learning model learns from the reviewers' decision ('relevant', 'irrelevant') and uses this knowledge in selecting the next publication to be screened by the reviewer.

Although prior studies on the application of active learning models in title and abstract screening have addressed important issues [@Yu2018; @Yu2019; @Miwa2014; @Wallace2010], the complex nature of the field is making it difficult to draw overarching conclusions about best practice [@OMara-Eves2015]. First, whilst there exists a scala of classification techniques, research to date has not yet evaluated techniques other than Support Vector Machines (SVM) [@Yu2018; @Yu2018; @Miwa2014; @Wallace2010]. Second, Miwa et al. [@Miwa2014] were the only study so far to make a direct comparison of model performance across research areass and found that active learning was more difficult on data from the social sciences compared to data from the medical sciences. The lack of replication on data from varying research contexts makes it impossible to draw conclusions about the general effectiveness of such techniques [@OMara-Eves2015; @Marshall2020]. The question remains how active learning models for screening prioritization perform across different (1) classification techniques and (2) review contexts. Hence, additional evaluations of active learning models are required.

<!-- on the reviewers screening decision ('relevant', 'irrelevant'), the model can incrementally improve its relevancy predictions on the remaining publications, and again select the next publication to be screened.   -->

<!-- The latest state-of-the-art methods implement active learning, where the model improves  -->
<!-- The algorithm aims to compute which papers in the pool need to be excluded and which need to be included, based on the reviewers decisions.  -->

<!-- The algorithm aims to detect whether a given publication is relevant or irrelevant to the systematic review and reorders the publications by their likeliness of relevance; the reviewer sees most relevant papers first!  -->
<!-- thereby its predictions can be  is allowed to choose the data points it would like to learn from, thus allowing it to learn more efficiently.  -->
<!-- Active learning systems attempt to overcome the labeling bottleneck by asking -->
<!-- queries in the form of unlabeled instances to be labeled by an oracle (e.g., a human -->
<!-- annotator). -->
<!--  that a machine learning algorithm can perform better with less training if it is allowed to choose the data from which it learns. An active learner may pose "queries," usually in the form of unlabeled data instances to be labeled by an "oracle" (e.g., a human annotator) that already understands the nature of the problem -->
<!-- in which a model can choose the data points it would like to learn from, thus allowing it to learn more efficiently -->
<!-- its predictions by interactively querying the user to provide labels to unlabeled data .  -->
<!-- In title and abstract screening this means that the active learning -->
<!-- is allowed to improve its predictions by choosing the data it wants to learn from.  -->

<!-- Although the effectiveness of active learning models in reducing workload in title and abstract screening has been extensively studied,  -->

The purpose of the current paper is to increase the evidence base of active learning models for reducing workload in title and abstract screening in systematic reviews. Combining latest insights from this area, we propose seven different active learning models for the purpose of identifying relevant publications in systematic review datasets. The models were chosen to maximize the number of identified relevant publications, while minimizing the number of publications needed to screen. 
Working towards a general consensus in this emerging field, model performance was asessed by conducting a retrospective simulation on six systematic review datasets. Datasets were collected from the fields of medicine, software engineering, psychology, behavioural public administration, and virology to assess generalizability of the models across research contexts. The models, datasets and simulations are implemented in a pipeline of active learning for prioritization screening, called $\texttt{ASReview}$ [@ASReview2020]. $\texttt{ASReview}$ is an open source and generic tool such that users can adapt and add modules as they like, encouraging fellow researchers to replicate findings from previous studies. All scripts and data used are openly published to facilitate usability and acceptability of AI-assisted title and abstract screening in the field of systematic review.

The remaining part of this paper is organized as follows. The Technical Details section will cover the workings of active learning models for study selection in systematic reviews on a conceptual level. 
The method section describes ... The results section reports ... The discusssion section summarises the findings, comments on them and  summarises the main findings, discusses 
discuss limitations, draw conclusion and

<!-- We demonstrate effectiveness of a generic/flexible tool for assisting researchers in the screening phase..? -->
<!-- To allow comparison of different models they need to be applied on the same dataset.  -->

<!-- "Screening is performed in two steps, first preliminarily based on title and abstract, -->
<!-- then based on full-text."  -->

<!-- add: current consensus active learning are state of the art: how do they woroK; by prioritizatio (put in a picture)  -->

<!-- #### identifying the niche  -->

<!-- By evaluating different configurations of an active learning model on different review contexts.  -->
<!-- status quo: prioritization + active learning? + certainty [@Miwa2014], but classifier  -->
<!-- Although several studies attempt to evaluate such methods,  -->
<!-- It makes  -->
<!-- although  tackle issues in the (complex) area,  -->
<!-- Furthermore, few studies have investigated the use of different classifiers in any systematic way -->
<!-- AI-aided title and abstract screening has been proven succesful in reducing workload, but a lack of replication and  -->

<!-- Furthermore, a comparison mode in public health and clinicial medicing [@Miwa2014] compared different classifier on different datasets and found differential performance on 'messier' social sciences datasets.   -->

<!-- #### Occupying -->
<!-- ##### (listing purpose of new research,  -->

<!-- The tool is generic such that any researcher can adjust modules to increase efficiency in this field.   -->
 <!-- [@OConnor2019] -->

<!-- ##### listing questions,  -->
<!-- This study .  -->
<!-- research questions:  -->
<!-- RQ1 – how -->
<!-- -	RQ1a – does classifier performance vary over different research areas? -->
<!-- (in what terms does it perform best)  -->

<!-- ASReview aims to reduce workload in the citation screening process by prioritizing the most relevant abstracts.  -->
<!-- ASReview combines in text classification  -->

<!-- Furthermore, this study makes a contribution to ...  by demonstrating a reproducible and flexible tool,  -->
<!-- Furthermore, a generic and reproducible workflow is presented to facilitate, hopefully to increase acceptability and usability by researchers performing systematic reviews.  -->

# Methods 
## Task Description
The screening process of a systematic review starts with all publications obtained in the search. The task is to identify which of these publications are relevant, by screening the publications at the title and abstract level.  In active learning for screening prioritiztaion, the screening process proceeds as follows:

- Start with the set of all publications $x$ with an unknown label, $\mathcal{U}$.
- The reviewer provides a label for a few publications $x \in \mathcal{U}$, creating an set of labeled publications $\mathcal{L}$. The publications can be labeled as relevant $x_R$ or irrelevant $x_I$.
- The active learning cycle starts:

  1. A classifier is trained on the labeled publications, $C = \texttt{train}(\mathcal{L})$

  2. The classifier predicts labels for all unlabelled publications, $C(\mathcal{U})$

  3. Based on the predictions by $C$, the model selects the most relevant publication $x \in \mathcal{U}$

  4. The model asks the reviewer to screen this publication, $x_?$

  5. The reviewer screens the publication and provides a label, $x_R$ or $x_I$

  6. The labeled publication is added to the training data, $\mathcal{L}$

  7. Back to step 1.

- In this active learning cycle, the model can incrementally improve its predictions on the remaining unlabeled publications. The relevant publications are identified as early in the process as possible. The reviewer and the model keep interacting until the reviewer decides to stop or until all publications in $\mathbf{U}$ have been labelled.

<!-- Active learning for screening prioritization denotes the scenario in which the reviewer is labeling references that are presented by a machine learning model. The machine learning model learns from the reviewers' decision and uses this knowledge in selecting the next reference that will be presented to the reviewer. -->

<!-- - then active learning cycle starts in which the user makes a decision after each iteration of the model which is then used to re-train the model. -->

<!-- - instance $x$, label $y$ -->
<!-- - utility measure $\phi_A(\cdot)$ -->
<!-- - $x^*_A$ best query instance according to $\phi_A(\cdot)$ -->

<!-- The reviewer starts labeling some instances in $\mathcal{U}$, creating labeled data set $\mathcal{L}$. -->
<!-- This is the starting point for the active learning model. -->
<!-- The model trains a classifier who predicts ...  -->
<!-- The reviewer and the are interacting in the active learning cycle. -->
<!-- Based on previous labelling decisions by the reviewer, the model constantly reorders the remaining publications in the dataset. -->

<!-- # Algorithm 1 -->
<!-- Just a sample algorithmn -->
<!-- \begin{algorithm}[H] -->
<!-- \DontPrintSemicolon -->
<!-- \SetAlgoLined -->
<!-- \KwResult{Write here the result} -->
<!-- \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output} -->
<!-- \Input{Write here the input} -->
<!-- \Output{Write here the output} -->
<!-- \BlankLine -->
<!-- \While{While condition}{ -->
<!--     instructions\; -->
<!--     \eIf{condition}{ -->
<!--         instructions1\; -->
<!--         instructions2\; -->
<!--     }{ -->
<!--         instructions3\; -->
<!--     } -->
<!-- } -->
<!-- \caption{While loop with If/Else condition} -->
<!-- \end{algorithm}  -->

<!-- meaning the reviewer the reviewer screens as little as possible while inclusion rate = high (not true but is nice statistic?) -->
<!-- The model is balancing two goals: obtain all relevant publications with as little as reviewing possible. -->
<!-- Active learning for screening prioritization denotes the scenario in which the reviewer is labeling references that are presented by a machine learning model. The machine learning model learns from the reviewers' decision and uses this knowledge in selecting the next reference that will be presented to the reviewer. -->

<!-- #### Assumptions  -->
<!-- 1) decisions of the original SR are **ground truth** (benchmark) (oracle) -->
<!-- study assumptions: -->
<!-- - simulation mode  -->
<!-- - assumptions (prior knowledge etc, oracle = reviewer). -->

<!-- A SR can be divided into phases.  -->
<!-- Everything starts with a **systematic search**, leading to -->
<!-- then citation screening is performed, then full-text screening -->
<!-- [@PRISMA-PGroup2015] -->

<!-- Selecting papers is a two-step process: abstract & fulltext screening -->

<!-- x -->

<!-- Now there are a few technical details. Many different versions of such algorithms exist. Many of such algorithms have been described in the active learning literature and have been applied in the systematic reviewing process.  Not exhaustive, but the algorithm can apply many different strategies to arrive at its predictions, which can be divided in following parameters: classifier, feature extraction strategy, balancing, query strateg y.  -->

<!-- The current section describes how common issues in identifying relevant publication can be dealt with.  -->
<!-- key components  -->

## Technical details
A more detailed account of the active learning models is given in the following section.
The structure and functions of the key components of the models will be introduced to clarify the choices made in the design of the current study.

#### Classification
To make predictions on the unlabeled publications, a classifier is trained on features from the set of previously labeled publications. 
<!-- The classifier predicts class of a given unlabeled publication.  -->
A technique widely used in classification tasks is the Support Vector Machine (SVM). SVMs separate the data into classes by finding a multidimensional hyperplane [@Tong2001; @Kremer2014]. SVMs have been proven to be effective in active learning models for screening prioritization [@Yu2018; Miwa2014]. Moreover, SVMs are the currently the only classifier implemented in ready-to-use software tools implementing active learning for screening prioritization (Abstrackr [@Wallace2012b], Colandr [@Cheng2018], FASTREAD [@Yu2018], Rayyan [@Ouzzani2016], and RobotAnalyst [@Przybyla2018]).

Whilst performance several classification techniques have been investigated in the AI-aided title and abstract screening field in general [SOURCE], the relatively new subfield of active learning for screening prioritization has not yet studied the performance of classifiers other than SVMs [@Yu2018; @Yu2018; @Cormack2014; @Cormack2015; @Miwa2014; @Wallace2010]. 
<!-- Whilst the SVM is a well-known technique, also classifiers have been employed in class prediction tasks in systematic reviews.  -->
<!-- In fact, research in this area to date has not yet determined the performance of classifiers other than SVMs. -->
The current study aims to address this gap by exploring performance of three classifiers besides SVM:

 - L2-regularized Logistic Regression (LR)  models the probabilities describing the possible outcomes by a logistic function. The L2 penalty is imposed on the coefficients to reduce the number of features upon which the given solution is dependent [@scikit-learn].
 - Naive Bayes (NB) is a supervised learning algorithm often used in text classification. Based on Bayes' Theorem, with the 'naive' assumption that all features are independent given the class value [@Zhang2004].
- Random Forests (RF) is a supervised learning algorithm where a large number of decision trees are fit on bootstrapped samples of the original data. All trees cast a vote on the class, which are aggregated into a class prediction for each instance [@Breiman2001].

These three classification techniques were selected because they are all widely adopted in previous studies on text classifcation [source]. Moreover, the amount of processing power is relatively low. They can run on a personal computer.

 <!-- Logistic models the posterior directly, naive bayes has higher bias but lower variance,  -->
 <!-- logistic might perform better when trainign set increases. [@Ng2002] -->
<!-- To carry out classification of publications, a classifier takes given input features $\mathbf{X}$ -->
<!-- The model utilizes features of previously labeled publications and  -->
<!-- The model utilizes L to predict labels for all abstracts (classifier). (training) -->
<!-- a new paper is selected (query) the one that is most likely to be relevant. & queried  -->
<!-- The reviewer is asked to label the query (relevant, irrelevant), which is then added to the training set.  -->
<!-- Now the model can update in (retraining) its predictions based on the new training dataset. This is caled retraining.  -->
<!-- The cycle starts over again..  -->
<!-- As the cycle repeats itselfs, labeled dataset grows, predicitons improve hopefully and soon (or sooner than at random) the relevant publications will we identified. s -->
<!-- We want to classify based on some features from the instances $x$, feature vector $\mathbf{X}$.  -->

#### Class imbalance problem
There are two classes in the dataset: relevant and irrelevant publications. Typically, only a fraction of the publications in the data belong to the relevant class. This poses a problem for training a classifier as there are far fewer examples of relevant than irrelevant publications to train on [@OMara-Eves2015]. Moreover, classifiers are well-suited to separate data into classes, but not to correctly identifying one class [@Wallace2010]. This is evident in the case of a dataset where only one percent of publications are relevant. A model would achieve 99% accuracy when classifying all publications as irrelevant, even though none of the relevant papers would have been identified. Therefore, the class imbalance problem causes the classifier to miss relevant publications.

Previous studies have addressed the class imbalance problem by rebalancing the training data in different ways [@OMara-Eves2015]. To decrease the class imbalance in the training data, the models in the current study rebalance the training set by Dynamic Supersampling (DS). DS decreases the number of irrelevant publications in the training data, whereas the number of relevant publications are increased (by copy) such that the size of the training data remains the same. The ratio between relevant and irrelevant publications is not fixed, but dynamically updated and depends on the size of the training data, the total number of publications, and the ratio between total number of relevant and irrelevant publications.

#### Word embeddings 
To predict publication class, the classifier uses information from the publications in the dataset. 
Examples of such information are titles and abstracts. However, a model cannot predict class from the titles and abstracts as they are; their textual content needs to be repressented numerically. The textual information needs to be mapped to feature vectors. This process of numerically representing textual content is called 'word embeddings'.

A classical example of word embeddings is a 'bag of words' (bow) representation. For each text in the data set, the number of occurrences of each word is stored. This leads to $n$ features, where $n$ is the number of distinct words in the texts [@scikit-learn]. The bag-of-words method is simplistic and will highly value often occuring but otherwise meaningless words such as "and". A more sophisticated approach is Term-frequency Inverse Document Frequency (TF-IDF). TF-IDF circumvents this problem by adjusting the term frequency in a text with the inverse docuement frequency, the frequency of a given word in the entire data set [@Ramos2003]. A downside of TF-IDF and other bow methods is that they do not take into account the ordering of the words, thereby ignoring semantics. An example of an approach that aims to overcome this weakness is Doc2Vec (D2V), capable of grasping the relations between words by learning to predict the words in the texts [@Le2014].

Miwa et al. found that active learning was more difficult on data from the social sciences compared to data the medical sciences and were able to link this difficulty to a natural difference in text complexity between these research areas [@Miwa2014]. As the study by Miwa et al. adopted a bow approach [@Miwa2014], we hypothesize that a more sophisticated word embeddings strategy has the potential to bridge the performance gap between these research areas.

#### Query strategy
The active learning model can adopt different strategies in selecting the next publication to be screened by the reviewer. 
A strategy mentioned before is selecting the publication with the highest probability of being relevant. In the active learning literature this is referred to as certainty-based active learning [@Settles2012]. Another well-known strategy is uncertainty-based active learning, where the instances that will be presented next will be those instances on which on which the model's classifications are the least certain, i.e. close to 0.5 probability [@Settles2012]. Traditionally, this strategy trains the most accurate model because the model can learn the most from instances it is uncertain about.

Even though uncertainty-based active learning generally leads to a more accurate model in the end, certainty-based active learning is the preferred strategy for the task at hand. The main reason is that this strategy is far better suited to the goal of prioritizing the relevant publications. Moreover, a study comparing performance of both strategies in found that the accuracy gain of uncertainty-based screening was not significant [@Miwa2014]. Furthermore, uncertainty-based active learning is far better equipped at dealing with imbalanced data in active learning [@Fu2011]. 
Therefore, certainty-based sampling is most equipped to the current scenario in which we are dealing with highly imbalanced datasets and where the goal is to identify all relevant publications as soon as possible.

## Models 
The seven models consist of the components described in the Technical Details section, adopting four different classification techniques and two different word embeddings approaches:

First, four models combining every classifier with TF-IDF word embeddings were investigated: 
 - SVM + TF-IDF
 - NB + TF-IDF
 - RF + TF-IDF 
 - SVM + TF-IDF

Second, the classifiers were combined with Doc2Vec word embeddings, leading to the following three models:^[The combination NB + D2V is impossible because NB can only process positive input features, whereas D2V]

 - SVM + Doc2Vec
 - RF + Doc2Vec
 - SVM + Doc2Vec

## Simulation study
For each of the seven models, performance was evaluated by simulating the model on the screening process of six systematic reviews. 
Performance of the seven models was evaluated by simulating their behaviour in the screening process of six systematic review datasets.
Put differently, 42 simulations were carried out. For every model - dataset combination, hyperparameters were optimized^[see the Appendix for more information]. To account for variance, every simulation was repeated for 15 trials. 
Simulations were run using ASReview's simulation mode [@ASReview2020]. There was no need for a human reviewer as the model could query the labels in the data instead.

Every simulation started with an initial training set of one relevant and one irrelevant publication to represent a 'worst case scenario' where the reviewer has minimal prior knowledge on the publications in the data. To account for bias, the initial training set was randomly sampled from the dataset for every of the 15 runs. Although varying over runs, the initial training sets were kept constant over datasets to allow for a direct comparison of models within datasets. A seed value was set to ensure reproducability. 
The classifier was retrained every time after a publication had been labeled. 
The simulation ended after all publications in the dataset had been labeled.

<!-- can retrieve the labels from the annotated dataset instead -->
<!-- The oracle is not the reviewer but the labels in the data.  -->
<!-- The researcher has some prior knowledge about the pool, some papers ought to be included in the SR.  -->

## Datasets 
The models were simulated on a convenience sample of six systematic review datasets. The data selection process was driven by two factors. Firstly, datasets were selected based on their background, given the need for datasets from diverse research areas. Secondly, datasets were selected by their availability, given the limited timespan of the current project. The datasets were retrieved from a collection of open systematic review datasets to be used for text mining purposes^[https://github.com/asreview/systematic-review-datasets, [@ASReview2020]].

Three out of six datasets originated from the _medical sciences_: Ace, Wilson, and Virus.
The Wilson dataset [@Appenzeller-Herzog2020] is on a review on effectiveness and safety of treatments of Wilson Disease, a rare genetic disorder of copper metabolism [@Appenzeller-Herzog2019]. The Ace dataset contains publications on the efficacy of Angiotensin-converting enzyme (ACE) inhibitors, a drug treatment for heart disease [@Cohen2006]. The Virus dataset is from a systematic review on studies that performed viral Metagenomic Next-Generation Sequencing (mNGS) in farm animals [@Kwok2020].
From the field of _software engineering_, the Software dataset contains publications from a review on fault prediction in source code [@Hall2012]. The Nudging dataset [@Nagtegaal2019a] belongs to a systematic review on nudging healthcare professionals [@Nagtegaal2019], stemming from the area of _behavioural public administration_, 
The PTSD dataset contains publications from the field of _psychology_. The corresponding systematic review is on studies applying latent trajectory analyses on posttraumatic stress after exposure to trauma [@vandeSchoot2017]. Of these six datasets, Ace, and Software have been used for model simulations in previous studies on AI-aided title and abstract screening, respectively [@Cohen2006] and [@Yu2018].

Data were preprocessed from their original source into a test dataset, containing title and abstract of the publications obtained in the initial search. Candidate studies with missing abstracts and duplicate instances were removed from the data. Test datasets were labelled to indicate which candidate studies were included in the systematic review, thereby indicating relevant publications. All test datasets consisted of thousands of candidate studies, of which only only a fraction was deemed relevant to the systematic review. Inclusion rates were centered around 1-2 percent with one outlier of about 5 percent (`r table_nums("datasets", display = "cite")`). 
<!-- The data preprocessing scripts can be found on the GitHub^[https://github.com/GerbrichFerdinands/asreview-thesis] -->

```{r}
# datasets table
drops <- readRDS(paste0(pathRdata, "drops.RDS"))
all <- readRDS(paste0(pathRdata, "all.RDS"))

nstudies <- function(all, datastage, col){
}

nostudies <- function(all, set, stage){
  sapply(all, function(x) x[set, stage])
}
inclrate <- function(all, set){
  sapply(all, function(x) round(x[set,"incl"]/x[set,"search"]*100,2))
}
datastats <- 
  tibble(Dataset = names(all), 
       #Citation = NA, # maybe add footnote citation with kableExtra i.o. this.
       # paper 
       `candidates_paper` = nostudies(all, "paper", "search"), 
       #`fulltext_paper` = nostudies(all, "paper", "ftext"),
       `incl_paper` = nostudies(all, "paper", "incl"),
       `inclrate_paper` = inclrate(all, "paper"),
       # test set 
       `candidates_test` =  nostudies(all, "test", "search"), 
       #`fulltext_test` =  nostudies(all, "test", "ftext"), 
       `incl_test` =  nostudies(all, "test", "incl"), 
       `inclrate_test` = inclrate(all, "test")
       ) 
```

`r table_nums("datasets")`
```{r}
datastats %>%
  kable(format = "latex",
        booktabs = TRUE,
        col.names = c("Dataset", rep(c("Candidate studies", 
                                       #"Studies selected for fulltext screening", 
                                       "Relevant studies", 
                                       "Inclusion rate (%)"),2))) %>%
  kable_styling(full_width = TRUE) %>%
  add_header_above(c("", "Original study" = 3, "Test collection" = 3))
```

## Evaluating performance
Model performance was visualized by plotting recall curves and further assessed by three different measures, Work Saved over Sampling (WSS), Relevant References Found (RRF), and Average Time to Discovery (ATD). Results were averaged over 15 trials for every simulation.

Plotting recall as a function of the number of screened publications offers insight in model performance throughout the screening process [@Cormack2014; @Yu2018]. The curves give informations in two directions. On the one hand they display the number of publications that need to be screened to achieve a certain level of recall (WSS), but on the other hand they present how many relevant publications are identified after screening a certain proportion of all publications (RRF).

WSS indicates the reduction in publications needed to be screened, at a given level of recall [@Cohen2006]. 
When measured at the typical recall level of 0.95 [@Cohen2006], WSS yields an estimate of the amount of work that can have be saved at the cost of failing to identify 5% of relevant publications. In the current study, WSS is computed at 0.95 and 1.00 recall level. 
RRF statistics are computed at 10%, representing the proportion of relevant studies that were identified after screening 10% of all publications.

Both RRF and WSS are sensitive to random effects as these statistics are strongly dependent on the position of the cutoff value. 
Moreover, WSS makes assumptions about acceptable recall levels whereas this might depend on the research question at hand [@OMara-Eves2015]. 
A statistic that is not dependent on some arbitrary cutoff value is the ATD, which is the average number of publications that are screened to find a relevant publication, divided by the total number of publications in the data. The ATD is proportional to the area above the recall curve.

<!-- FN are far more problematic/unwanted/detrimental than FP: a researcher can label the FP as irrelevant whereas the FN are possibly never seen thereby missing relevant studies running the risk of an invalid systematic review.  -->
<!-- The goal is twofold: we want to identify all relevant papers, as fast as we can.  -->
<!-- Tradeoff: identifying all relevant papers and reducing workload.  -->
<!-- what about class imbalance.  -->

Analyses were carried out using R [@RCoreTeam2019], version 3.6.1. 
All datasets accompanying the systematic reviews are openly published. 
This study was approved by the Ethics Committee of the Faculty of Social and Behavioural Sciences of Utrecht University, filed as an amendement under study 20-104.
All simulations were run using through Cartesius (EINF-156).

<!-- $$\texttt{WSS} = \frac{TN + FN}{N} - (1- recall) $$  -->

<!-- $$\texttt{ATD} = \frac{L}{N}$$ -->
<!-- $$\texttt{ATD} = \frac{L(relevant) }{N}$$ -->

<!-- Is performance related to some characteristic (n, inclusion rate, ...) -->

<!-- ? How to compare outcomes of 3 different optimization strategies?  -->

<!-- The cost of a false negative outweighs the cost of a false positive.  -->
<!-- Note that we assume the oracle/original user to hold the truth.  -->
<!-- This is of course not always the case.  -->
<!-- There are two classes in the data: exlusions and inclusions. -->

<!-- Datasets from the medical and social sciences, software engineering and public administration.  -->
<!-- Medical sciences SRs are viewed as more 'strict'/'structured' and social sciences more messy.  -->
<!-- Models are evaluated by performing a simulation on data from six existing systematic reviews from various research areas. -->

---

### Optimizing hyperparameters
Every model component contains hyperparameters, leading to a unique set of hyperparameters for each model. To maximize model performance, we need to find optimal values for the hyperparameters. 
For every model the optimal hyperparameter values are determined by optimizing on the data $d$.
The hyperparameters are optimized by running several hundreds of optimization trials, in which hyperparameter values are sampled from their possible parameter space.
A description of all hyperparameters and their sample space can be found in the appendix.

Maximum model performance is defined as the average time it takes to find an inclusion in the data, or more specific: the loss function minimizes the average number of papers needed to screen to find an inclusion (e.g. the area above the curve in the inclusion plot).

The optimization data $d$ consists of (a subset from) the six systematic review datasets $D$ mentioned above. 
Three different approaches in composing $d$ are explored:

- __one__, where hyperparameters are optimized on only one of the six datasets, $d \in D$. Such hyperparameters are expected to lead to maximum performance in the same dataset $d$.
- __n__, where hyperparameters are optimized on all six data sets, $d = D$. This optimization approach intends to serve in producing the most optimal hyperparameters overall. 
- __n-1__, where hyperparameters are optimized on all six datasets but one, $d \subset D$. Serving as a sensitivity analysis for the former condition, e.g. how sensitive are the hyperparamters. also as a cross-validation later on: hyperparameters obtained by training data, test data is never seen before.
where d are all datasets but the one where we want to simulate later on. 
This results in $6+6+1=13$ sets of hyperparameters for every model.

Results were visually inspected to check if an optimum (minimal loss) has been reached. 
More trials were run if the loss still seemed to go down at a quick pace.

The hyperparameter values that were found to lead to a minimum loss value were visually inspected.

\newpage
## Background
[@Yu2018], [@Yu2018] simulated 32 svm classifiers, on software engineering. 
A popular classifier is SVM. succes with HUTM (fastread), uncertainty, mix of weighting and agressive undersampling, 
In terms of Yu et al, we adopt .CT.

SVM - tf-idf on medical data, uncertainty sampling, agressive undersampling. [@Wallace2010]

abstrackr svm certainty

SVM + Weighting + uncertainty (bow) produced good methods  [@Miwa2014]
Also include social sciences data besides medical data.

[@Cohen2006] perceptron-based classifier (neural network)

SVM on legal documents (no balancing, certainty ) [@Cormack2014] 
in limitations section mentions that LR yields about same results, nb inferior results.

[@Kilicoglu2009] - SVM, naive bayes, boosting and combinations. future work should optimize parameters.
  "Regarding the base classifiers used in identifying method- ologically rigorous studies, boosting consistently strikes the best balance between precision and recall, whereas naive Bayes in general performs well on recall (demonstrating a tradeoff between recall and precision), as does polynomial SVM on precision. The AUC results are mixed, although boosting has a slight edge overall. These results demonstrate that different classifiers can be used to satisfy different information needs (SVM for specificity, naive Bayes for sensitivity, and boosting for balance between the two, for example)."

Our extensions is that we try different classifiers, on more datasets.

When no balancing is applied, the training data set = labeled data set $\mathcal{L}$
The model can query the labels, who serve as the reviewer, active learning then perform active learning to detect inclusions.

a machine learning-based citation classification tool to reduce workload in systematic reviews of drug class efficacy. 
Using a perceptron classifier, WSS@95% = 56.61 in [@Cohen2006]. (5x2 crossvalidation). Can we beat this? 
The data

Openness, reproducible,

Benefits:

Adopting some sort of stopping criterion (outside scope of the current thesis) the reviewer can quit reviewing after having read only a fraction of candidate studies. 
meaning the screening process can be finished after reading a fraction of all candidate studies. Saving hours of time and resources.

paper organization:

The goal is to gain insight in classifiers other than the widely applied SVM, overall various research areas. 
So not only medical sciences.

Although considerable research has been devoted to ..., less attention has been paid to the comparison of different classifiers. 
Few studies have evaluated different classifiers in any systematic way
different models to use haven't been investigated in a systematic way (only in software engineering)

1) the lack of replication of methods is making it impossible to draw any overall conclusions about best practice/ best approaches of the problem of reducing workload in screening. 
2) screening prioritization is appealing to systematic reviewers because ...
3) 
4) implemantation: usabliity and acceptability of such tools amongst researchers conducting a systematic review.

lack of studies investigating the effect of different methods over different research areas. (yu but only software engineering, miwa perhaps?)
There is also a significant lack of examples outside of healthcare with the exception of one example in software engingeering

However, there is a need for consensus () and transparancy?

So far, however, there has been little discussion about the different classifiers to be used. Support Vector Machine has been the default in almost all studies.

Several algorithms to assist the reviewer in the abstract screening process have been proposed. (they are available in many shapes/sorts).

<!-- RQ2 different hyperparameter optimizations? -->
<!-- Classifiers come with hyperparameters. We have to make a choice on how to set these hyperparameters. What we do is create 3 sets, optimizing in three ways, aggregate results obtained.   -->

For truly evaluate the effectiveness of such methods,

Such a solution can save time and resources. Time saving is not the only benefit: and help to minimize bias..

#### Proposed solution
This study is about how machine learning models can increase efficiency in systematic reviews.
Their main objective is to identify ..

The strategy to reduce workload proposed in the current study is by prioritizing publications that are deemed most relevant to the systematic review. 
As the relevant publications are screened first, the reviewing process can be quit earlier.

The stage of abstract screening where abstracts are systematically screened is where a lot is to be gained. This stage is the target of possible learning algorithms that can assist the reviewer in selecting the relevant papers. Together with the reviewer /human machine interaction.  The algorithm aims to compute which papers in the pool need to be excluded and which need to be included, based on the reviewers decisions. It learns from the reviewers decisions and asks the reviewer to provide more labels, incrementally improving its class predictions.

The goal of the algorithm defined in the current study is to reduce to number of abstracts needed to screen (maybe not right term, bit biomedical). To be more specific, the algorithm aims to present the reader with the primary studies as soon as possible. This means that at some point you probably have seen all relevant abstracts and are only viewing excluded papers, which means you can stop reviewing much earlier (theoretically spoken). Also reviewing is now much more fun. As compared to when you have to review all abstracts and you perhaps see only one relevant abstract every other week/day.

### Background
... starts with a search for potentially relevant publications. 
This initial set of candidate studies need to be manually screened to identify the publications relevant for answering the research question. 
Because the initial set often consists of thousands of papers and the

To gather the findings relevant to answering the research question, a systematic search is performed.
A systematic search starts with collecting all publications that meet pre-specified eligibility criteria.
From this collection of candidate studies the researcher has to identify the publications relevant for answering the research question. 
Of all candidate studies only a fraction is relevant [...]. 
As more and more papers are published and reproducibility crisis has emerged, 
A systematic search often results in thousands of candidate studies.
Relevant publications are then identified by screening title and abstract of all candidate studies. 
This screening process is a manual task often executed by multiple reviewers to ensure reliability. 
This is a time consuming process that weighs heavily on resources.

Most often the SVM classifier is used, popular and very good results. Also lots of other configurations. However, other classifiers have not been tested a lot (polygon thing by cohe, naïve bayes and random forest by …), but mostly SVM still. Also, most research in the medical sciences (well there are some exceptions of course [conversation between cohen and matwill]

A solution is ...

It is important reflect on research by giving an overview of research areas which is typically done by a systematic review […].

To review a specific research area, one starts out with an initial search of thousands of academic papers. All these papers abstracts need to be screened to find an initial batch of possibly relevant papers. With now hopefully only a couple of hundred papers left, the researcher needs to read these papers full-text to arrive at a final selection of papers that are relevant for the final systematic review [this is prisma process?]. This whole processes costs this and this much time [shelmilt].

----

## active learning for systematic reviews 
corpus = all the text:

Active learning = 
increasing classification performance with every query. 
The query strategy determines the way unlabeled papers are queried to the researcher.

[@modAL2018]

\newpage

### Optimization results
For every model, 13 sets of hyperparameters were optimized.

*to include: plotting the loss reduction over trials*

Optimal values of the hyperparameters were visually inspected.

As an example, the hyperparameters of the RF_TF-IDF model are presented in Figure 1. 
A panel displays the optimal values for a certain hyperparameter, where the blue colored dots represent the __one__ condition, the green dots the __n-1__ condition and the orange dot represents the optimal hyperparameter value when optmizing over all datasets (__n__). The x-axis represents the possible parameter space where the vertical grey lines mark the boundaries of the hyperparameter space (if possible). Note that the feature parameters ngram_max, split_ta, use_keywords and model parameters max_features and n_estimators are categorical.

Overall, the optimal values are distributed over the parameter space. Outlying values all belong to the __one__ condition where optimization was dependent on one dataset only. Fore example, the class_weight parameter has an outlying value of 11.3 that belongs to the nudging dataset. (Why this is the case I can only speculate, but it is worth mentioning that this dataset has a relatively high inclusion rate  of 5.41%, compared to the other datasets).

```{r}
include_graphics(
  "figs/rf_tfidf.eps",
)
```

## Appendix x - Hyperparameters and their sample space

Classifier hyperparameters

    class_weight: normal(0,1) constrained to be > 0
    class weight of the inclusions.

  Logistic Regression

    c: float. normal(0,1), constrained to be > 0

  Support Vector Machine

    gamma: ["auto", "scale"],
    Gamma parameter of the SVM model.

    C: 
    C parameter of the SVM model.

    kernel:
    SVM kernel type.["linear", "rbf", "poly", "sigmoid"]

  Naive Bayes

    "model_param.alpha", # exp(normal(0,1))

  Random Forest

    max_features: int (between 6 and 10)
        Number of features in the model.

    n_estimators: int between 10 and 100 
        Number of estimators.

    model_param.n_estimators" #quniform(10,100,1)

Balance strategy hyperparameters:
    Dynamic supersampling:

    a: float
        Governs the weight of the 1's. Higher values mean linearly more 1's
        in your training sample.
    alpha: float
        Governs the scaling the weight of the 1's, as a function of the
        ratio of ones to zeros. A positive value means that the lower the
        ratio of zeros to ones, the higher the weight of the ones.
    b: float
        Governs how strongly we want to sample depending on the total
        number of samples. A value of 1 means no dependence on the total
        number of samples, while lower values mean increasingly stronger
        dependence on the number of samples.

Feature extraction strategy hyperparameters

    split_ta: 0 or 1
       whether titles and abstracts are split

    use_keywords: 0 or 1
        whether keywords should be used

  TF-IDF

    ngram_max: 1, 2 or 3
        Can use up to ngrams up to ngram_max. For example in the case of 
        ngram_max=2, monograms and bigrams could be used.

  Doc2Vec

    vector_size: int (between 32 and 127)
        Output size of the vector.

    epochs: int (between 20 and 50)
        Number of epochs to train the doc2vec model.

    min_count: int (between 1 and 3)
        Minimum number of occurences for a word in the corpus for it to be 
        included in the model.

    window: int (between 5 and 9)
        Maximum distance over which word vectors influence each other.

    dm_concat: int 0 or 1
        Whether to concatenate word vectors or not.

    dm: int 
        Model to use.
        0: Use distribute bag of words (DBOW).
        1: Use distributed memory (DM).
        2: Use both of the above with half the vector size and concatenate them.

    dbow_words: int 0 or 1
        Whether to train the word vectors using the skipgram method.

# The software 
ASReview takes the following parameters/arguments: 
 We now have 75 combinations. 
for every for every model (5), for every dataset (5) and for every set of optimized hyperparameters (3), a simulation study consisting trials is performed. From these $5*5*3=75$ simulation studies, performance of the different models is evaluated.

```{r}
tab %>%
kable(format = "latex", 
      booktabs = TRUE, 
      col.names = "Configurations") %>%
  kable_styling(full_width = TRUE)
```

Use these inputs to predict relevance of papers.

### Stage 1: hyperparameter optimization

Or, more specific:

```{r}
tabbie <-
  combins %>%
  filter(query_strategy == "max",
         feature_extraction %in% c("doc2vec", "tfidf") & model %in% c("dense_nn","svm", "nb", "rf", "lr")) %>%
  select(-query_strategy)

tabbie <- tabbie[c(1,3,5,6,9),]
row.names(tabbie) <- NULL

tabbie %>%
  kable(col.names = names(words)[c(1,3)], 
        format = "latex",
        booktabs = TRUE)
```

\newpage

# Results

```{r}
simulations <- readRDS("../../simulation_study/R/01_hyperparameter_sets.RDS")[,,1]

kable(simulations)
```

\newpage
# Discussion

- we look for final inclusions but we screen only the abstracts (do they satisfy the imformation need  (blake (page 19 omara et evs)))

future research: 
- stopping rule is not discussed
- computation/retraining time

strengths:

- open data
- different research areas
- different models on same dataset
- different datasets on same model

limitations

future research
- all models save time, difficult to distinguish performance over datasets, especially when applied on a dataset of which no prior information is known (e.g. inclusions isn't known in practice). Perhaps go for other criteria like the fastest model, replicate study with computation time?

Simulating the title and abstract screening process, models are evaluted on their capability/speed of detecting the final inclusions. 
However, in a manual SR these final inclusions are selected after reading the fulltext. Information the text mining tool does not have. 
To truly assess the added value of such a tool, models should be evaluated on their capability of detecting the abstract inclusions. 
Call for systematic reviewers to openly publish need for open data containing abstract inclusions, not only final inclusions!

\newpage
# Appendix A - list of definitions

### Feature Extraction Strategies 
split_ta = overall hyperparameter

#### TF-IDF

##### hyperparameters 
    ngram_max: int
            Can use up to ngrams up to ngram_max. For example in the case of
            ngram_max=2, monograms and bigrams could be used.

#### Doc2Vec
Predicts words from context. 
Aims at capturing the relations between word (man-woman, king-queen).
[@Le2014]. Using a neural network.

using Continuous Bag-of-Words (CBOW), Skip-Gram model, ....
Word vector _W_ and extra: document vector _D_, trained to predict words in the text.

From gensim [@Rehurek2010].

        Arguments
        ---------
        vector_size: int
            Output size of the vector.
        epochs: int
            Number of epochs to train the doc2vec model.
        min_count: int
            Minimum number of occurences for a word in the corpus for it to
            be included in the model.
        workers: int
            Number of threads to train the model with.
        window: int
            Maximum distance over which word vectors influence each other.
        dm_concat: int
            Whether to concatenate word vectors or not.
            See paper for more detail.
        dm: int
            Model to use.
            0: Use distribute bag of words (DBOW).
            1: Use distributed memory (DM).
            2: Use both of the above with half the vector size and concatenate
            them.
        dbow_words: int
            Whether to train the word vectors using the skipgram metho

#### SBERT

BERT-base model with mean-tokens pooling [@Reimers2019]

#### embeddingIdf 
This model averages the weighted word vectors of all the words in the text,
in order to get a single feature vector for each text. The weights are provided by the inverse document frequencies

### Models 
#### Naive Bayes 
Naive Bayes assumes all features are independent given the class value. [@Zhang2004]

ASReview uses the ` MultinomialNB` from the scikit-learn package [@scikit-learn], that implements the naive Bayes algorithm for multinomially distributed data. 
` nb`

Hyperparameters

  * alpha - accounts for features not present in learning samples and prevents zero probabilities in further computations.

#### Random Forests
A number of decision trees are fit on bootstrapped samples of the original data, 
[@Breiman2001]
RandomForestClassifier from sklearn

Arguments
        ---------
        n_estimators: int
            Number of estimators.
        max_features: int
            Number of features in the model.
        class_weight: float
            Class weight of the inclusions.
        random_state: int, RandomState
            Set the random state of the RNG.
        """

#### Support Vector Machine

 Arguments
        ---------
        gamma: str
            Gamma parameter of the SVM model.
        class_weight:
            class_weight of the inclusions.
        C:
            C parameter of the SVM model.
        kernel:
            SVM kernel type.
        random_state:
            State of the RNG.

#### Logistic Regression

#### __Dense Neural Network__

### Query Strategies 
* Max - Choose the most likely samples to be included according to the model
* Uncertainty - choose the most uncertain samples according to the model (i.e. closest to 0.5 probability) [@Lewis1994] 
* Random - randomly selects abstracts with no regard to model assigned probabilities. 
* Cluster - Use clustering after feature extraction on the dataset. Then the highest probabilities within random clusters are sampled

The following combinations are simulated:

* cluster
* max
* cluster * random 
* cluster * uncertainty
* max * cluster
* max * random
* max * uncertainty

### Balance Strategies

### amount of training data

* n_instances = number of papers queried each query
* n_queries = number of queries
* n_prior_included: 5
* n_prior_excluded:

# Combinations

This leads to  `r nrow(combins)` combinations of configurations.

- Naive bayes only goes with tfidf feature extraction.
- For the feature extraction strategies we will focus on doc2vec and tfidf. (but will compute all 4)
- This leads to 3 * 7  * 4 * 3 + 1 * 7 * 1 * 3 =  273 combinations.

See appendix A for a table containing all 273 combinations.

## Cross-validation
Should give an accurate estimate of maximum performance / future systematic reviews to be performed.

# Appendix B - combinations

```{r}
names(combins) <- c("Model", "Query Strategy", "Feature extraction strategy") #, "Training data [included/excluded]")
combins %>%
  kable(row.names = FALSE, format = "latex", longtable = TRUE, booktabs = TRUE) %>%
  kable_styling(latex_options = c("repeat_header")) 
```

# Appendix C - supercomputer Cartesius

500,000 SBU

Running on Cartesius is charged in System Billing Units (SBUs), and charging is based on the wall clock time of a job. On fat and thin nodes, an SBU is equal to using 1 core for 1 hour (a core hour), or 1 core for 20 minutes on a GPU node. Since compute nodes are allocated exclusively to a single job at a time, you will be charged for all cores on that node - even if you are using less.

In the current study, the classifier and the feature extraction strategy are varied, whereas the query and balance strategy remain fixed. 
In the current study only a fraction of all possible configurations are tested for the sake of brevity. 
There are many more options available and open to exploration.

# References