### 2020-05-04 Untitled/Untitled.Rmd
updates intro + mmethods + results

---
title: Title here

author: 
- name: Author 1
  affiliation: Department of YYY, University of XXX
  email: abc@def

- name: Author 2
  affiliation: Department of ZZZ, University of WWW
  email: djf@wef

keywords:
- key
- dictionary
- word

abstract: |
  The text of your summary.  Should not exceed 225 words.

bibliography: bibliography.bib
output: rticles::arxiv_article
month: "`r format(Sys.Date(), '%b')`"
year: "`r format(Sys.Date(), '%Y')`"
referee: true # Papers submitted to Biometrics should ALWAYS be prepared
              # using the referee option!!!! Turn off only to preview
              # two column-format
---

# Introduction {#intro}

Your text comes here. Separate text sections with

# Section title {#sec:1}

Text with citations by @heagerty2000time, [@pepe2003statistical].

## Subsection title {#sec:2}

as required [@hoerl1970ridge; @zou2005regularization]. Don't forget to give each section
and subsection a unique label (see Sect. \ref{sec:1}).

#### Paragraph headings

Use paragraph headings as needed.

## Equations

Here is an equation:

$$ f_{X}(x) = \left(\frac{\alpha}{\beta}\right)\left(\frac{x}{\beta}\right)^{\alpha-1}e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; \alpha,\beta,x > 0 $$

Here is another:
\begin{align}
a^2+b^2=c^2
\end{align}

Inline equations: $\sum_{i = 2}^\infty\{\alpha_i^\beta\}$

# Figures and tables

## Figures coming from R

#### Normal figure embedded in text

```{r fig2, fig.width = 4, fig.height = 4, fig.cap = "Output from `pdf()`", echo = FALSE, dev = "pdf"}
plot(runif(25) ~ runif(25))
```

\clearpage

## Tables coming from R

```{r tab1, results = 'asis'}
print(xtable::xtable(head(mtcars)[,1:4], 
caption = "Caption centered under table", label = "tab1"), 
comment = FALSE, timestamp = FALSE, caption.placement = "top")
```

Table \ref{tab1} shows these numbers. Some of those numbers are plotted in Figure \ref{fig:fig1}.

```{r tab2}
head(mtcars[,1:4])
```

# References