### 2020-05-11 results/README.Rmd

---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(rjson)
library(tidyverse)
library(knitr)
library(flextable)
library(officer)
```

Script for reading simulation files
```{python}
# to be found on harddrive results folder
```

Script for reading results 
```{r}
data <- readRDS("../simulation_study/R/00_datasets.RDS")
models <- list.files("/Volumes/Gerbrich/asreview/results/one_seed/statistics/")
names(models) <- c("NB + TF-IDF", "LR + D2V", "LR + TF-IDF", "RF + D2V", "RF + TF-IDF", "SVM + D2V", "SVM + TF-IDF" )

# function that reads results for a model M 
read_results <- function(m){
   files = list.files(paste0("/Volumes/Gerbrich/asreview/results/one_seed/statistics/", m), recursive = TRUE)
  # names of the files are the data
  names(files) <- str_split(files, "/", simplify = TRUE)[,1]
  # read data
  dat <- lapply(files, function(x) fromJSON(file = paste0("/Volumes/Gerbrich/asreview/results/one_seed/statistics/", m, "/", x)))

  # extract wss, rrf, and loss
  dat <- map(dat, `[`, c("wss", "rrf", "loss"))

  # transorm into dataframe
  dat <- map_dfr(dat, ~ as.data.frame(.x), .id = "dataset")

  # add model name
  dat <- dat %>% 
    mutate(model = names(models[models == m]))

  return(dat)
}
```

Parse 
```{r, eval = FALSE}
dat$virus
# extract only wss, rrf, and loss values from list
test <- map(dat, `[`, c("wss", "rrf", "loss"))
# transorm into dataframe
test <- map_dfr(test, ~ as.data.frame(.x), .id = "dataset")
# add model name
test <- test %>% mutate(model = "NB + TF-IDF")
```

Apply over all datasets and models 
```{r}
# extract results for all models
# list for models separately
results <- lapply(models, read_results)

# all in one dataframe
results <- do.call("rbind", results)

# convert loss (ttd) to percentage 
results$loss <- results$loss*100
# save results file 
saveRDS(results, "output/results.RDS")

results %>%
  group_by(dataset) %>%
  arrange(loss)
```

Create table
```{r}

tabres <- 
  results %>%
  pivot_wider(names_from = dataset, values_from = c("wss.95", "wss.99", "wss.100", "rrf.5", "rrf.10", "rrf.20", "loss"))

```

```{r}

test <- results %>% pivot_longer(cols = c(starts_with("wss"), starts_with("rrf"), "loss"), names_to = "stats", values_to = "value")

tab <- 
  test %>%
  tab_rows(model) %>%
  tab_cells(value) %>%
  tab_cols(dataset %nest% stats) %>%
  tab_stat_mean() %>%
  tab_pivot()

```

```{r}
tabres %>%
  select(model, starts_with("wss.95"), starts_with("loss"))

gtest <- tabres %>%
  gt()

gtest <- results %>%
  gt()

```
```{r}

metricsat <- c("WSS@95", "WSS@99", "WSS@100", "RRF@5", "RRF@10", "RRF@20", "ATD")
metrics <- c("WSS", "RRF", "TTD")
at <- c("95", "99", "100", "5", "10", "20", "0")

typology <- data.frame(
  col_keys = names(tabres)[2:43],
  what = rep(data,7),
  measure = rep(metricsat, each = 6),
  what = c(rep("WSS", 18), rep("RRF", 18), rep("ttd", 6)),
  measure = c(rep(c("@95", "@99", "@100"), each = 6), rep(c("@5", "@10", "@20"), each = 6), rep("", 6)),
  stringsAsFactors = FALSE
)

test <- flextable(tabres)
test <- set_header_df(test, mapping = typology, key = "col_keys")
test <- merge_h(test, part = "header")

test <- merge_v(test, part = "header")

test <- theme_booktabs(test)
test <- autofit(test)
test <- fix_border_issues(test)
```

```{r}
# reorder columns in tabres
# which of the vars do we need?

# select_and_shuffle <- function(metrics, tab){
#   # metrics we got 
#   metricsat <- c("WSS@95", "WSS@99", "WSS@100", "RRF@5", "RRF@10", "RRF@20", "ATD")
#   # which metrics we want 
#   selector <- which(metricsat %in% metrics)
#   selector <- selector + 1
#   # sequence to reorder 
#   print(selector)
#   selector <- matrix(selector)
#   seqq <- apply(selector, 1, function(x) seq(x, 43, 6))
#   
#   seqq <- do.call("c", seqq)
#   print(seqq)
#   # reorder the selected columns
#   tab <- tab[, c(1, seqq)]
#   
#   return(tab)
#   
# }
# 
# tabres[, c(1, c(apply(matrix(2:7), 1, function(x) seq(x, 43, 6))))]
# 
# metri <- metricsat[c(1,5,7)]
```

```{r}
stabres <- 
tabres %>%
  select(model,
         starts_with("wss.95"),
         starts_with("rrf.10"), 
         starts_with("loss")) %>%
  select(model,
         ends_with("ace"),
         ends_with("nudging"),
         ends_with("ptsd"),
         ends_with("software"),
         ends_with("virus"),
         ends_with("wilson")
         )

saveRDS(stabres, "output/tabresults.RDS")
```

```{r}

metricsat <- c("WSS@95", "RRF@10", "TTD")
metrics <- c("WSS", "RRF", "TTD")
at <- c("@95", "@10", "")

Data <- c("Ace", "Nudging", "PTSD", "Software", "Virus", "Wilson")

typology <- data.frame(
  col_keys = names(stabres)[2:19],
  what = rep(Data, each = length(metrics)),
  #measure = rep(metricsat, each = 6),
  what = rep(c(rep("WSS@95"), rep("RRF@10"), rep("TTD")), 6),
  #measure = c(rep(c("@95", "@10", ""), each = 6)),
  stringsAsFactors = FALSE
)
num_keys <- names(stabres)[2:19]
test <- flextable(stabres)
# format numbers
test <- colformat_num(x = test, j = num_keys, big.mark = ",", digits = 2, na_str = "missing")
test <- colformat_char(x=test, j = "model")
# set all columns 
test <- set_header_df(test, mapping = typology, key = "col_keys")

test
test <- merge_h(test, part = "header")
test
#border_v = fp_border(color="purple")
#test <- border_inner_v(test, part="all", border = border_v )
test <- merge_v(test, part = "header")

test

test <- theme_booktabs(test)
test <- autofit(test, add_w = 0.2)
test <- fix_border_issues(test)

test
test <- align(test, align = "left", part = "all" )
test
```

```{r}
# make table for only ace

# add ptsd logistic + tf-idf 
ptsd_lctd <- fromJSON(file = paste0("/Volumes/Gerbrich/asreview/results/one_seed/statistics/LCTD/ptsd/seed_lctd_ptsd_.json"))
ptsd_lctd <- ptsd_lctd$.
ptsd_lctd$wss
```

Now again for with seed 
```{r}
#files = list.files("/Volumes/Gerbrich/asreview/results/one/statistics/", recursive = TRUE)

```