### 2020-05-13 datasets/readme.Rmd
update research archive

---
title: "Preprocessing datasets"
output: github_document
---

This notebook processes six systematic review datasets into datasets suitable for simulating the screening process.

- Wilson - [@Appenzeller-Herzog2020] is on a review on effectiveness and safety of treatments of Wilson Disease, a rare genetic disorder of copper metabolism [@Appenzeller-Herzog2019]. 
- Ace -  contains publications on the efficacy of Angiotensin-converting enzyme (ACE) inhibitors, a drug treatment for heart disease [@Cohen2006]. 
- Virus -  is from a systematic review on studies that performed viral Metagenomic Next-Generation Sequencing (mNGS) in farm animals [@Kwok2020]. 
- Software - from the software engineering field, contains publications on fault prediction in source code [@Hall2012].
- Nudging dataset - [@Nagtegaal2019a] on nudging healthcare professionals [@Nagtegaal2019], stemming from the area of 
- PTSD - on studies applying latent trajectory analyses on posttraumatic stress after exposure to traumatic events [@vandeSchoot2017].

Data were preprocessed from their original source into a test dataset, containing title and abstract of the publications obtained in the initial search. Candidate studies with missing abstracts and duplicate instances were removed from the data.

Every dataset has 2 versions: one for hyperparameter optimization `test_datasets` and one for simulation `sim_datasets`.

Every dataset is a .csv file with the following columns: 
  - title
  - abstract
  - label_included
  - keywords? (only for hyperparameter optimization, not for simulation).

```{r, include = FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
```

### Set-up
```{r, echo = FALSE}
datasets <- c("ace", "nudging", "ptsd", "software", "virus", "wilson")
nd <- length(datasets)
# create storage for statistics
template <- data.frame(search = rep(0,4), # papers obtained in systematic search
                       ftext = rep(0, 4), # number of abstract inclusions
                       incl = rep(0,4), # papers included in final systematic review

                       row.names = c("paper", # numbers reported in the publication
                                     "raw", # numbers published on raw dataset (by author,  mostly on OSF)
                                     "asreview", # numbers on asreview repository 
                                     "test")) # numbers in test dataset

# statistics on duplicates and dropping in ASReview dataset
drops <- data.frame(n = rep(0,nd), # candidates
                    na = rep(0,nd),# number of missing abstracts
                    narate = rep(0,nd), # percentage of missing abstracts
                    dup = rep(0,nd),  # number of duplicate abstracts 
                    duprate = rep(0,nd), # percentage of missing abstracts
                    row.names = datasets)

# some functions to extract statistics
data_descr <- function(data){
  return(c(nrow(data), # all papers obtained in the systematic search
           ifelse(sum(data$inclusion_code > 0), sum(data$inclusion_code > 0), NA), # number of abstract inclusions
           sum(data$included))) # papers included in systematic review
}

nostudies <- function(all, set, stage){ # number of studies
  sapply(all, function(x) x[set, stage])
}

inclrate <- function(all, set){ # inclusion rate
  sapply(all, function(x) round(x[set,"incl"]/x[set,"search"]*100,2))
}
```

## ACE
```{r}
ace <- template
# paper ------------------------------------------------------------------------
ace["paper", ] <- c(2544, NA, round(2544*0.0160))

# raw --------------------------------------------------------------------------
# original paper not to linked directly. cannot be retrieved? 
# checked dataset online by cohen et al. 
#https://dmice.ohsu.edu/cohenaa/systematic-drug-class-review-data.html
ace["raw", ] <- c(2544, NA, 41)

# asreview ---------------------------------------------------------------------
ace_asr <- read.csv("https://raw.github.com/asreview/systematic-review-datasets/master/datasets/Cohen_EBM/output/ACEInhibitors.csv",
                     header=T)
# replace empty abstracts by NA values.

ace_asr[ace_asr$abstract == "", "abstract"] <- NA
# stats on missingness in this data
drops["ace",] <- ace_asr %>%
           summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    na_rate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    dup_rate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)

ace["asreview", ] <- c(nrow(ace_asr),
                       NA,
                       sum(ace_asr$label_included))
```

### Create test dataset
```{r}
# drop missings and duplicates. 
ace_test <- ace_asr %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  distinct(abstract, .keep_all = TRUE) # remove entries with duplicate abstracts

# data on 
ace["test", ] <- c(nrow(ace_test),
                   NA,
                   sum(ace_test$label_included))
```

### Some statistics 
```{r}
# missingness and duplicates in raw data on ASReview
drops["ace",]
ace
```

## nudging dataset 
```{r}
# by Rosanna Nagtegaal
nudging <- template
nudging["paper", ] <- c(2006, 377, 100) 
```

```{r eval = TRUE}
# data is not public yet

# raw --------------------------------------------------------------------------
n_raw <- read.csv("../../scripting/nagtegaal_preprocessing/data/output/nagtegaal.csv")
n_raw <- read.csv("../../scripting/nagtegaal_preprocessing/data/output/nagtegaal.csv")
#n_rr <- read.csv2("https://dataverse.harvard.edu/api/access/datafile/:persistentId?persistentId=doi:10.7910/DVN/WMGPGZ/HY6N2S",
                 # sep = ",")
#n_rr <- fread("https://dataverse.harvard.edu/api/access/datafile/:persistentId?persistentId=doi:10.7910/DVN/WMGPGZ/HY6N2S",
              #    sep = ",", encoding = "UTF-8")
# missing statistics
drops["nudging",] <- n_raw %>%
          summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    na_rate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    dup_rate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)
# asreview ---------------------------------------------------------------------

# remove missing abstracts 
nudging_test <- n_raw %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  filter(abstract != "", abstract != "NA") %>%
  distinct(abstract, .keep_all = TRUE) %>% # remove entries with duplicate abstracts
  mutate(id = X)

nudging["test", ] <- data_descr(nudging_test)

```

### Create test dataset

### Some statistics
```{r}

drops["nudging",]
nudging
```

## PTSD dataset
```{r}
ptsd <- template 
ptsd["paper", ] <- c(6185, 363, 38) # from flowchart

# raw --------------------------------------------------------------------------
ptsd["raw", ]

# asreview ---------------------------------------------------------------------
ptsd_asr <- read.csv("https://raw.github.com/asreview/systematic-review-datasets/master/datasets/Van_de_Schoot_PTSD/output/PTSD_VandeSchoot_18.csv", 
                     header=T)

ptsd["asreview", ] <- data_descr(ptsd_asr)

# statistics on duplicates from asr data to test
drops["ptsd", ] <- ptsd_asr %>%
  summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    narate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)
```

### Create test dataset
Now, remove duplicate entries and empty abstracts to arrive at test set. 
```{r}
ptsd_test <- ptsd_asr %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  filter(abstract != "", abstract != "NA") %>%
  distinct(abstract, .keep_all = TRUE)# remove entries with duplicate abstracts

```

There are 38 inclusions in this dataset, should be 34. 
```{r}
# get rid of 4 extra inclusions (see log book)
excl <- 
  ptsd_test %>%
  filter(included == 1, 
         str_detect(authors, "Sterling, M., Hendrikz, J., Kenardy, J.") |
         str_detect(authors, "Hou") |
         str_detect(authors, "Mason") |
         str_detect(authors, "Pérez")) %>%
  select(id)

indices <- ptsd_test$id %in% excl

# get rid of 4 papers
ptsd_test[indices, "included"] <- 0
```

```{r}
# finalize dataset  
ptsd_test <- ptsd_test %>%
  select(id, authors, title, abstract, keywords, included, inclusion_code) # select relevant columns

ptsd["test", ] <- data_descr(ptsd_test)
```

Derive statistics in final test set.
```{r}
drops["ptsd", ]
ptsd
```

## software dataset
```{r}
hall <- template

# paper ------------------------------------------------------------------------
hall["paper", ] <- c(8911, NA, 104)

# raw --------------------------------------------------------------------------
h_raw <- read.csv("https://zenodo.org/record/1162952/files/Hall.csv",
                  header=T)

hall["raw", ] <- c(length(h_raw$Document.Title),
                   NA,
                   sum(h_raw$label == "yes"))

# asreview ---------------------------------------------------------------------
hall_asr <- read.csv("https://raw.githubusercontent.com/asreview/systematic-review-datasets/master/datasets/Four%20Software%20Engineer%20Data%20Sets/output/Software_Engineering_Hall.csv",
                     header=T)

hall["asreview", ] <- data_descr(hall_asr)
```

### Creating test dataset
```{r}
hall_test <- hall_asr %>%
   drop_na(abstract) %>% # drop entries with missing abstracts
   filter(abstract != "") %>%
   distinct(abstract, .keep_all = TRUE) # remove entries with duplicate abstracts

hall_test <- hall_test %>% # add id
  mutate(id = 1:nrow(hall_test))

hall["test", ] <- data_descr(hall_test)
```

### Statistics 
```{r}
drops["software", ] <- 
  hall_asr %>%
  summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    narate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)

hall
```

## Virus dataset
```{r}
virus <- template
virus["paper",] <- c(2481, 132, 120)
v_raw <- read.csv("../../sims/data/virus.csv")
v_raw$abstract[v_raw$abstract == ""] <- NA

virus["raw", ] <- data_descr(v_raw)

drops["virus", ] <- 
  v_raw %>%
  summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    narate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)

v_test <- v_raw %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
   filter(abstract != "") %>%
   distinct(abstract, .keep_all = TRUE)# remove entries with duplicate abstracts

virus["test",] <- data_descr(v_test)
```

## Wilson dataset

```{r}
wilson <- template
# paper ------------------------------------------------------------------------
wilson["paper", ] <- c(3453, 174, 26)

# raw --------------------------------------------------------------------------
# w_ftext <- read.delim("../datasets/raw/DOKU_All FT-Screening_20200116_cap.txt")
# w_incl <- read.csv("../datasets/raw/DOKU_All Included_20200116_cap.csv")
# w_all <- read.csv("../datasets/raw/DOKU_All TiAb-Screening_20200116_cap.csv")
# wilson["raw", ] <- c(nrow(w_all), nrow(w_ftext), nrow(w_incl))
wilson["raw", ] <- c(3453, 174, 26) # looked it up

# asreview ---------------------------------------------------------------------
w_asr <- read.csv("https://raw.github.com/asreview/systematic-review-datasets/master/datasets/Appenzeller-Herzog_Wilson/output/output_csv_wilson.csv", header=T)

w_asr[w_asr$abstract == "", "abstract"] <- NA
drops["wilson", ] <- w_asr %>%
   summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    narate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)
# add inclusion_code label 
w_asr$inclusion_code <- w_asr$label_abstract_screening + w_asr$label_included
w_asr$included <- w_asr$label_included
wilson["asreview", ] <- data_descr(w_asr)
```

```{r}
# create test dataset
w_test <- w_asr %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  distinct(abstract, .keep_all = TRUE) # remove entries with duplicate abstracts

wilson["test",] <- data_descr(w_test)
```

Statistics
```{r}
drops["wilson",]

wilson
```

# Save descriptive statistics on all datasets

```{r eval = TRUE}
# put everything together in list. 
all <- list(Ace = ace,
            Nudging = nudging,
            PTSD = ptsd,
            Software = hall,
            Virus = virus,
            Wilson = wilson)

# save datafile, to serve as data for descriptive table in manuscript
saveRDS(all, file = "data_statistics/all.RDS")

saveRDS(drops, file = "data_statistics/drops.RDS")
```

# Statistics
All candidate papers, paper selected for full text screening, papers included in final review and inclusion rate, over various versions of the dataset.  
```{r, echo = FALSE}
tibble(Dataset = names(all), 
       # paper 
       `candidates_paper` = nostudies(all, "paper", "search"), 
       `fulltext_paper` = nostudies(all, "paper", "ftext"),
       `incl_paper` = nostudies(all, "paper", "incl"),
       `inclrate_paper` = inclrate(all, "paper"),

       # raw data 
       `candidates_raw` = nostudies(all, "raw", "search"), 
       `fulltext_raw` = nostudies(all, "raw", "ftext"),
       `incl_raw` = nostudies(all, "raw", "incl"),
       `inclrate_raw` = inclrate(all, "raw"),

       # asreview repository 
       `candidates_asr` = nostudies(all, "asreview", "search"), 
       `fulltext_asr` = nostudies(all, "asreview", "ftext"),
       `incl_asr` = nostudies(all, "asreview", "incl"),
       `inclrate_asr` = inclrate(all, "asreview"),

       # test set 
       `candidates_test` =  nostudies(all, "test", "search"), 
       `fulltext_test` =  nostudies(all, "test", "ftext"), 
       `incl_test` =  nostudies(all, "test", "incl"), 
       `inclrate_test` = inclrate(all, "test")) %>% 
  t() %>% 
  kable("markdown") 
# %>%
#   kable_styling(full_width = TRUE) %>%
#   #scroll_box(width = "100%", height = "200px") %>%
#   pack_rows("Paper", 2, 5) %>%
#   pack_rows("Raw data", 6, 9) %>%
#   pack_rows("ASReview data", 10, 13) %>%
#   pack_rows("Test data set", 14, 17)  
```

Descriptives on missingness and duplicate abstracts in the ASReview test data set. 
```{r, echo = FALSE}
drops %>%
  kable("markdown",
        col.names = c("n", "NA", "NA rate (%)", "duplicates", "duplicate rate (%)"),
        digits = 2) 
# %>%
#   kable_styling(full_width = TRUE)

```

# Write all test dataset files 
```{r, eval = FALSE}
# ace
write.csv(ace_test %>%  select(pubmedID, authors, title, abstract, keywords, label_included), "test_datasets/ace.csv", row.names = FALSE)

# nudging (no keywords available)
write.csv(nudging_test %>% select(id, title, abstract, included), "test_datasets/nudging.csv", row.names = FALSE)

# ptsd
write.csv(ptsd_test %>% select(id, authors, title, abstract, keywords, included), "test_datasets/ptsd.csv", row.names = FALSE)

# hall (no keywords available )
write.csv(hall_test, "test_datasets/software.csv", row.names = FALSE)

# virus 
write.csv(v_test %>% select(id, authors, title, abstract, keywords, included), "test_datasets/virus.csv", row.names = FALSE)

# wilson
write.csv(w_test %>% select(id, authors, title, abstract, keywords, included), "test_datasets/wilson.csv", row.names = FALSE)
```

# write test dataset files for simulation (removing keywords)
```{r, eval = FALSE}
# ace
write.csv(ace_test %>%  select(pubmedID, title, abstract, label_included), "sim_datasets/ace.csv", row.names = FALSE)
# nudging (no keywords available)
write.csv(nudging_test %>% select(id, title, abstract, included), "sim_datasets/nudging.csv", row.names = FALSE)
# ptsd
write.csv(ptsd_test %>% select(id, title, abstract, included), "sim_datasets/ptsd.csv", row.names = FALSE)
# hall (no keywords available )
write.csv(hall_test, "sim_datasets/software.csv", row.names = FALSE)
# virus 
write.csv(v_test %>% select(id, title, abstract, included), "sim_datasets/virus.csv", row.names = FALSE)
# wilson
write.csv(w_test %>% select(id, title, abstract, included), "sim_datasets/wilson.csv", row.names = FALSE)
```

To check:
```{bash , eval = FALSE}
asreview stat test_datasets/*
```

returns 
```
************  ace.csv  ************

Number of papers:            2235
Number of inclusions:        41 (1.83%)
Number of exclusions:        2194 (98.17%)
Number of unlabeled:         0 (0.00%)
Average title length:        123
Average abstract length:     1623
Average number of keywords:  15.0
Number of missing titles:    0 (of which 0 included)
Number of missing abstracts: 0 (of which 0 included)

************  nudging.csv  ************

Number of papers:            1847
Number of inclusions:        100 (5.41%)
Number of exclusions:        1747 (94.59%)
Number of unlabeled:         0 (0.00%)
Average title length:        109
Average abstract length:     1831
Average number of keywords:  None
Number of missing titles:    0 (of which 0 included)
Number of missing abstracts: 0 (of which 0 included)

************  ptsd.csv  ************

Number of papers:            5031
Number of inclusions:        38 (0.76%)
Number of exclusions:        4993 (99.24%)
Number of unlabeled:         0 (0.00%)
Average title length:        104
Average abstract length:     1537
Average number of keywords:  9.9
Number of missing titles:    1 (of which 0 included)
Number of missing abstracts: 0 (of which 0 included)

************  software.csv  ************

Number of papers:            8896
Number of inclusions:        104 (1.17%)
Number of exclusions:        8792 (98.83%)
Number of unlabeled:         0 (0.00%)
Average title length:        81
Average abstract length:     896
Average number of keywords:  None
Number of missing titles:    0 (of which 0 included)
Number of missing abstracts: 0 (of which 0 included)

************  virus.csv  ************

Number of papers:            2304
Number of inclusions:        114 (4.95%)
Number of exclusions:        2190 (95.05%)
Number of unlabeled:         0 (0.00%)
Average title length:        103
Average abstract length:     1345
Average number of keywords:  20.3
Number of missing titles:    0 (of which 0 included)
Number of missing abstracts: 0 (of which 0 included)

************  wilson.csv  ************

Number of papers:            2333
Number of inclusions:        23 (0.99%)
Number of exclusions:        2310 (99.01%)
Number of unlabeled:         0 (0.00%)
Average title length:        83
Average abstract length:     1325
Average number of keywords:  31.3
Number of missing titles:    1 (of which 0 included)
Number of missing abstracts: 0 (of which 0 included)

```