### 2020-05-14 datasets/readme.Rmd
readme update

---
title: "Preprocessing datasets"
output: github_document
bibliography: ../manuscript/manuscript/asreview.bib
---

This notebook processes six systematic review datasets into datasets suitable for simulating the screening process.

- __Wilson__ - is on a review on effectiveness and safety of treatments of Wilson Disease, a rare genetic disorder of copper metabolism. Dataset: [@Appenzeller-Herzog2020]. Paper: [@Appenzeller-Herzog2019]. 
- __Ace__ -  contains publications on the efficacy of Angiotensin-converting enzyme (ACE) inhibitors, a drug treatment for heart disease Paper and dataset: [@Cohen2006]. 
- __Virus__ -  is from a systematic review on studies that performed viral Metagenomic Next-Generation Sequencing (mNGS) in farm animals [@Kwok2020]. 
- __Software__ - from the software engineering field, contains publications on fault prediction in source code [@Hall2012].
- __Nudging__ dataset - [@Nagtegaal2019a] on nudging healthcare professionals [@Nagtegaal2019], stemming from the area of 
- __PTSD__ - on studies applying latent trajectory analyses on posttraumatic stress after exposure to traumatic events [@vandeSchoot2017].

Data were preprocessed from their original source into a test dataset, containing title and abstract of the publications obtained in the initial search. Candidate studies with missing abstracts and duplicate instances were removed from the data.

Every dataset has 2 versions: one for hyperparameter optimization (`test_datasets`), and one for simulation (`sim_datasets`).

Every simulation dataset is a .csv file with the columns `title`, `abstract`, `label_included`, the last one indicating whether a publication was relevant or not. This is the label that is queried by the active learning model when simulating a systematic review. The reason for this sparse datasets (no keywords are included etc) is twofold. 1) This is by PRISMA guidelines the only information that should be used to identify relevant publications from a search and 2) I wanted equal information for all datasets, and for some datasets this was the only information that was available. For hyperparameter optimization I used all information accessible to arrive at better hyperparameters, with future projects in mind.

```{r, include = FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
```

### Preprocessing datasets
For every datasets, 3 numbers are given: total number of publications, abstract inclusions and final inclusions. For some datasets there are discrepancies between how these numbers are reported in the publication and how they actually are in the raw dataset. Therefore, these numbers are given for how they are reported in the manuscript, how they are found in the raw data, and how they are found in the test dataset, e.g. after removing duplicates and missing abstracts. Moreover, statistics on missingness and duplicates are given.

```{r, echo = FALSE}
datasets <- c("ace", "nudging", "ptsd", "software", "virus", "wilson")
nd <- length(datasets)
# create storage for statistics
template <- data.frame(search = rep(0,4), # papers obtained in systematic search
                       ftext = rep(0, 4), # number of abstract inclusions
                       incl = rep(0,4), # papers included in final systematic review

                       row.names = c("paper", # numbers reported in the publication
                                     "raw", # numbers published on raw dataset (by author,  mostly on OSF)
                                     "asreview", # numbers on asreview repository 
                                     "test")) # numbers in test dataset

# statistics on duplicates and dropping in ASReview dataset
drops <- data.frame(n = rep(0,nd), # candidates
                    na = rep(0,nd),# number of missing abstracts
                    narate = rep(0,nd), # percentage of missing abstracts
                    dup = rep(0,nd),  # number of duplicate abstracts 
                    duprate = rep(0,nd), # percentage of missing abstracts
                    row.names = datasets)

# some functions to extract statistics
data_descr <- function(data){
  return(c(nrow(data), # all papers obtained in the systematic search
           ifelse(sum(data$inclusion_code > 0), sum(data$inclusion_code > 0), NA), # number of abstract inclusions
           sum(data$included))) # papers included in systematic review
}

nostudies <- function(all, set, stage){ # number of studies
  sapply(all, function(x) x[set, stage])
}

inclrate <- function(all, set){ # inclusion rate
  sapply(all, function(x) round(x[set,"incl"]/x[set,"search"]*100,2))
}
```

## Ace dataset
```{r}
ace <- template

# information given in the paper -----------------------------------------------
ace["paper", ] <- c(2544, NA, round(2544*0.0160))

# raw --------------------------------------------------------------------------
# original paper not to linked directly. cannot be retrieved? 
# checked dataset online by cohen et al. 
#https://dmice.ohsu.edu/cohenaa/systematic-drug-class-review-data.html
ace["raw", ] <- c(2544, NA, 41)

# asreview dataset -------------------------------------------------------------
ace_asr <- read.csv("https://raw.github.com/asreview/systematic-review-datasets/master/datasets/Cohen_EBM/output/ACEInhibitors.csv",
                     header=T)
```

### Create test dataset
```{r}
# replace empty abstracts by NA values. 
ace_asr[ace_asr$abstract == "", "abstract"] <- NA
# stats on missingness in this data
drops["ace",] <- ace_asr %>%
           summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    na_rate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    dup_rate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)

ace["asreview", ] <- c(nrow(ace_asr),
                       NA,
                       sum(ace_asr$label_included))

# drop missings and duplicates. 
ace_test <- ace_asr %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  distinct(abstract, .keep_all = TRUE) # remove entries with duplicate abstracts

# data on 
ace["test", ] <- c(nrow(ace_test),
                   NA,
                   sum(ace_test$label_included))
```

### Some statistics 
On missingness (# and %) and duplicates (# and %): 
```{r, echo = FALSE}
# missingness and duplicates in raw data on ASReview
drops["ace",] %>% knitr::kable(format = "markdown", digits = 2)
```

On total number of publications, abstract inclusions (this information was not available for this dataset) and final inclusions. 
```{r, echo = FALSE}
ace %>% knitr::kable(format = "markdown")
```

## Nudging dataset 
```{r}
# by Rosanna Nagtegaal
nudging <- template
nudging["paper", ] <- c(2006, 377, 100) 
```

```{r eval = TRUE}
# data is not public yet

# raw --------------------------------------------------------------------------
n_raw <- read.csv("../../scripting/nagtegaal_preprocessing/data/output/nagtegaal.csv")
n_raw <- read.csv("../../scripting/nagtegaal_preprocessing/data/output/nagtegaal.csv")
#n_rr <- read.csv2("https://dataverse.harvard.edu/api/access/datafile/:persistentId?persistentId=doi:10.7910/DVN/WMGPGZ/HY6N2S",
                 # sep = ",")
#n_rr <- fread("https://dataverse.harvard.edu/api/access/datafile/:persistentId?persistentId=doi:10.7910/DVN/WMGPGZ/HY6N2S",
              #    sep = ",", encoding = "UTF-8")
# missing statistics
drops["nudging",] <- n_raw %>%
          summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    na_rate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    dup_rate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)
# asreview ---------------------------------------------------------------------

# remove missing abstracts 
nudging_test <- n_raw %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  filter(abstract != "", abstract != "NA") %>%
  distinct(abstract, .keep_all = TRUE) %>% # remove entries with duplicate abstracts
  mutate(id = X)

nudging["test", ] <- data_descr(nudging_test)

```

### Some statistics
On missingness (# and %) and duplicates (# and %):

```{r, echo = FALSE}
drops["nudging",]%>% knitr::kable(format = "markdown")
```
On total number of publications, abstract inclusions (this information was not available for this dataset) and final inclusions. 
```{r, echo = FALSE}
nudging %>% knitr::kable(format = "markdown")
```

## PTSD dataset
```{r}
ptsd <- template 
ptsd["paper", ] <- c(6185, 363, 38) # from flowchart

# raw --------------------------------------------------------------------------
ptsd["raw", ]

# asreview ---------------------------------------------------------------------
ptsd_asr <- read.csv("https://raw.github.com/asreview/systematic-review-datasets/master/datasets/Van_de_Schoot_PTSD/output/PTSD_VandeSchoot_18.csv", 
                     header=T)

ptsd["asreview", ] <- data_descr(ptsd_asr)

# statistics on duplicates from asr data to test
drops["ptsd", ] <- ptsd_asr %>%
  summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    narate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)
```

### Create test dataset
Now, remove duplicate entries and empty abstracts to arrive at test set. 
```{r}
ptsd_test <- ptsd_asr %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  filter(abstract != "", abstract != "NA") %>%
  distinct(abstract, .keep_all = TRUE)# remove entries with duplicate abstracts

```

There are 38 inclusions in this dataset, should be 34. 
```{r}
# get rid of 4 extra inclusions (see log book)
excl <- 
  ptsd_test %>%
  filter(included == 1, 
         str_detect(authors, "Sterling, M., Hendrikz, J., Kenardy, J.") |
         str_detect(authors, "Hou") |
         str_detect(authors, "Mason") |
         str_detect(authors, "Pérez")) %>%
  select(id)

indices <- ptsd_test$id %in% excl

# get rid of 4 papers
ptsd_test[indices, "included"] <- 0
```

```{r}
# finalize dataset  
ptsd_test <- ptsd_test %>%
  select(id, authors, title, abstract, keywords, included, inclusion_code) # select relevant columns

ptsd["test", ] <- data_descr(ptsd_test)
```

### Some statistics
On missingness (# and %) and duplicates (# and %):

```{r, echo = FALSE}
drops["ptsd",]%>% knitr::kable(format = "markdown")
```
On total number of publications, abstract inclusions (this information was not available for this dataset) and final inclusions. 
```{r, echo = FALSE}
ptsd %>% knitr::kable(format = "markdown")
```

## Software dataset
```{r}
hall <- template

# paper ------------------------------------------------------------------------
hall["paper", ] <- c(8911, NA, 104)

# raw --------------------------------------------------------------------------
h_raw <- read.csv("https://zenodo.org/record/1162952/files/Hall.csv",
                  header=T)

hall["raw", ] <- c(length(h_raw$Document.Title),
                   NA,
                   sum(h_raw$label == "yes"))

# asreview ---------------------------------------------------------------------
hall_asr <- read.csv("https://raw.githubusercontent.com/asreview/systematic-review-datasets/master/datasets/Four%20Software%20Engineer%20Data%20Sets/output/Software_Engineering_Hall.csv",
                     header=T)

hall["asreview", ] <- data_descr(hall_asr)
```

### Creating test dataset
```{r}
hall_test <- hall_asr %>%
   drop_na(abstract) %>% # drop entries with missing abstracts
   filter(abstract != "") %>%
   distinct(abstract, .keep_all = TRUE) # remove entries with duplicate abstracts

hall_test <- hall_test %>% # add id
  mutate(id = 1:nrow(hall_test))

hall["test", ] <- data_descr(hall_test)
```

```{r}
drops["software", ] <- 
  hall_asr %>%
  summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    narate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)

```

### Some statistics
On missingness (# and %) and duplicates (# and %):

```{r, echo = FALSE}
drops["hall",]%>% knitr::kable(format = "markdown")
```

On total number of publications, abstract inclusions (this information was not available for this dataset) and final inclusions. 
```{r, echo = FALSE}
hall %>% knitr::kable(format = "markdown")
```

## Virus dataset
```{r}
virus <- template
virus["paper",] <- c(2481, 132, 120)
v_raw <- read.csv("../../sims/data/virus.csv")
v_raw$abstract[v_raw$abstract == ""] <- NA

virus["raw", ] <- data_descr(v_raw)

drops["virus", ] <- 
  v_raw %>%
  summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    narate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)

v_test <- v_raw %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
   filter(abstract != "") %>%
   distinct(abstract, .keep_all = TRUE)# remove entries with duplicate abstracts

virus["test",] <- data_descr(v_test)
```

### Some statistics
On missingness (# and %) and duplicates (# and %):

```{r, echo = FALSE}
drops["virus",]%>% knitr::kable(format = "markdown")
```

On total number of publications, abstract inclusions (this information was not available for this dataset) and final inclusions. 
```{r, echo = FALSE}
virus %>% knitr::kable(format = "markdown")
```

## Wilson dataset

```{r}
wilson <- template
# paper ------------------------------------------------------------------------
wilson["paper", ] <- c(3453, 174, 26)

# raw --------------------------------------------------------------------------
# w_ftext <- read.delim("../datasets/raw/DOKU_All FT-Screening_20200116_cap.txt")
# w_incl <- read.csv("../datasets/raw/DOKU_All Included_20200116_cap.csv")
# w_all <- read.csv("../datasets/raw/DOKU_All TiAb-Screening_20200116_cap.csv")
# wilson["raw", ] <- c(nrow(w_all), nrow(w_ftext), nrow(w_incl))
wilson["raw", ] <- c(3453, 174, 26) # looked it up

# asreview ---------------------------------------------------------------------
w_asr <- read.csv("https://raw.github.com/asreview/systematic-review-datasets/master/datasets/Appenzeller-Herzog_Wilson/output/output_csv_wilson.csv", header=T)

w_asr[w_asr$abstract == "", "abstract"] <- NA
drops["wilson", ] <- w_asr %>%
   summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    narate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)
# add inclusion_code label 
w_asr$inclusion_code <- w_asr$label_abstract_screening + w_asr$label_included
w_asr$included <- w_asr$label_included
wilson["asreview", ] <- data_descr(w_asr)
```

```{r}
# create test dataset
w_test <- w_asr %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  distinct(abstract, .keep_all = TRUE) # remove entries with duplicate abstracts

wilson["test",] <- data_descr(w_test)
```

### Some statistics
On missingness (# and %) and duplicates (# and %):

```{r, echo = FALSE}
drops["wilson",]%>% knitr::kable(format = "markdown")
```

On total number of publications, abstract inclusions (this information was not available for this dataset) and final inclusions. 
```{r, echo = FALSE}
wilson %>% knitr::kable(format = "markdown")
```

# Save descriptive statistics on all datasets

```{r eval = TRUE}
# put everything together in list. 
all <- list(Ace = ace,
            Nudging = nudging,
            PTSD = ptsd,
            Software = hall,
            Virus = virus,
            Wilson = wilson)

# save datafile, to serve as data for descriptive table in manuscript
saveRDS(all, file = "data_statistics/all.RDS")
saveRDS(drops, file = "data_statistics/drops.RDS")
```

# Statistics
All candidate papers, paper selected for full text screening, papers included in final review and inclusion rate, for the test datasets. 
```{r, echo = FALSE, eval = TRUE, warning = FALSE}
tibble(Dataset = names(all), 
       # paper 
       # `candidates_paper` = nostudies(all, "paper", "search"), 
       # `fulltext_paper` = nostudies(all, "paper", "ftext"),
       # `incl_paper` = nostudies(all, "paper", "incl"),
       # `inclrate_paper` = inclrate(all, "paper"),
       # 
       # # raw data 
       # `candidates_raw` = nostudies(all, "raw", "search"), 
       # `fulltext_raw` = nostudies(all, "raw", "ftext"),
       # `incl_raw` = nostudies(all, "raw", "incl"),
       # `inclrate_raw` = inclrate(all, "raw"),
       # 
       # # asreview repository 
       # `candidates_asr` = nostudies(all, "asreview", "search"), 
       # `fulltext_asr` = nostudies(all, "asreview", "ftext"),
       # `incl_asr` = nostudies(all, "asreview", "incl"),
       # `inclrate_asr` = inclrate(all, "asreview"),
       # 
       # test set 
       `candidates_test` =  nostudies(all, "test", "search"), 
       `fulltext_test` =  nostudies(all, "test", "ftext"), 
       `incl_test` =  nostudies(all, "test", "incl"), 
       `inclrate_test` = inclrate(all, "test")) %>% 
  t() %>% 
  kable("markdown", digits = 2) 
# %>%
#   kable_styling(full_width = TRUE) %>%
#   #scroll_box(width = "100%", height = "200px") %>%
#   pack_rows("Paper", 2, 5) %>%
#   pack_rows("Raw data", 6, 9) %>%
#   pack_rows("ASReview data", 10, 13) %>%
#   pack_rows("Test data set", 14, 17)  
```

Descriptives on missingness and duplicate abstracts in the raw datasets: 
```{r, echo = FALSE}
drops %>%
  kable("markdown",
        col.names = c("n", "NA", "NA rate (%)", "duplicates", "duplicate rate (%)"),
        digits = 2)

```

# Write all test dataset files 
```{r, eval = FALSE}
# ace
write.csv(ace_test %>%  select(pubmedID, authors, title, abstract, keywords, label_included), "test_datasets/ace.csv", row.names = FALSE)

# nudging (no keywords available)
write.csv(nudging_test %>% select(id, title, abstract, included), "test_datasets/nudging.csv", row.names = FALSE)

# ptsd
write.csv(ptsd_test %>% select(id, authors, title, abstract, keywords, included), "test_datasets/ptsd.csv", row.names = FALSE)

# hall (no keywords available )
write.csv(hall_test, "test_datasets/software.csv", row.names = FALSE)

# virus 
write.csv(v_test %>% select(id, authors, title, abstract, keywords, included), "test_datasets/virus.csv", row.names = FALSE)

# wilson
write.csv(w_test %>% select(id, authors, title, abstract, keywords, included), "test_datasets/wilson.csv", row.names = FALSE)
```

# write test dataset files for simulation (removing keywords)
```{r, eval = FALSE}
# ace
write.csv(ace_test %>%  select(pubmedID, title, abstract, label_included), "sim_datasets/ace.csv", row.names = FALSE)
# nudging (no keywords available)
write.csv(nudging_test %>% select(id, title, abstract, included), "sim_datasets/nudging.csv", row.names = FALSE)
# ptsd
write.csv(ptsd_test %>% select(id, title, abstract, included), "sim_datasets/ptsd.csv", row.names = FALSE)
# hall (no keywords available )
write.csv(hall_test, "sim_datasets/software.csv", row.names = FALSE)
# virus 
write.csv(v_test %>% select(id, title, abstract, included), "sim_datasets/virus.csv", row.names = FALSE)
# wilson
write.csv(w_test %>% select(id, title, abstract, included), "sim_datasets/wilson.csv", row.names = FALSE)
```

# References