### 2020-05-14 simulation_study/simulation/log.Rmd

---
title: "Simulations log"
author: "Gerbrich Ferdinands"
date: "5/14/2020"
output: github_document
---
This notebook contains all simulations run on the Cartesius supercomputer.

# Run simulations

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

createLines <- function(cond, d, hp, seed = FALSE){
  comm <- "./cart_simulate.py simdata/%s.csv --config_file simconfig/%s/%s/%s-%s.ini --state_file simoutput/%s/%s/%s/results.h5 -r 15"
  files <- data.frame(c = condition, 
                    f = c("nb_max_double_tfidf",
                    "rf_max_double_tfidf",
                    "svm_max_double_tfidf",
                    "logistic_max_double_tfidf",
                    "nn-2-layer_max_double_tfidf",
                    "nb_max_double_doc2vec",
                    "rf_max_double_doc2vec",
                    "svm_max_double_doc2vec",
                    "logistic_max_double_doc2vec",
                    "nn-2-layer_max_double_doc2vec"))

  # create condition grid
  cons <- expand.grid(d = d, hp = hp, c = cond, stringsAsFactors = FALSE)
  cons <- left_join(cons, files)

  # write 
  script <- sprintf(comm, 
        # input data
        cons$d, 
        # config file details
        cons$hp, cons$c, cons$f, cons$d,
        # output file details
        cons$hp, cons$c, cons$d)

  if(seed){
    script <- paste0(script, " --init_seed 42")
  }
return(script)
}
transfer <- function(hp, c, seed = FALSE){
    cons <- expand.grid(hp = hp, c = c, stringsAsFactors = FALSE)

    if(seed){
      txt <- "scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/simoutput/%s/%s /Volumes/Gerbrich/asreview/simoutput/%s_seed/%s"
    } else {
      txt <- "scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/simoutput/%s/%s /Volumes/Gerbrich/asreview/simoutput/%s/%s"
    }
    result <- sprintf(txt, hp, c, hp, c)
    return(result)
}

```

```{r, include=FALSE}
sims <- readRDS("../R/01_hyperparameter_sets.RDS")
d <- colnames(sims)
condition <- rownames(sims)
classifier <- c("nb", "rf", "svm", "logistic", "nn-2-layer")
names(classifier) <- c("B", "R", "S", "L", "N")
```

# create config files 
#### 01.04.2020
```{bash, eval = FALSE}
# get config files from surfsara to hard drive 
scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/config /Volumes/Gerbrich/asreview/config/optimization_output
```

Manually update config files to simulation 
Push finished ones to GH repository
Fetch github repo on surfsara

ASReview version 0.8.1

```{r}
sims[condition[c(1,2,4)], d, "one"] <- 1 # simulation configs are up
```

## hp = one

#### BCTD
```{r}
bctd <- createLines(condition[1], d=d, hp = "one")
write.table(bctd, file = "scripts/one/bctd.txt", sep = "\t", quote = FALSE,
            row.names = FALSE)
names(bctd) <- c("7883184", "7883202", "7883207", "7883209", "7883213", "7883214")
```

#### RCTD, LCTD
```{r}
rctd_lctd <- createLines(condition[c(2,4)], d=d, hp = "one")
write.table(rctd_lctd, file = "scripts/one/rctd_lctd.txt", sep = "\t", quote = FALSE,
            row.names = FALSE)
#one_rctd_virus etc 
names(rctd_lctd) <- c(rep("0", 6), "7949940", "7949952", "7949953", "7949954", "7949956", "7949959")
```

```{r}
sims[condition[c(1,2,4)], d, "one"] <- 2 # simulations are running 
```

# transfer simulation files to harddrive

```{bash, eval = FALSE}
scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/config /Volumes/Gerbrich/asreview/config/optimization_output

./cart_simulate.py simdata/wilson.csv --config_file simconfig/one/LCTD/logistic_max_double_tfidf-wilson.ini --state_file simoutput/one/LCTD/wilson/results.h5 -r 15
```

```{r}
transfer("one", condition[c(1,2,4)])
```

```{r}
# output is on harddrive
sims[condition[1], , "one"] <- 3
```

#### 02.04.2020
get config files 'all' BCTD, LCTD.

scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/config /Volumes/Gerbrich/asreview/config/optimization_output

```{r}
# output is on harddrive
sims[condition[2], , "one"] <- 3
```

#### 03.04.2020
SCDD model

```{r}
# get config files
# manually transform
# push to github 
sims[condition[8], d, "one"] <- 1 # simulation configs are up

# run simulations 
scdd <- createLines(condition[8], d, "one")
write.table(scdd, file = "scripts/one/scdd.txt", sep = "\t", quote = FALSE,
           row.names = FALSE)
 names(scdd) <- c("7950186", "7950187", "7950188", "7950189", "7950190", "7950191")
 sims[condition[8], d, "one"] <- 2 # simulations are running 
```

```{r}
 sctd <- createLines(condition[3], d, "one")
 write.table(sctd, file = "scripts/one/sctd.txt", sep = "\t", quote = FALSE,
            row.names = FALSE)
names(sctd) <- c("7950483", "7950484", "7950485", "7950487", "7950489", "7950490")
sims[condition[3], d, "one"] <- 2 # simulations are running
```

```{r}
rcdd <- createLines(condition[7], d, "one")
write.table(rcdd, file = "scripts/one/rcdd.txt", sep = "\t", quote = FALSE,
            row.names = FALSE)
names(rcdd) <- c("7950491", "7950493", "7950495", "7950496", "7950497", "7950499")
sims[condition[7], d, "one"] <- 2 # simulations are running
```

```{r}
lcdd <- createLines(condition[9], d, "one")
write.table(lcdd, file = "scripts/one/lcdd.txt", sep = "\t", quote = FALSE, row.names = FALSE)
names(lcdd) <- c("7965136", "7965141", "7965143", "7965145", "7965146", "7965147")
sims[condition[9], d, "one"] <- 2 # simulations are running
```

10.04.2020

```{r}
# LCTD and SCDD both failedsimulations both failed? 
# ran again 
# transfer to harddrive
transfer("one", condition[c(4,7:9)])

# sctd is not finished yet! 
sims[condition[4],,"one"] <- 3 # output is on harddrive

# sims[condition[7],,"one"] <- 3 # output is on harddrive 
# sims[condition[8],,"one"] <- 3 # output is on harddrive 
# too little space for 9, so run again

sims[condition[9],,"one"] <- 3 # output is on harddrive 
# but ace wasn't finished yet! 
```

Repeat all failed simulations

```{r}
# run again 
# transfer to harddrive 
transfer("one", condition[8])
```

15.04.2020 
```{r}
transfer("one", condition[3]) # SCTD 
sims[condition[8], , "one"] <- 3 # output is on harddrive

# BCTD run ace dataset again 
createLines(condition[1], "ace", "one") # 8018165
#scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/simoutput/one/BCTD/ace /Volumes/Gerbrich/asreview/simoutput/one/BCTD
sims[condition[1], "ace", "one"] <- 3 # output is on harddrive

# RCDD needs to go all over again
names(rcdd) <- c("8018262", "8018263", "8018264", "8018265", "8018267", "8018268")
transfer("one", "RCDD")

# SCDD need ace ptsd software virus again
# ace 8019246
# ptsd 8019247
# software 8019250
# virus 8019251
transfer("one", "SCDD")

# LCDD only ace 
createLines(condition[9], "ace", "one") # 8019423
#scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/simoutput/one/LCDD/ace /Volumes/Gerbrich/asreview/simoutput/one/LCDD

# LCTD
# software 
createLines("LCTD", "software", "one") # 8020733
# scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/simoutput/one/LCTD/software /Volumes/Gerbrich/asreview/simoutput/one/LCTD

# SCTD 
# ace ptsd software
sctd <- createLines("SCTD", c("ace", "ptsd", "software"), "one")
names(sctd) <- c("8025808", "8025809", "8025810")
#scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/simoutput/one/SCTD/ace /Volumes/Gerbrich/asreview/simoutput/one/SCTD
createLines("SCTD", "ptsd", "one") # 8066095
createLines("SCTD", "software", "one")
```

---

23.04.2020
Simulations for ace dataset again but now with seed. 
```{r, message = FALSE}
seed_ace <- createLines(condition[c(1, 2, 3,4,7,8,9)], "ace", "one", seed = TRUE) %>%
  noquote()
names(seed_ace) <- c(".", "8119459", "8119460", "8119461", "8119462", "8119463", "8119464")

seed_nudging <- createLines(condition[c(1, 2, 3,4,7,8,9)], "nudging", "one", seed = TRUE) %>%
  noquote()
names(seed_nudging) <- c("8119472", "8119472", "8121720", "8121721", "8121722", "8121723", "8121725")

# transfer sctd's later!(3)
noquote(transfer("one", condition[c(1,2,3,4,7,8,9)], seed = TRUE))
# ace: scp -r gerbrich@cartesius.surfsara.nl:asreview/simulations/automated-systematic-review-simulations/scripts/simoutput/one/SCTD/ace /Volumes/Gerbrich/asreview/simoutput/one_seed/SCTD

```

```{r}
seed_virus <- createLines(condition[c(1, 2, 3,4,7,8,9)], "virus", "one", seed = TRUE) %>%
  noquote()
names(seed_virus) <- c("8125572", "8125573", "8125574", "8125575", "8125581", "8125582", "8125583")

seed_wilson <- createLines(condition[c(1, 2, 3,4,7,8,9)], "wilson", "one", seed = TRUE) %>%
  noquote()
names(seed_wilson) <- c("8125584", "8125585", "8125586", "8125587", "8125588", "8125589", "8125590")
```

```{r}
seed_ptsd <- createLines(condition[c(1, 2, 3,4,7,8,9)], "ptsd", "one", seed = TRUE) %>%
  noquote()

# sctd later

seed_software <- createLines(condition[c(1, 2, 3,4,7,8,9)], "software", "one", seed = TRUE) %>%
  noquote()
# sctd later
seed_software
names(seed_software) <- c("8129589", "8129732", "8129588", "8131305", "8133472", "8134718", "8149271")

# ptsd sctd
seed_ptsd # 8149296
```

```{r template, eval = FALSE}
# get config files
# manually transform
# push to github 
sims[condition[1], d, "one"] <- 1 # simulation configs are up
# run simulations 
createLines()
sims[condition[1], d, "one"] <- 2 # simulations are running 
# transfer to harddrive 
transfer()
sims[condition[1], , "one"] <- 3 # output is on harddrive
```