### 2020-05-14 simulation_study/README.Rmd
update table sims

---
title: "Simulation study scripts"
output: github_document
bibliography: ../manuscript/manuscript/asreview.bib
---

This directory contains all steps taken in hyperparameter optimization and simulating automated systematic reviews.

```{r, include = FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
source("R/functions.R")
#source("R/00_condition_grid.R")
```

# Models 
We are simulating 7 models on 6 systematic review datasets. Each model has an abbreviation to ease the programming process: 
```{r, eval = TRUE}
conditiongrid <- readRDS("R/00_conditiongrid.RDS")
conditiongrid <- conditiongrid[c(1:4,7:9),]
conditiongrid %>%
  mutate(Model = c("NB + TF-IDF", "LR + D2V", "LR + TF-IDF", "RF + D2V", "RF + TF-IDF", "SVM + D2V", "SVM + TF-IDF" )) %>%
  mutate(Abbreviation = condition) %>%
  select(Model, Abbreviation) %>%
  kable(format = "markdown")
```
The first letter denotes the classification strategy, NB, LR, RF, or SVM. The second letter stands for C of certainty sampling. The third letter denotes the feature extraction strategy, TF-IDF or Doc2vec. The fourth letter stands for D of Dynamic Resampling.

For every model, first the hyperparameters were optimized. The scripts and results for that can be found in the `hpoptimization` directory. The `simulation` directory contains scripts for simulating an automated systematic review, using the hyperparameters obtained in the optimization.

# Requirements 
This study requires having several packages installed, like ASReview version 0.9.3 [@ASReview2020]. 
All these requirements are listed in the `requirements.txt` file. Assuming you have Python 3.7 or higher, you can run the following in your bash terminal to install all requirements:

```{bash, eval = FALSE}
pip install -r requirements.txt
```

This automatically installs asreview version 0.9.3 and several dependencies. Moreover, a specific branch of the asreview simulation extension should be installed. Run the following in your bash terminal:
```{bash, eval = FALSE}
git clone https://github.com/GerbrichFerdinands/asreview-thesis-simulation.git
```

And then, within the newly created directory, the following:
```{bash, eval = FALSE}
# install visualization package 
pip install . 
```

# References