### 2020-05-14 simulation_study/simulation/README.Rmd

---
title: "Simulations"
output: github_document
---

This directory contains all scripts used to simulate systematic reviews:

- `config` - contains the hyperparameters for every model-dataset combination. These files are called upon in the simulation command. They are copies from the config files in the `hpoptimization/config` directory, except that the `n_instances` parameter has been set to 1, so that the model is retrained after labelling one publication instead of 20. 
- `scripts` -  contains all scripts run on the Cartesius supercomputer. The `log.Rmd` file tracks when all these scripts were executed and connects them to a jobID on Cartesius (just for documentation).
- `simoutput` - where the output of the simulation files is stored. As mentioned in the readme in the root directory, the files resulting from this simulation are enormous (over 500GB in total), therefore they cannot be stored on GitHub and the `simoutput` directory remains empty. Instead, the simulation output files are published on the Open Science Framework: https://osf.io/7mr2g/ and https://osf.io/ag2xp/.

## Running a simulation
All simulations were run with seed value `42`. If you want to run a simulation locally, for example the NB + TF-IDF model on the nudging dataset, for 15 trials, run the following in your bash terminal:

```{bash, eval = FALSE}
asreview batch ../../datasets/sim_datasets/nudging.csv --config_file config/one/BCTD/nb_max_double_tfidf-nudging.ini --state_file simoutput/one/BCTD/nudging/results.h5 -r 15 --init_seed 42
```

The simulation will output 15 `.h5` files. Please refer to the `results` folder in the root directory of this repository for a demonstration of how to analyze the resulting files. Code for processing these files can be found in the `results` directory, which also contains already A processed version of the simulation output files is available in the `results` directory.