### 2020-07-20 manuscript/manuscript/Ferdinands, G - MSBBSS.Rmd
corrctions July 20 2020

---
title: Active learning for efficient systematic reviews
subtitle: Evaluating models across research areas 
author: "Gerbrich Ferdinands"
# keywords:
header-includes:
    - \usepackage{setspace}\doublespacing
    - \usepackage[left]{lineno}
    - \usepackage{amssymb}
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  bookdown::pdf_document2:
    extra_dependencies: subfig, algorithm2e, float
    number_sections: no
    toc: no
  word_document: default
bibliography: asreview.bib
link-citations: yes
csl: elsevier-vancouver.csl
---

<!-- --- -->
<!-- title: Manuscript drafts -->
<!-- author: "Gerbrich Ferdinands" -->
<!-- date: '`r format(Sys.time(), "%d %B, %Y")`' -->
<!-- output: -->

<!--   bookdown::pdf_document2: -->
<!--     extra_dependencies: subfig, algorithm2e, float -->
<!--     number_sections: no -->
<!--     toc: no -->
<!--   word_document: default -->
<!-- bibliography: asreview.bib -->
<!-- link-citations: yes -->
<!-- csl: systematic-reviews.csl -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.ncol = 2, out.width = "50%", fig.align = "center", dev = "png", fig.pos = "H") 
options(tinytex.verbose = TRUE)

library(tidyverse)
library(knitr)
library(xtable)
library(kableExtra)
library(captioner)
library(glue)
library(rbbt)
library(officer)
library(flextable)
# load model parameters 
source("asr-parameters.R")

#pathRdata <- "data/"
pathRdata <- "../../datasets/data_statistics/"
```

```{r}
# configurations table 
tab <- lapply(words, FUN = function(x) paste(x, collapse = ", "))
tab <- do.call(rbind, tab)
# remove row training data (28.02)
colnames(tab) <- c("Configurations")
tab <- tab[-4,]
```

```{r}
# combinations table
params[[4]] <- NULL # remove training data
combins <- expand.grid(params) %>%
  arrange(model, feature_extraction)

# rows to drop
droprows <- combins$model == "nb" & combins$feature_extraction != "tfidf"
combins <- combins[!droprows, ]
```

```{r all tables, results = 'hide'}
table_nums <- captioner(prefix = "Table")
# table  on datasets statistics
table_nums(name = "datasets", caption = "Statistics on the datasets obtained from six original systematic reviews.")
table_nums(name = "atd", caption = "ATD values ($\\bar x (\\hat s)$) for all model-dataset combinations, and median (MAD) for all datasets.")
table_nums(name = "wss", caption = "WSS@95 values ($\\bar x (\\hat s)$) for all model-dataset combinations, and median (MAD) for all datasets.")
table_nums(name = "rrf", caption = "RRF@10 values ($\\bar x, (\\hat s)$) for all model-dataset combinations, and median (MAD) for all datasets.")

```

\newpage

\linenumbers
# Abstract
#### Background 
the context and purpose of the study

#### Methods
how the study was performed and statistical tests used

#### Results
the main findings

#### Conclusions
brief summary and potential implications

keywords:

# Background
Systematic reviews are top of the bill in research. A systematic review brings together all studies relevant to answer a specific research question [@PRISMA-PGroup2015]. Systematic reviews inform practice and policy [@Gough2002] and are key in developing clinical guidelines [@Chalmers2007]. However, systematic reviews are costly because they involve the manual screening of thousands of titles and abstracts to identify publications relevant to answering the research question.

<!-- An experienced reviewer takes on average 30 seconds to screen one title and abstract, whereas an inexperienced reviewer takes even longer [@Wallace2010]. A typical review of 5000 publications would require over 40 hours of uninterrupted screening, while typically only a fraction of the publications is relevant (mean = 2.94%, IQR=2.5 @Borah2017]). -->
Conducting a systematic review typically requires over a year of work by a team of researchers [@Borah2017]. Nevertheless, systematic reviewers are often bound to a limited budget and timeframe. Currently, the demand for systematic reviews exceeds the available time and resources by far [@Lau2019]. Especially when the need for guidelines is urgent it is extremely challenging to provide a review that is both timely and comprehensive.

To ensure a timely review, reducing workload in systematic reviews is essential. With advances in Machine Learning (ML), there has been wide interest in tools to reduce workload in systematic reviews [@Harrison2020]. Various learning models have been proposed, aiming to predict whether a given publication is relevant or irrelevant to the systematic review. Previous findings suggest that such models potentially reduce workload with 30-70% at the cost of losing 5% of relevant publications, i.e. 95% recall [@OMara-Eves2015].
<!-- To date, several studies have begun to examine the use of learning algorithms for screening prioritization. -->

A well-established approach in increasing efficiency in title and abstract screening is screening prioritization [@Cohen2009; @Shemilt2014]. In screening prioritization, the learning model presents the reviewer with the publications that are most likely to be relevant first, thereby expediting the process of finding all of the relevant publications. Such an approach allows for substantial time-savings in the screening process as the reviewer can decide to stop screening after a sufficient number of relevant publications have been retrieved [@Yu2019]. Moreover, the early retrieval of relevant publications facilitates a faster transition of those publications to the next steps in the review process [@Cohen2009]. 
<!-- Additionally, several studies report increasing efficiency beyond saving time [@OMara-Eves2015].  -->

<!-- assist the reviewer in the screening process -->
<!-- Several studies have proposed active learning models for screening prioritization -->
Recent studies have demonstrated the effectiveness of screening prioritization by means of active learning models [@Yu2018; @Yu2019; @Miwa2014; @Cormack2014; @Cormack2015; @Wallace2010; @Gates2018a]. With active learning, the ML model can iteratively improve its predictions on unlabelled data by allowing the model to select the records from which it wants to learn [@Settles2012]. The model queries these records to a human annotator who provides them with a label, from which the model then updates its predictions. The general assumption is that by letting the model select which records are labelled, the model can achieve higher accuracy more quickly while requiring the human annotator to label as few records as possible [@Settles2009]. Active learning has proven to be an efficient strategy in large unlabelled datasets where labels are expensive to obtain [@Settles2009]. This makes the screening phase in systematic reviewing an ideal candidate solution for such models because typically, labelling a large number of publications is very costly. When active learning is applied in the screening phase, the reviewer screens publications that are suggested by an active learning model. Subsequently, the active learning model learns from the reviewers' decision ('relevant', 'irrelevant') and uses this knowledge to update its predictions and to select the next publication to be screened by the reviewer.

<!-- Combined with some sort of stopping criterion, the reviewer c -->
<!-- _Adopting some sort of stopping criterion (outside scope of the current thesis) the reviewer can quit reviewing after having read only a fraction of candidate studies. meaning the screening process can be finished after reading a fraction of all candidate studies. Saving hours of time and resources._ -->

<!-- , the complex nature of the field is making it difficult to draw overarching conclusions about best practice [@OMara-Eves2015]. -->

The application of active learning models in reducing workload of systematic reviews has been extensively studied [@Yu2018; @Yu2019; @Miwa2014; @Wallace2010; @Gates2018a]. Whilst previous studies have evaluated active learning models in many forms and shapes [@Yu2018; @Yu2019; @Miwa2014; @Cormack2014; @Cormack2015; @Wallace2010; @Singh2018; @Carvallo; @Yimin2007], ready-to-use software tools implementing such active learning models (Abstrackr [@Wallace2012b], Colandr [@Cheng2018], FASTREAD [@Yu2018], Rayyan [@Ouzzani2016], and RobotAnalyst [@Przybyla2018]) currently use the same classification technique to predict relevance of publications, namely Support Vector Machine (SVM). Findings show that different classification techniques can serve different needs in the retrieval of relevant publications, for example the desired balance between recall and precision [@Kilicoglu2009; @Aphinyanaphongs2004]. Therefore, is essential to evaluate different classification techniques in the context of active learning models. The current study investigates active learning models adopting four classification techniques: Naive Bayes, Logistic Regression, Support Vector Machine, and Random Forest. These are widely adopted techniques in text classification [@Aggarwal2012] and are suitable for creating interpretable ML models [@Molnar2020]. Moreover, due to their relatively short computation time they are fit for software tools to be used in scientific practice, as they allow the reviewer to be timely presented with the next publication.

<!-- low computation time makes these techniques fit to practial use in a software tool when run on a personal computer, as retraining a model can be done in a timely manner. as default of asreview is on local computer to be applicable -->

<!-- Whilst the performance of several classification techniques has been investigated in the ML-aided title-and-abstract screening field in general [@Kilicoglu2009; @Aphinyanaphongs2004], the relatively new subfield of active learning for screening prioritization has not yet studied the performance of classifiers other than SVMs [@Yu2018; @Yu2019; @Cormack2014; @Cormack2015; @Miwa2014; @Wallace2010]. -->

<!-- no study in this area has made a comparison between models adopting different classification techniques. -->
<!-- Since a plethora of techniques exist [@Aggarwal2012], it would be of interest to incorporate such models to this area as well.  -->
Another component that influences model performance is the way how the textual content of titles and abstracts are represented in a model, called the feature extraction strategy [@Zhang2011; @Carvallo; @Singh2018]. One of the more sophisticated feature extraction strategies is called Doc2vec (D2V), also known as Paragraph vector [@Le2014]. D2V learns continuous distributed vector representations for pieces of text. The idea of distributed representation is that every word can be known by its neighborhood. D2V can overcome problems with more simplistic feature extraction strategies such as TF-IDF, which ignores the the semantics and the ordering of the words [@Le2014]. Therefore, it is of interest to compare models adopting D2V to models adopting TF-IDF.

<!-- [@Yu2018; @Yu2019; @Miwa2014; @Wallace2010], -->

Lastly, previous studies have mainly focussed on reviews from a single scientific field, like medicine [@Wallace2010; @Gates2018a] and computer science [@Yu2018, @Yu2019]. Model replications on reviews from varying research contexts are essential to draw conclusions about the general effectiveness of active learning models [@OMara-Eves2015; @Marshall2020]. As far as known to the authors, Miwa et al [@Miwa2014] were the only researchers to make a direct comparison between two systematic reviews from different research areas, namely the social and the medical sciences. They found that active learning was more difficult on data from the social sciences due to the different nature of the vocabularies used. Thus, it is of interest to evaluate model performance across different research contexts, namely the social science, humanities, medical science and computer science.

Taken together, for a more comprehensive understanding of active learning models in the context of systematic reviewing, a methodical evaluation of such models is required. The current study aims to address this issue by answering the following research questions:

<!-- (1) across different classification techniques, (2) feature extraction strategies, and (3) review contexts.  -->

<!-- Therefore, the current study aims to address these issues by answering the following research questions: -->

|    __RQ1__ What is the performance of active learning models across four different classification techniques?
|    __RQ2__ What is the performance of active learning models across two different feature extraction strategies? 
|    __RQ3__ Does the performance of active learning models differ across six systematic reviews from four different research areas?

The purpose of the current paper is to increase the evidence base on active learning models for reducing workload in title and abstract screening in systematic reviews. We adopt four different classification techniques (NB, LR, SVM, and RF) and two different feature extraction strategies (TF-IDF and D2V) for the purpose of maximizing the number of identified relevant publications, while minimizing the number of publications needed to screen. Models were assessed by conducting a simulation on six systematic review datasets. Datasets were collected from the fields of medical science [@Cohen2006; @Appenzeller-Herzog2019; @Kwok2020], computer science [@Yu2018], humanities [@Nagtegaal2019] and social science [@vandeSchoot2017], to assess generalizability of the models across research contexts. The models, datasets and simulations are implemented in a pipeline of active learning for screening prioritization, called $\texttt{ASReview}$ [@vandeSchoot2020]. $\texttt{ASReview}$ is an open source and generic tool in which users can adapt and add modules as they like, encouraging fellow researchers to replicate findings from previous studies. All scripts and data used are openly published to facilitate usability and acceptability of ML-assisted title and abstract screening in the field of systematic review.

The remaining part of this paper is organized as follows. The methods section describes the study that was designed to answer the research questions and elaborates on the technical details of active learning models for identifying relevant publications in the context of systematic reviews. The findings of the simulation study are reported in the results section. The implications of the findings in context of previous research are discussed in the discussion section, followed by this study's main conclusions in the conclusion section.

# Methods

## Technical details 
What follows is a more detailed account of the active learning models. The structure and functions of the key components of the models are introduced to clarify the choices made in the design of the current study.

### Task description
The screening process of a systematic review starts with all publications obtained in the search. The task is to identify which of these publications are relevant, by screening their titles and abstracts. In active learning for screening prioritization, the screening process proceeds as follows:

<!-- Just a sample algorithmn -->
<!-- \begin{algorithm}[H] -->
<!-- \DontPrintSemicolon -->
<!-- \SetAlgoLined -->
<!-- \KwResult{Write here the result} -->
<!-- \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output} -->
<!-- \Input{Write here the input} -->
<!-- \Output{Write here the output} -->
<!-- \BlankLine -->
<!-- \While{While condition}{ -->
<!--     instructions\; -->
<!--     \eIf{condition}{ -->
<!--         instructions1\; -->
<!--         instructions2\; -->
<!--     }{ -->
<!--         instructions3\; -->
<!--     } -->
<!-- } -->
<!-- \caption{While loop with If/Else condition} -->
<!-- \end{algorithm} -->

- Start with the set of all unlabelled records (titles and abstracts), $\mathcal{U}$.
- The reviewer provides a label for a few, for example 5-10 records $x \in \mathcal{U}$, creating a set of labelled records $x \in \mathcal{L}$ such that $x \notin \mathcal{U}$. The label can be either Relevant $\langle x, \texttt{R}\rangle$ or Irrelevant $\langle x, \texttt{I}\rangle$.
- The active learning cycle starts:

  1. A classifier $C$ is trained on the labelled records $\mathcal{L}$, $C = \texttt{train}(\mathcal{L})$

  2. The classifier predicts relevancy scores for all unlabelled records $\mathcal{U}$, $C(\mathcal{U})$

  3. Based on the predictions by $C$, the model selects the most relevant record $x^* \in \mathcal{U}$

  4. The model queries the reviewer to screen this record, $\langle x^*, \texttt{?}\rangle$

  5. The reviewer screens the record and provides a label, $\langle x^*, \texttt{R}\rangle$ or $\langle x^*, \texttt{I}\rangle$

  6. The newly labelled record is added to the training data, such that $x^* \in \mathcal{L}$ and $x^* \notin \mathcal{U}$

  7. Back to step 1

  8. Repeat this cycle until the reviewer decides to stop [@Yu2019] or until all records have been labelled, $\mathcal{U} = \varnothing$.

In this active learning cycle, the model incrementally improves its predictions on the remaining unlabelled title and abstracts. Relevant titles and abstracts are identified as early in the process as possible.

This case is an example of pool-based active learning, as the next record to be queried is selected by predicting relevancy for all records in a fixed pool [@Settles2012]. Another form of active learning is stream-based active learning, in which data is regarded as a stream instead of a fixed pool, in which the model selects one record at a time and then decides whether or not to query this record. This approach of active learning is preferred when it is expensive or impossible to exhaustively search the data for selecting the next query. A possible application of stream-based active learning is living systematic reviews, as the review is continually updated as new evidence becomes available. For an example see the study by Wynants et al., @Wynants2020.

### Class imbalance problem
There are two classes in the dataset: relevant and irrelevant publications. Typically, the inclusion rate is low as only a fraction of the publications belong to the relevant class (2.94%, [@Borah2017]). The class imbalance causes the classifier to miss relevant publications, because there are far fewer examples of relevant than irrelevant publications to train on [@OMara-Eves2015]. Moreover, classifiers can achieve high accuracy but still fail to identify any of the relevant publications [@Wallace2010]. This is evident in the case of a systematic review dataset where only three percent of publications are relevant.

<!-- A model would achieve 97% accuracy when classifying all publications as irrelevant, even though none of the relevant papers would have been correctly identified.  -->

Previous studies have addressed the class imbalance problem by rebalancing the training data in various ways [@OMara-Eves2015]. To decrease the class imbalance in the training data, the models in the current study rebalance the training set by a technique we propose to call Dynamic Resampling (DR). DR undersamples the number of irrelevant publications in the training data, whereas the number of relevant publications are oversampled such that the size of the training data remains the same. The ratio between relevant and irrelevant publications in the rebalanced training data is not fixed, but dynamically updated depending on the number of publications in the available training data, the total number of publications in the dataset, and the ratio between relevant and irrelevant publications in the available training data.

<!-- This works because..  -->
<!-- This technique is related to aggressive undersampling in such a way that.  -->

### Classification
To make predictions on the unlabelled publications, a classifier is trained on features from the training data. The performance of the following four classifiers is explored:
<!-- The classifier predicts class of a given unlabelled publication.  -->
<!-- Whilst the SVM is a well-known technique, also classifiers have been employed in class prediction tasks in systematic reviews.  -->
<!-- In fact, research in this area to date has not yet determined the performance of classifiers other than SVMs. -->

 - Support Vector Machine (SVM) - SVMs separate the data into classes by finding a multidimensional hyperplane [@Tong2001; @Kremer2014].

 - L2-regularized Logistic Regression (LR) models the probabilities describing the possible outcomes by a logistic function. The coefficients are regularizes/model is regularized. 
 A penalty is imposed on the size of the coefficients, shrinking coefficients of features with a minor contribution to the solution towards zero.

 - Naive Bayes (NB) is a supervised learning algorithm often used in text classification. Based on Bayes' theorem, with the 'naive' assumption that all features are independent given the class value [@Zhang2004].

- Random Forests (RF) is a supervised learning algorithm where a large number of decision trees are fit on bootstrapped samples of the original data. All trees cast a vote on the class, which are aggregated into a class prediction for each instance [@Breiman2001].

### Feature extraction
To predict relevance of a given publication, the classifier uses information from the publications in the dataset. Examples of such information are titles and abstracts. However, a model cannot make predictions from the titles and abstracts as they are; their textual content needs to be represented numerically. The textual information needs to be mapped to feature vectors. This process of numerically representing textual content is referred to as 'feature extraction'.

Term-Frequency Inverse Document Frequency (TF-IDF) stores the term frequency - the number of occurrences of each word - for each text in the dataset. The term frequency is adjusted with the inverse document frequency, the number of documents which contain the given word [@Ramos2003]. A disadvantage of TF-IDF and other bag-of-words methods is that they do not take the ordering of words into account, thereby ignoring semantics. An example of an approach that aims to overcome this weakness is D2V. D2V learns continuous distributed vector representations for pieces of text capable of grasping semantics by learning to predict the words in the texts [@Le2014].

### Query strategy
The active learning model can adopt different strategies in selecting the next publication to be screened by the reviewer. A strategy mentioned before is selecting the publication with the highest probability of being relevant. In the active learning literature this is referred to as certainty-based active learning [@Settles2012]. Another well-known strategy is uncertainty-based active learning, where the instances that are presented next are those instances on which the model's classifications are the least certain, i.e. close to 0.5 probability [@Settles2012]. Traditionally, this strategy trains the most accurate model because the model can learn the most from instances it is uncertain about. However, a study comparing the performance of both strategies in detecting relevant publications found that the accuracy gain of uncertainty-based screening was not significant [@Miwa2014].

Certainty-based active learning is the preferred strategy for the task at hand. Firstly, the aim of the screening process is to find relevant publications, not to train a good model. Secondly, certainty-based active learning is far better equipped at dealing with imbalanced data in active learning, as it aims to present only records that belong to the relevant class [@Fu2011].

## Simulation study
The section below describes the simulation study that was carried out to answer the research questions.

### Set-up
To address __RQ1__, four models combining every classifier with TF-IDF feature extraction were investigated:

 1. SVM + TF-IDF
 2. NB + TF-IDF
 3. RF + TF-IDF 
 4. LR + TF-IDF

To address __RQ2__, the classifiers were combined with D2V feature extraction, leading to the following three models:

 5. SVM + D2V
 6. RF + D2V
 7. LR + D2V

The combination NB + D2V could not be tested because the multinomial naive Bayes classifier^[https://scikit-learn.org/stable/modules/generated/sklearn.naive_bayes.MultinomialNB.html#sklearn.naive_bayes.MultinomialNB] can only handle a feature matrix with positive values, whereas the D2V feature extraction approach^[https://radimrehurek.com/gensim/models/doc2vec.html] produces a feature matrix that can also contain negative values. The performance of the seven models was evaluated by simulating every model on six systematic review datasets, addressing __RQ3__. Hence, 42 simulations were carried out, representing all model-dataset combinations.

Instead of having a human reviewer label publications manually, the screening process was simulated by retrieving the labels in the data. Each simulation started with an initial training set of one relevant and one irrelevant publication to represent a 'worst case scenario' where the reviewer has minimal prior knowledge on the publications in the data. The model was retrained every time after a publication had been labelled. A simulation ended after all publications in the dataset had been labelled. To account for sampling variance, each simulation was repeated 15 times. To account for bias due to the content of the initial publications, the initial training set was randomly sampled from the dataset for each of the 15 trials. Although varying over trials, the 15 initial training sets were kept constant over datasets to allow for a direct comparison of models within datasets. A seed value was set to ensure reproducibility. The simulation study was carried out using the \texttt{ASReview} simulation extension [@ASReview2020]. For each simulation, hyperparameters were optimized through a ... tpe to arrive at maximum model performance.  todo
<!-- About a 1000 runs, retraining after labelling 20, after visual inspection.  --> 
Scripts for the hyperparaemter optimization and resulting hyperparameter values can be found at GitHub.

Simulations were carried out in ASReview, version 0.9.3 [@ASReview2020]. Analyses were carried out using \texttt{R}, version 3.6.1 [@RCoreTeam2019]. Scripts and data are stored in the GitHub repository for this paper^[https://github.com/asreview/paper-evaluating-models-across-research-areas].The output resulting from the simulation was stored on the Open Science Framework page of this paper ^[https://osf.io/7mr2g/], as their size exceeded the storage limit of GitHub by far. Due to their large number, the simulations were carried out on Cartesius, the Dutch national supercomputer. Access was granted by SURF via a grant (ID EINF-156).

### Datasets 
The models were simulated on a convenience sample of six systematic review datasets. The data selection process was driven by two factors. Firstly, datasets were collected from various research areas to assess generalizability of the models across research contexts (RQ3). Secondly, all original data files should be openly published with a CC-BY license.

todo to different section. and are available through the [ $\texttt{ASReview}$ GitHub page](https://github.com/asreview/systematic-review-datasets).

The models are simulated on a convenience sample of six systematic review datasets. The data selection process was driven by two factors. Firstly, datasets are collected from various research areas to assess generalizability of the models across research contexts (RQ3). Secondly, all original data files have to be openly published with a CC-BY license. Datasets are available through this paper’s GitHub page.

The Wilson dataset [@Appenzeller-Herzog2020] - from the field of _medicine_ - is on a review on the effectiveness and safety of treatments of Wilson Disease, a rare genetic disorder of copper metabolism [@Appenzeller-Herzog2019]. From the same scientific field, the ACE dataset contains publications on the efficacy of Angiotensin-converting enzyme (ACE) inhibitors, a drug treatment for heart disease [@Cohen2006]. Additionally, the Virus dataset is from a systematic review on studies that performed viral Metagenomic Next-Generation Sequencing (mNGS) in farm animals [@Kwok2020]. From the field of _computer science_, the Software dataset contains publications from a review on fault prediction in software engineering [@Hall2012]. The Nudging dataset [@Nagtegaal2019a] belongs to a systematic review on nudging healthcare professionals [@Nagtegaal2019], stemming from the _humanities_. From _social science_, the PTSD dataset contains publications on studies applying latent trajectory analyses on posttraumatic stress after exposure to traumatic events [@vandeSchoot2017]. Of these six datasets, ACE and Software have been used for model simulations in previous studies on ML-aided title and abstract screening [@Cohen2006; @Yu2018].

Data were preprocessed from their original source into a dataset, containing title and abstract of the publications obtained in the initial search. Candidate studies with missing abstracts and duplicate instances were removed from the data. Preprocessing scripts and resulting datasets can be found at GitHub repository for this paper. Datasets were labelled to indicate which candidate studies were included in the systematic review, thereby indicating relevant publications. All datasets consisted of thousands of candidate studies, of which only a fraction was deemed relevant to the systematic review. For the Virus and the Nudging dataset, the inclusion rate was about 5 percent. For the remaining six datasets, inclusion rates were centered around 1-2 percent. (`r table_nums("datasets", display = "cite")`). 
<!-- The data preprocessing scripts can be found on the GitHub^[https://github.com/GerbrichFerdinands/asreview-thesis] -->

```{r, results = 'asis'}
# datasets table
drops <- readRDS(paste0(pathRdata, "drops.RDS"))
all <- readRDS(paste0(pathRdata, "all.RDS"))

nstudies <- function(all, datastage, col){
}

nostudies <- function(all, set, stage){
  sapply(all, function(x) x[set, stage])
}
inclrate <- function(all, set){
  sapply(all, function(x) round(x[set,"incl"]/x[set,"search"]*100,2))
}
datastats <- 
  tibble(dataset = names(all), 
       #Citation = NA, # maybe add footnote citation with kableExtra i.o. this.
       # paper 
       `candidates_paper` = nostudies(all, "paper", "search"), 
       #`fulltext_paper` = nostudies(all, "paper", "ftext"),
       `incl_paper` = nostudies(all, "paper", "incl"),
       `inclrate_paper` = inclrate(all, "paper"),
       # test set 
       `candidates_test` =  nostudies(all, "test", "search"), 
       #`fulltext_test` =  nostudies(all, "test", "ftext"), 
       `incl_test` =  nostudies(all, "test", "incl"), 
       `inclrate_test` = inclrate(all, "test")
       )

test <- 
datastats %>%
  select(dataset, candidates_test, incl_test) %>%
  mutate(excl = candidates_test - incl_test)

test <- rbind(test,test) %>%
  mutate(relevancy = c(rep("incl", 6), rep("excl", 6))) 
a <- c(test[1:6,"incl_test"], test[7:12, "excl"])
test[,"number"] <- c(a$incl_test, a$excl)
```

```{r, results = 'asis', caption = "Datasets obtained from six original systematic reviews.", eval = FALSE}
test %>%
  ggplot(aes(x=dataset, y = number, fill = relevancy)) +
  geom_col(position = "stack") +
  # scale_fill_brewer(palette="Set1") +
  scale_fill_manual(values= c("#CC6666","#66CC99"))+
  theme_bw() +
  theme(legend.position = "none") +
  labs(y="Candidate studies", x = "") +
  coord_flip()

#ggsave("datasets.png")
```

\begin{center}
`r table_nums("datasets")` 
\end{center}
```{r, results = 'asis'}
datastats <- datastats %>%
  select(dataset, candidates_test, incl_test, inclrate_test) 
colnames(datastats) <- c("dataset", rep(c("Candidate publications", 
                                       #"Studies selected for fulltext screening", 
                                       "Relevant publications", 
                                       "Inclusion rate (%)"),1))
datastats <- datastats[c(2,3,4,1,5,6), ]
datastats[datastats == "Ace"] <- "ACE"
saveRDS(datastats, "tables/tab1_data.RDS")

print(xtable(datastats, digits = c(0,0,0,0,1)),
      include.rownames=FALSE, comment = FALSE, booktabs = TRUE,
      format.args = list(big.mark = ",", decimal.mark = "."))
```

### Evaluating performance
Model performance was assessed by three different measures, Work Saved over Sampling (WSS), Relevant References Found (RRF), and Average Time to Discovery (ATD). 
WSS indicates the reduction in publications needed to be screened, at a given level of recall [@Cohen2006]. Typically measured at a recall level of 0.95, WSS@95 yields an estimate of the amount of work that can be saved at the cost of failing to identify 5% of relevant publications. In the current study, WSS is computed at 0.95 recall. RRF@10 represents the proportion of relevant publications that are found after screening 10% of all publications.

Both RRF and WSS are sensitive to the position of the cutoff value. Moreover, WSS makes assumptions about acceptable recall levels whereas this level might depend on the research question at hand [@OMara-Eves2015]. Therefore, we introduce the ATD, the average proportion of publications needed to screen to find a relevant publication. The ATD is an indicator of performance throughout the entire screening process instead of performance at some arbitrary cutoff value.

Furthermore, model performance was visualized by plotting recall curves. Plotting recall as a function of the proportion of screened publications offers insight in model performance throughout the entire screening process [@Cormack2014; @Yu2018]. The curves give information in two directions. On the one hand they display the number of publications that need to be screened to achieve a certain level of recall, but on the other hand they present how many relevant publications are identified after screening a certain proportion of all publications (RRF).

For each simulation, the RRF@10, WSS@95, and ATD are reported as means over 15 trials. To indicate the spread of performance within simulations, the means are accompanied by an estimated^[The metrics for all individual 15 trials deviate slightly from the overal mean over 15 trials because of pre-averaging in the \texttt{ASReview} source code. As the analyses across all trials did not produce information on the 15 separate runs, the standard deviation of the mean, $\hat s$, was estimated by computing the standard deviation within the individual 15 trials.] standard devation $\hat s$. To compare the overall performance across datasets, median performance is reported for every dataset, accompanied by the Median Absolute Deviation (MAD), indicating variability between models within a certain dataset. Recall curves are plot for each simulation, representing the average recall over 15 trials $\pm$ the standard error of the mean.

# Results 
This section proceeds as follows: Firstly, the results of the Nudging dataset are discussed in detail to provide a basis for answering the research questions. Secondly, the results are presented for each research question over all datasets.

```{r, eval = TRUE}
# measures for averages over 15 runs 
results <- readRDS("../../results/output/results.RDS")
stabres <- readRDS("../../results/output/tabresults.RDS")
```

```{r}

# table for in manuscript

sdruns <- readRDS("../../results/output/sdruns.RDS")

nicetab <- function(results, statistic){
  test <- results %>% select(model, dataset, all_of(statistic))
  sdname <- paste0("sd", statistic)

  test <- left_join(test, sdruns[,c("model", "dataset", sdname)], by = c("model", "dataset"))

  test[,statistic] <- sprintf("%.1f", round(test[,statistic],1)) 
  test[,sdname] <-  sprintf("%.2f", round(test[,sdname],2)) 
  test$tab <- with(test, paste0(test[,statistic], " (", test[,sdname], ")"))

  tab <- test %>%
      select(model, dataset, tab) %>%
      pivot_wider(names_from = dataset, values_from = c("tab"))

  tab <- tab %>%
    select(model, nudging, ptsd, software, ace, virus, wilson)
  names(tab) <- c("", "Nudging", "PTSD", "Software", "ACE", "Virus", "Wilson")
  return(tab)
}

# rrf10tab <- tabruns %>%
#   select(dataset, model, mean_rrf10, sd_rrf10) %>%
#   pivot_wider(names_from = dataset, values_from = c("mean_rrf10","sd_rrf10"))
# 
# atdtab <- tabruns %>% 
#   select(dataset, model, mean_loss, sd_loss) %>%
#   pivot_wider(names_from = dataset, values_from = c("mean_loss","sd_loss"))
# tabruns %>%
#   select(model, starts_with(mean_wss95), starts_with(sd_wss95),
#          starts_with(mean_rrf10), starts_with(sd_rrf10), 
#          starts_with(mean_loss), starts_with(sd_loss))
```

## Evaluation on the Nudging dataset
Figure 1a shows the recall curves for all simulations on the Nudging dataset. As desribed in the previous section, these curves plot recall as a function of the proportion of publications screened. The curves represent the average recall over 15 trials $\pm$ the standard error of the mean in the direction of the y-axis. The x-axis is cut off at 40% since at this point in screening all models had already reached 95% recall. The dashed horizontal lines indicate the RRF@10 values, the dashed vertical lines the WSS@95 values. The dashed grey diagonal line corresponds to the expected recall curve when publications are screened in a random order. Desirable model performance was defined as maximizing recall while minimizing the number of publications needed to screen.

The recall curves were used to examine model performance throughout the entire screening process and to make a visual comparison between models within datasets. For example in Figure 1a, after screening about 30% of the publications all models had already found 95% of the relevant publications. Moreover, after screening 5% the green curve - representing the RF + TF-IDF model - splits away from the others and remains to be the lowest of all curves until about 30% of publications have been screened. Hence, from screening 5 to 30 percent of publications, the RF + TF- IDF model was the slowest in finding the relevant publications. The ordering of the remaining recall curves changes throughout the screening process, but maintain to show relatively similar performance at face value.

Figure 1b shows a subset of the recall curves in Figure 1a, namely the curves for the first four models only to allow for a visual comparison across classification techniques adopting the TF-IDF feature extraction strategy. Figure 1c shows recall curves for the remaining three models to compare the models using D2V feature extraction. Figures 1d to 1f plot recall curves for models adopting the TF-IDF feature extraction strategy to recall curves for their D2V-using counterparts to allow for a comparison between models adopting TF-IDF and models adopting D2V feature extraction.

```{r, eval = TRUE, fig-sub-nudging, fig.cap = "Recall curves of different models for the Nudging dataset, indicating how fast the model finds relevant publications during the process of screening publications. Figure a displays curves for all seven models at once. The remaining figures display curves for several subsets of those models to allow for a more detailed inspection.", fig.subcap=c('All seven models',  "Models adopting TF-IDF feature extraction", "Models adopting D2V feature extraction",'D2V vs TF-IDF for LR classifier', 'D2V vs TF-IDF for RF classifier', "D2V vs TF-IDF for SVM classifier")}
# path
# nudgingplots <- list.files("../../results/one_seed/plots/nudging", full.names=TRUE)
f <- c("inclusion_all_nudging", "inclusion_tfidf_nudging", "inclusion_d2v_nudging", "inclusion_d2v_vs_tfidf_logistic_nudging", "inclusion_d2v_vs_tfidf_rf_nudging", "inclusion_d2v_vs_tfidf_svm_nudging")
nudgingplots <- paste0("../../results/one_seed/plots/nudging/", f, ".pdf")
include_graphics(nudgingplots)
```
<!-- is apparent that models adopting the TF-IDF feature extraction strategy are outperformed by their Doc2Vec-using counterparts:  -->
<!-- The best performing model in terms of ATD is NB + TF-IDF (9.25, 0.29), LR + TF-IDF (9.46, 0.19), SVM + TF-IDF (62.20),and RF + TF-IDF (11.6, 0.43). SVM + D2V (8.71, 0.33), LR + D2V (8.73, 0.47), RF + D2V (10.17, 0.88).  -->

<!-- provides the RRF@10 values for all model-dataset combinations.  -->
It can be seen from `r table_nums("atd", display = "cite")` that in terms of ATD, the best performing models on the Nudging dataset were SVM + D2V and LR + D2V, both with an ATD of 8.9%. This indicates that the average proportion of publications needed to screen to find a relevant publication was 8.9% for both models. In the SVM + D2V model, the standard deviation was 0.33 , whereas for the LR + D2V model $\hat s =$ 0.47. This indicates that for the SVM + D2V model, the ATD values of individual trials were closer to the overal mean compared to the LR + D2V model, meaning that the SVM + D2V model performed more stable across different initial training datasets. Median ATD for this dataset was 9.6% with an MAD of 1.06, indicating that for half of the models, the ATD was within 1.06 distance from the median ATD.

\begin{center}
`r table_nums("atd")`
\end{center}

```{r, results="asis"}
tabatd <- nicetab(results, "loss") 
tabatd <- tabatd[c(7, 1, 5, 3, 6, 4, 2),]
# add range rows 
mad <- results %>% group_by(dataset) %>% summarise(median = sprintf("%.1f", round(median(loss), 1)), mad = sprintf("%.2f", round(mad(loss), 2)))

mad <- with(mad, paste0(median, " (", mad, ")"))

tabatd <- rbind(tabatd, (c("median (MAD)", mad[c(2:4, 1, 5,6)])))

saveRDS(tabatd, file = "tables/tab2_atd.RDS")

print(xtable(tabatd, align = c("r", "r", rep("c", 6))), 
      include.rownames=FALSE, comment = FALSE, booktabs = TRUE, hline.after = c(0,7))

```

As `r table_nums("wss", display = "cite")` shows, the highest WSS@95 value on the Nudging dataset was achieved by the NB + TF-IDF model with a mean of 71.7%, meaning that this model reduced the number of publications needed to screen with 71.7% at the cost of losing 5% of relevant publications. The estimated standard deviation of 1.37% indicates that in terms of WSS@95, this model performed the most stable across trials. The model with the lowest WSS@95 value was RF + TF-IDF ($\bar x$ = 64.9, $\hat s =$ 2.50). Median WSS@95 of these models was 66.9%, with a MAD of 3.05%, indicating that WSS@95 values of models varied the most within this dataset.

\begin{center}
`r table_nums("wss")` 
\end{center}
```{r, eval = TRUE, results="asis"}
tabwss95 <- nicetab(results, "wss.95") 
tabwss95 <- tabwss95[c(7, 1, 5, 3, 6, 4, 2),]
# add range rows 
mad <- results %>% group_by(dataset) %>% summarise(median = sprintf("%.1f", round(median(wss.95), 1)), 
                                                   mad = sprintf("%.2f", round(mad(wss.95), 2)))
# insert mad 
mad <- with(mad, paste0(median, " (", mad, ")"))

tabwss95 <- rbind(tabwss95, (c("median (MAD)", mad[c(2:4, 1, 5,6)])))
# insert N per dataset 
saveRDS(tabwss95, file = "tables/tab3_wss95.RDS")

print(xtable(tabwss95, align = c("r", "r", rep("c", 6))), 
      include.rownames=FALSE, comment = FALSE, booktabs = TRUE, hline.after = c(0,7))

  # kable(format = "latex") %>%
  # kableExtra()
```

As can be seen from the data in `r table_nums("rrf", display = "cite")`, LR + D2V was the best performing model in terms of RRF@10, with a mean of 67.5 indicating that after screening 10% of publications, on average 67.5% of all relevant publications had been identified, with a standard deviation of 2.59. The worst performing model was RF + TF-IDF ($\bar x =$ 53.6, $\hat s =$ 2.71). Median performance was 62.6, with an MAD of 3.89 indicating again that RRF@10 values were most dispersed for models within this dataset.

\begin{center}
`r table_nums("rrf")` 
\end{center}
```{r, results="asis"}
tabrrf10 <- nicetab(results, "rrf.10")

tabrrf10 <- tabrrf10[c(7, 1, 5, 3, 6, 4, 2),]
# add range rows 
mad <- results %>% group_by(dataset) %>% summarise(median = sprintf("%.1f", round(median(rrf.10), 1)), mad = sprintf("%.2f", round(mad(rrf.10), 2)))

mad <- with(mad, paste0(median, " (", mad, ")"))

tabrrf10 <- rbind(tabrrf10, (c("median (MAD)", mad[c(2:4, 1, 5,6)])))
saveRDS(tabrrf10, file = "tables/tab4_rrf10.RDS")

print(xtable(tabrrf10, align = c("r", "r", rep("c", 6))), 
      include.rownames=FALSE, comment = FALSE, booktabs = TRUE, hline.after = c(0,7))

```
<!-- (67.74, 2.40), followed by SVM + D2V (67.61, 2.82), NB + TF-IDF (65.59, 2.57), RF + D2V (62.83, 5.54), LR + TF-IDF (62.22, 2.72), SVM + TF-IDF (60.40, 3.29), AND RF + TF-IDF (53.74, 2.71) . -->

```{r, eval = FALSE}
stabres %>%
  select(model, starts_with("wss.95")) %>%
  flextable() %>%
  colformat_num(j = 2:7, big.mark = ",", digits = 2, na_str = "NA") %>%
  colformat_char(j=1) %>%
  set_header_labels(wss.95_ace = "ACE", wss.95_nudging = "Nudging", wss.95_ptsd = "PTSD", wss.95_software = "Software", wss.95_virus = "Virus", wss.95_wilson = "Wilson") %>%
  theme_booktabs() %>%
  autofit()

stabres %>%
  select(model, starts_with("rrf.10")) %>%
  flextable() %>%
  colformat_num(j = 2:7, big.mark = ",", digits = 2, na_str = "NA") %>%
  colformat_char(j=1) %>%
    set_header_labels(rrf.10_ace = "ACE", rrf.10_nudging = "Nudging", rrf.10_ptsd = "PTSD", rrf.10_software = "Software", rrf.10_virus = "Virus", rrf.10_wilson = "Wilson") %>%

  theme_booktabs() %>%
  autofit()

stabres %>%
  select(model, starts_with("loss")) %>% 
  flextable() %>%
  colformat_num(j = 2:7, big.mark = ",", digits = 2, na_str = "NA") %>%
  colformat_char(j=1) %>%
  set_header_labels(loss_ace = "ACE", loss_nudging = "Nudging", loss_ptsd = "PTSD",loss_software = "Software", loss_virus = "Virus", loss_wilson = "Wilson") %>%
  theme_booktabs() %>%
  autofit() 
```

```{r, eval = FALSE}
atdrank <-
  results %>%
  group_by(dataset) %>%
  mutate(atd_rank = dense_rank(loss)) %>%
  select(dataset, model, loss, atd_rank)

# %>%
#   filter(model == "SVM + TF-IDF")

wssrank <- 
results %>%
  group_by(dataset) %>%
  mutate(wss95rank = dense_rank(-wss.95)) %>%
  select(dataset, model, wss95rank)

rrfrank <-
results %>%
  group_by(dataset) %>%
  mutate(rrf10rank = dense_rank(-rrf.10)) %>%
  select(dataset, model, rrf10rank)

tabruns %>%
  group_by(dataset) %>%
  mutate(atd_rank = dense_rank(mean_loss)) %>%
  filter(model == "RF + TF-IDF")

tabruns %>%
  mutate(wss95rank = dense_rank(-mean_wss95)) %>%
  filter(model == "RF + TF-IDF")

tabruns %>%
  mutate(rrf10rank = dense_rank(-mean_rrf10)) %>%
  filter(model == "RF + TF-IDF")

```

## Overall evaluation 
Recall curves for the simulations on the five remaining datasets are presented in Figure 2. For the sake of conciseness, recall curves are only plotted once per dataset, like in Figure 1a. Please refer to Additional file 1 for figures presenting subsets of recall curves for the remaining datasets, like in Figure 1b-f.

First of all, for all datasets, the models were able to detect the relevant publications much faster compared to when screening publications at random order as the recall curves exceed the expected recall at screening at random order by far. Even the worst results outperform this reference condition. Across simulations, the ATD was at maximum 11.8% (in the Nudging dataset), the WSS@95 at least 63.9% (in the Virus dataset), and the lowest RRF@10 was 53.6% (in the Nudging dataset). Interestingly, all these values were achieved by the RF + TF-IDF model.

Similar to the simulations on the Nudging dataset (Figure 1b), the ordering of recall curves changes throughout the screening process, indicating that model performance is dependent on the number of publications that have been screened. Moreover, the ordering of models in the Nudging dataset (Figure 1b) does not replicate on the remaining five datasets (Figure 2).

```{r, eval = TRUE, fig-sub-alldata, fig.cap = "Recall curves of all seven models for all  indicating how fast the model finds relevant publications during the process of screening publications.  on (a) the PTSD, (b) Software, (c) ACE, (d) Virus, and (e) Wilson dataset.", fig.subcap=c("PTSD","Software","ACE","Virus","Wilson")}
d <- c("ptsd", "software", "ace", "virus", "wilson")
files <- c("../../results/one_seed/plots/ptsd/inclusion_all_ptsd.pdf" ,paste0("../../results/one_seed/plots/", d[2:5], "/nolegend_inclusion_all_", d[2:5], ".pdf"))

include_graphics(files)
```

### RQ1 - Comparison across classification techniques
The first reserach question was aimed at evaluating the first four models adopting either the NB, SVM, LR or RF classification technique, combined with TF-IDF feature extraction. When comparing ATD-values of the models (Table 2), the NB + TF-IDF model ranked first in the ACE, Nudging, PTSD, Virus and Wilson dataset, and second in the PTSD and the Software dataset, in which the LR + TF-IDF model achieved the lowest ATD value. The RF + TF-IDF ranked last in all of the datasets except in the ACE dataset, where the SVM + TF-IDF model achieved the highest ATD-value.

```{r, eval = FALSE}
atdrank %>% 
  filter(model %in% c("NB + TF-IDF", "RF + TF-IDF", "SVM + TF-IDF", "LR + TF-IDF")) %>% 
  arrange(dataset, atd_rank)
```

```{r, eval = FALSE}
wssrank %>% 
  filter(model %in% c("NB + TF-IDF", "RF + TF-IDF", "SVM + TF-IDF", "LR + TF-IDF")) %>% 
  arrange(dataset, wss95rank)
```
Additionally, in terms of WSS@95 (Table 3) the ranking of models was strikingly similar across all datasets. In the ACE, Nudging, Software, and Virus dataset, the highest WSS@95 value was always achieved by the NB + TF-IDF model, followed by LR + TF-IDF, SVM + TF-IDF, and RF + TF-IDF. In the PTSD dataset this ranking applied as well, except that the LR + TF-IDF and NB + TF-IDF showed equal WSS@95 values. The ordering of the models for the Wilson dataset was NB + TF-IDF, RF + TF-IDF, LR + TF-IDF and SVM + TF-IDF.

```{r, eval = FALSE}
rrfrank %>% 
  filter(model %in% c("NB + TF-IDF", "RF + TF-IDF", "SVM + TF-IDF", "LR + TF-IDF")) %>% 
  arrange(dataset, rrf10rank)
```
Moreover, in terms of RRF@10 (Table 4) the NB + TF-IDF model achieved the highest RRF@10 value in the ACE, Nudging, and Wilson dataset. LR + TF-IDF ranked first in the PTSD dataset, SVM + TF-IDF was the best performing model within the Wilson dataset. The RF + TF-IDF model was again the worst performing model within all datasets, with an exception for the Software dataset. In this dataset, NB + TF-IDF ranked fourth, the remaining three models achieved an equal RRF@10 score.

Taken together, these results show that while all four models perform quite well, the NB + TF-IDF shows high performance on all measures across all datasets, whereas the RF + TF-IDF model never performed best on any of the measures across all datasets. 
<!-- The RF + TF-IDF model did not perform well on any of the datasets. This model was listed the lower half of the ranking across all datasets for both WSS@95, RRF@10^[With one exception for the RRF@10 in the Software dataset, however the range for this model-dataset combination is relatively small min(RRF@10) = 98.2, max(RRF@10) = 99.3.], and ATD, where the model ranked lowest within four datasets. -->

<!-- all datasets. Interestingly this model held it position when comparing ATD-values of all seven models at once, except for the Wilson dataset, where it ranked second, outperformed by   -->

<!-- the NB + TF-IDF belonged to the top four performing models in every dataset in terms of ATD, ranking first in the Wilson, Ace, and Virus dataset.  -->

<!-- For the PTSD dataset, the top performing classifiers are SVM + D2V (1.36, 0.05) AND LR + D2V () -->
<!-- For the Software dataset, models perform very similar with the -->

<!-- On the PTSD dataset  classification strategies had good performance as WSS@95%.  -->
<!-- No direct comparison between datasets can be made so we summarise model performance in terms of ranking. -->

### RQ2 - Comparison across feature extraction techniques
This section is concerned with the question of how models using different feature extraction strategies relate to each other. The recall curves for the Nudging data (Figure 1d-f) show a clear trend of the models adopting D2V feature extraction outperforming their TF-IDF counterparts. This trend also shows from the WSS@95 and RRF@10 values indicated by the vertical and horizontal lines in the figure. Likewise, the ATD values (Table 2) indicate that for the models adopting a particular classification technique, the model adopting D2V feature extraction always achieved a smaller ATD-value than the model adopting TF-IDF feature extraction.

In contrast, this pattern of models adopting D2V outperforming their TF-IDF counterparts in the Nudging dataset does not replicate across other datasets. Whether evaluated in terms of recall curves, WSS@95, RRF@10 or ATD, the findings were mixed. Neither one of the feature extraction strategies showed superior performance within certain datasets nor within certain classification techniques.

```{r, eval = FALSE}
wssrank %>% 
  filter(model != "NB + TF-IDF") %>% 
  arrange(dataset, wss95rank)

atdrank %>% 
  filter(model != "NB + TF-IDF") %>% 
  arrange(dataset, atd_rank)

atdrank %>% 
  arrange(model, atd_rank)

```

```{r, eval = FALSE}
wssrank %>% 
  arrange(wss95rank, model, dataset)

atdrank %>% 
  arrange(atd_rank, model, dataset)

rrfrank %>% 
  arrange(rrf10rank, model, dataset)

```

### RQ3 - Comparison across research contexts 
First of all, models showed much higher performance for some datasets than for others. Whilst performances on the PTSD (Figure 2a) and the Software dataset (Figure 2b) were quite high, performances were much lower across models for the Nudging (Figure 1a) and Virus (Figure 2d) datasets. There does not seem to be a clear distinction between the datasets from the biomedical sciences (ACE, Virus, and Wilson) and datasets from other fields (Nudging, PTSD, Software). The PTSD, Software and Nudging dataset also demonstrated high performances in terms of the median ATD, WSS@95 and RRF@10 values for these models (Table 2, 3, and 4).

Secondly, variability of between-model performance differed across datasets. For the PTSD (Figure 2a), Software (Figure 2b), and the Virus (Figure 2d) datasets, recall curves form a tight group meaning that within these datasets, the models perform relatively similar. For the Nudging (Figure 1a), ACE (Figure 2c), and Wilson (Figure 2e) dataset, the recall curves are much further apart, indicating that model performance is much more dependent on the classification technique and feature extraction strategy. The MAD values of the ATD, WSS@95 and RRF@10 confirm that within the PTSD, Software and Virus datasets, model performance is less spread out than within the Nudging, ACE and Wilson dataset.

Moreover, the curves for the ACE (Figure 2c) and Wilson (Figure 2e) datasets show a larger standard error of the mean compared to other the other datasets. For these datasets, model performance seemed to be more dependent on the initial training dataset compared to others.

<!-- For the PTSD dataset, WSS@95% ranges from 84.20% (RF + TF-IDF) to 93.75% (LR + TF-IDF).  -->

<!-- Overall these results indicate that first of all, model performance was mostly data dependent, however, the ordering of models is fairly similar.  -->
<!-- - the ordering of models is not the same in every dataset -->
<!-- - the spread between models performance differs over datasets, (high in ) -->
<!-- - social sciences vs medical sciences -->

<!-- #### plot:  -->
<!-- ## Overall Comparison -->
<!-- - models in general? -->
<!-- - which models perform better in which contexts? -->

<!-- all models perform better  -->
<!-- which models perform well overall.  -->

<!-- ranges in value: some datasets is matters more which models you choose, for some it matters less!  -->

<!-- #### across datasets  -->
<!-- For the SVM + TF-IDF model, WSS@95 ranged from 71.11 - 92.76 for the NB + TF-IDF model, WSS@95 ranged from 71.69 - 93.52, for -->

```{r, eval = FALSE}
metricsat <- c("WSS@95", "RRF@10", "ATD")
metrics <- c("WSS", "RRF", "ATD")
at <- c("@95", "@10", "")
Data <- c("ACE", "Nudging", "PTSD", "Software", "Virus", "Wilson")

typology <- data.frame(
  col_keys = names(stabres)[2:19],
  what = rep(Data, each = length(metrics)),
  #measure = rep(metricsat, each = 6),
  what = rep(c(rep("WSS@95"), rep("RRF@10"), rep("ATD")), 6),
  #measure = c(rep(c("@95", "@10", ""), each = 6)),
  stringsAsFactors = FALSE
)
num_keys <- names(stabres)[2:19]
test <- flextable(stabres)
###

Data <- Data[1]
data <- "ace"
acetab <- select(stabres, 
                  model, 
                  ends_with(data))
typology <- data.frame(
  col_keys = names(acetab)[2:ncol(acetab)],
  what = rep(Data, each = length(metrics)),
  #measure = rep(metricsat, each = 6),
  what = rep(c(rep("WSS@95"), rep("RRF@10"), rep("ATD")), length(Data)),
  #measure = c(rep(c("@95", "@10", ""), each = 6)),
  stringsAsFactors = FALSE
)
num_keys <- names(acetab)[2:length(acetab)]
test <- flextable(acetab)
```

```{r, eval = FALSE}
# format numbers
test <- colformat_num(x = test, j = num_keys, big.mark = ",", digits = 2, na_str = "missing")
test <- colformat_char(x=test, j = "model")
# set all columns 
test <- set_header_df(test, mapping = typology, key = "col_keys")
test <- merge_h(test, part = "header")
#border_v = fp_border(color="purple")
#test <- border_inner_v(test, part="all", border = border_v )
test <- merge_v(test, part = "header")
test <- theme_booktabs(test)
test <- autofit(test, add_w = 0.2)
test <- fix_border_issues(test)
test <- flextable::align(test, align = "left", part = "header" )

plot(test)
```

# Discussion 
The current study set out to evaluate performance of active learning models for the purpose of identifying relevant publications in systematic review datasets. 
<!-- This study adds to existing literature by evaluating models across four different classification techniques and two different feature extraction strategies. Moreover, as previous studies have paid little attention to the generalizability of active learning models across different research contexts, models were evaluated across six systematic review datasets from various research areas.  -->
It has been one of the first attempts to examine different classification strategies and feature extraction strategies in active learning models for systematic reviews. Moreover, this study has provided a deeper insight into the performance of active learning models across research contexts.

## Active learning-based screening prioritization
Overall, the findings confirm the great potential of active learning models in reducing workload for systematic reviewers. All models were able to detect 95% of the relevant publications after screening less than 40% of the total number of publications, indicating that active learning models can save more than half of the workload in the screening process. In a previous study, the ACE dataset was used to simulate a model that did not use active learning, finding a WSS@95 value of 56.61% [@Cohen2006], whereas the models in the current study achieved far superior WSS@95 values varying from 68.6% to 82.9% in this dataset. Active learning models clearly outperformed a model which did not use active learning. In addition, the Software dataset was used to simulate an active learning model [@Yu2018] and reached WSS@95 of 91%, strikingly similar the WSS@95 values found in the current study which ranged from 90.5% to 92.3%.

### Classification techniques
The first research question in this study sought to evaluate models adopting different classification techniques. The most obvious finding to emerge from these evaluations was that the NB + TF-IDF model consistently performed as one of the best models. The results suggest that whilst SVM - the classifier standardly used in software tools for active learning in systematic reviews - performed fairly well, the LR and NB classification techniques are good if not superior alternatives. Note that NB and LR were always good methods for text classifciation tasks [@Mitchell1997].
<!-- Across simulations the NB + TF-IDF model demonstrated high performance on  -->

### Feature extraction strategy 
The overall results on models adopting D2V versus TF-IDF feature extraction strategy remain inconclusive. According to these findings, adopting D2V instead of the well-established TF-IDF feature extraction strategy does not lead to better performing models. Given these results, although preliminary, preference goes out to the TF-IDF feature extraction technique as this relatively simplistic technique will lead to more interpretable model.

### Research contexts 
Simulating models on a heterogenous collection of systematic review datasets revealed that model performance is very data-dependent. Within some datasets, models achieved much higher overall performance than within other datasets. Moreover, for some datasets, differences between models were much larger than for other datasets. It has been suggested that active learning is more difficult for datasets from the social sciences compared to data from the medical sciences [@Miwa2014]. This does not appear to be the case as performance within the biomedical datasets (Wilson, Virus, ACE) was not in any way superior to performance within the datasets from the social sciences (PTSD and Nudging). An issue that emerges from these findings is that difficulty of active learning was not confined to any particular research area. A possible explanation for this is that difficulty of active learning could be attributed to factors more directly related to the systematic review at hand, such as the inclusion rate and the complexity of inclusion criteria used to identify relevant publications [@Gates2018a; @Rathbone2015]. Although the current study did not investigate the inclusion criteria of systematic reviews, the datasets on which the active learning models performed worst, Nudging and Virus, were interestingly also the datasets with the highest inclusion rates, 5.4% and 5.0%, respectively.

<!-- as differences in model performance could not be linked to non-biomedical datasets only from the biomedical areas (Nudging, Virus, and Wilson) did not perform did the datasets in which models pachieved highest performance  -->

<!-- although the nature of this study was rather exploratory, the results yield important insights on the ...  -->

<!-- Since models are more data dependent than model dependent (although for some datasets model performance varies more widely than for others). Maybe it doesn't really matter which model you pick although if you want to be save, the nb is a one size fits all model.

## Strengths and limitations
<!-- The findings from this study have extened our knowledge on performance of different classification strategies We added to this emerging field of research by expanding model evaluations across classification techniques and feature extraction strategies. The findings from this study has extended our knowledge on  -->

<!-- what does this mean for reviewing?? save a lot of time ! -->
<!-- and if you've limited screening time then why not use such an prioritization model.  -->

<!-- This thesis adds to the existing knowledge on active learning models for sreening publications by evaluating  -->
<!-- We evaluted active learning models using established methods (certainty sampling, rebalancing, threw some new methods in the mix and evaluated them on heterogenous   -->
<!-- what did we add to other research  -->

<!-- We were able to reach WSS@95 of 63.9-91.7% (instead of 30-70 as stated in [@OMara-Eves2015]). Depending on the model-dataset combination.  -->
<!-- current study adds to this field by expanding -->
<!-- Across all classification techniques, feature extraction strategies and review contexts, active learning models exceeded screening at random order by far.  -->

## Limitations and future research
When applied in systematic review practice, the succes of active learning models stands or falls down with the generalizability of model performance across unseen datasets. It is important to bear in mind that model hyperparameters were optimized for each model-dataset combination. Thus, the observed results reflect maximum model performance for the datasets at hand. The question remains whether model performance generalizes to datasets for which the hyperparameters were not optimized. Further research should be undertaken to determine the sensitivity of model performance to the hyperparameter values.

Additionally, whilst the sample of datasets in the current study is diverse compared to previous studies, the sample size does not allow for investigating how characteristics of the data - such as inclusion rate - relate to model performance. To build confidence in active learning models for screening publications, it is essential to identify how characteristics of the data influence model performance. Such a study requires more data on systematic reviews. Therefore, we call systematic reviewers to make an effort to openly publish their datasets.

<!-- the datasets are by no means representative of the current body of systematic reviews.  -->

<!-- When models are applied in a real world setting there is no way to know of optimal parameter. Future research will focus on a cross-validation  -->
<!-- study did not seek to address results to differences in ... ?  -->

<!-- The current study was designed to detecting the final inclusions in a systematic review.  -->
<!-- It is interesting to note that studies use only abstract + title infor -->
<!-- Since the active learning models base their predictions only on metadata... -->
<!-- Screening publications in systematic reviews is typically a two-step process. First, titles and abstracts are screened to identify potentially relevant publications, called abstract inclusions. Second, the fulltexts of these publications are read to identify the relevant publications. This implies that the relevant publications are selected based on information that the models do not have. To truly assess the added value of active learning models in title-and-abstract screening, models should be evaluated on their capability of detecting the abstract inclusions instead of relevant publications only. However, this data is typically not available. Hence, greater efforts are needed to provide information on the abstract inclusions in openly published systematic review datasets.  -->

<!-- just exploratory, no fancy statistical analysis, however the goal was not to look for the best model. -->
<!-- strengths: -->

<!-- - open data -->
<!-- - different research areas -->
<!-- - different models on same dataset -->
<!-- - different datasets on same model -->

An unanticipated finding was that the runtime of simulations varied widely across models, indicating that some models need more time to retrain after a publication has been labelled than other models. This finding has important implications for the practical application of such models, as an efficient model should be able to keep up with the decision-making speed of the reviewer. Further studies taking into account retraining time will need to be undertaken.

<!-- Now that we know that the potential of models/that the models show good performance, the next step is to identify factors that drive performance. When the only information available is an unlabelled dataset and we only want to try one model,  When applied in practice, it is  -->
<!-- - challenge is how to define BEFORE knowing the answers, what the model is.. In practice, knowing what the best mdoel is after labelling has no value whatsoever, the model + hyperparameters need to be defined a priori.  -->
<!-- - difficult to distinguish performance over datasets, especially when applied on a dataset of which no prior information is known (e.g. inclusions isn't known in practice).  -->

# Conclusions 
Overall, the findings of this study confirm that active learning models show great potential of retrieving relevant publications in a systematic review dataset, while minimizing the number of publications needed to screen. The results shed new light on the performance of different classification techniques, indicating that the NB classification technique is superior to the widely used SVM As model performance differs vastly across datasets, this study raises the question what causes models to yield more workload savings for some systematic review datasets than for others. In order to facilitate the applicability of active learning models in systematic review practice, it is essential to identify how dataset characteristics relate to model performance.

<!-- This study shows that active learning models are   -->
<!-- As all models exceeded screening at random order by far, they show great potential in reducing workload of systematic reviewers.  -->
<!-- of more precise estimate of  -->

# Declarations
## List of abbreviations

ATD

D2V
LR
MAD
ML
NB
PTSD
RF
RRF
SD
SEM
SVM
TF-IDF
WSS

## Ethics approval and consent to participate
This study has been approved by the Ethics Committee of the Faculty of Social and Behavioural Sciences of Utrecht University, filed as an amendment under study 20-104.

## Consent for publication
Not applicable.

## Availability of data and materials 
As reported in the main text, all data and materials are available through the GitHub repository for this paper, https://github.com/asreview/paper-evaluating-models-across-research-areas. This repository contains all systematic review datasets used during this study and their preprocessing scripts, scripts and data on the hyperparamter optimization, scripts on the simulations, scripts for analyzing the results of the simulations, and the source files for this manuscript. All output files of the simulation study are stored on the Open Science Framework page of this paper, https://osf.io/7mr2g/.

## Competing interests
The author declares that they has no competing interests.

## Funding
Computing hours on the Cartesius supercomputer were funded by SURFsara. SURFsara had no role whatsoever in the design of the current study, nor in the data collection, analysis and interpretation, nor in writing the manuscript.

## Author's contributions
RS developed the DR balance strategy and ATD metric.

All authors read and approved the final manuscript.

## Acknowledgements
I am grateful for all researchers who have made great efforts to openly publish the data on their systematic reviews, special thanks go out to Rosanna Nagtegaal. I would also like to thank dr. Caroline van Baal for supporting me in writing this paper, and prof. dr. René Eijkemans, for being the second grader of this paper Finally, I would like to express my appreciation to my supervisors prof. dr. Rens van de Schoot, Jonathan de Bruin, and dr. Raoul Schram. Your door was always open and your enthusiasm was contagious.
<!-- \newpage -->

<!-- # Appendix A - list of definitions -->

<!-- ### Feature Extraction Strategies  -->
<!-- split_ta = overall hyperparameter  -->

<!-- #### TF-IDF -->

<!-- ##### hyperparameters  -->
<!--     ngram_max: int -->
<!--             Can use up to ngrams up to ngram_max. For example in the case of -->
<!--             ngram_max=2, monograms and bigrams could be used. -->

<!-- #### Doc2vec -->
<!-- Predicts words from context.  -->
<!-- Aims at capturing the relations between word (man-woman, king-queen). -->
<!-- [@Le2014]. Using a neural network.  -->

<!-- using Continuous Bag-of-Words (CBOW), Skip-Gram model, .... -->
<!-- Word vector _W_ and extra: document vector _D_, trained to predict words in the text. -->

<!-- From gensim [@Rehurek2010]. -->

<!--         Arguments -->
<!--         --------- -->
<!--         vector_size: int -->
<!--             Output size of the vector. -->
<!--         epochs: int -->
<!--             Number of epochs to train the Doc2vec model. -->
<!--         min_count: int -->
<!--             Minimum number of occurences for a word in the corpus for it to -->
<!--             be included in the model. -->
<!--         workers: int -->
<!--             Number of threads to train the model with. -->
<!--         window: int -->
<!--             Maximum distance over which word vectors influence each other. -->
<!--         dm_concat: int -->
<!--             Whether to concatenate word vectors or not. -->
<!--             See paper for more detail. -->
<!--         dm: int -->
<!--             Model to use. -->
<!--             0: Use distribute bag of words (DBOW). -->
<!--             1: Use distributed memory (DM). -->
<!--             2: Use both of the above with half the vector size and concatenate -->
<!--             them. -->
<!--         dbow_words: int -->
<!--             Whether to train the word vectors using the skipgram metho -->

<!-- #### SBERT -->

<!-- BERT-base model with mean-tokens pooling [@Reimers2019] -->

<!-- #### embeddingIdf  -->
<!-- This model averages the weighted word vectors of all the words in the text, -->
<!-- in order to get a single feature vector for each text. The weights are provided by the inverse document frequencies -->

<!-- ### Models  -->
<!-- #### Naive Bayes  -->
<!-- Naive Bayes assumes all features are independent given the class value. [@Zhang2004]  -->

<!-- ASReview uses the ` MultinomialNB` from the scikit-learn package [@scikit-learn], that implements the naive Bayes algorithm for multinomially distributed data.  -->
<!-- ` nb` -->

<!-- Hyperparameters -->

<!--   * alpha - accounts for features not present in learning samples and prevents zero probabilities in further computations.  -->

<!-- #### Random Forests -->
<!-- A number of decision trees are fit on bootstrapped samples of the original data,  -->
<!-- [@Breiman2001] -->
<!-- RandomForestClassifier from sklearn  -->

<!-- Arguments -->
<!--         --------- -->
<!--         n_estimators: int -->
<!--             Number of estimators. -->
<!--         max_features: int -->
<!--             Number of features in the model. -->
<!--         class_weight: float -->
<!--             Class weight of the inclusions. -->
<!--         random_state: int, RandomState -->
<!--             Set the random state of the RNG. -->
<!--         """ -->

<!-- #### Support Vector Machine -->

<!--  Arguments -->
<!--         --------- -->
<!--         gamma: str -->
<!--             Gamma parameter of the SVM model. -->
<!--         class_weight: -->
<!--             class_weight of the inclusions. -->
<!--         C: -->
<!--             C parameter of the SVM model. -->
<!--         kernel: -->
<!--             SVM kernel type. -->
<!--         random_state: -->
<!--             State of the RNG. -->

<!-- #### Logistic Regression  -->

<!-- #### __Dense Neural Network__ -->

<!-- ### Query Strategies  -->
<!-- * Max - Choose the most likely samples to be included according to the model -->
<!-- * Uncertainty - choose the most uncertain samples according to the model (i.e. closest to 0.5 probability) [@Lewis1994]  -->
<!-- * Random - randomly selects abstracts with no regard to model assigned probabilities.  -->
<!-- * Cluster - Use clustering after feature extraction on the dataset. Then the highest probabilities within random clusters are sampled -->

<!-- The following combinations are simulated:  -->

<!-- * cluster -->
<!-- * max -->
<!-- * cluster * random  -->
<!-- * cluster * uncertainty -->
<!-- * max * cluster -->
<!-- * max * random -->
<!-- * max * uncertainty -->

<!-- ### Balance Strategies  -->

<!-- ### amount of training data -->

<!-- * n_instances = number of papers queried each query -->
<!-- * n_queries = number of queries -->
<!-- * n_prior_included: 5 -->
<!-- * n_prior_excluded: -->

<!-- # Combinations  -->

<!-- This leads to  `r nrow(combins)` combinations of configurations.  -->

<!-- - Naive bayes only goes with tfidf feature extraction. -->
<!-- - For the feature extraction strategies we will focus on Doc2vec and tfidf. (but will compute all 4) -->
<!-- - This leads to 3 * 7  * 4 * 3 + 1 * 7 * 1 * 3 =  273 combinations.  -->

<!-- See appendix A for a table containing all 273 combinations.  -->

<!-- ## Cross-validation -->
<!-- Should give an accurate estimate of maximum performance / future systematic reviews to be performed.  -->

<!-- # Appendix B - combinations -->

<!-- ```{r} -->
<!-- names(combins) <- c("Model", "Query Strategy", "Feature extraction strategy") #, "Training data [included/excluded]") -->
<!-- combins %>% -->
<!--   kable(row.names = FALSE, format = "latex", longtable = TRUE, booktabs = TRUE) %>% -->
<!--   kable_styling(latex_options = c("repeat_header"))  -->
<!-- ``` -->

<!-- # Appendix C - supercomputer Cartesius -->

<!-- 500,000 SBU -->

<!-- Running on Cartesius is charged in System Billing Units (SBUs), and charging is based on the wall clock time of a job. On fat and thin nodes, an SBU is equal to using 1 core for 1 hour (a core hour), or 1 core for 20 minutes on a GPU node. Since compute nodes are allocated exclusively to a single job at a time, you will be charged for all cores on that node - even if you are using less.  -->

<!-- In the current study, the classifier and the feature extraction strategy are varied, whereas the query and balance strategy remain fixed.  -->
<!-- In the current study only a fraction of all possible configurations are tested for the sake of brevity.  -->
<!-- There are many more options available and open to exploration. -->
\nolinenumbers

# References

<div id="refs"></div>

# Additional file 1 
Recall curves plot separately for the PTSD, Software, ACE, Virus and Wilson datasets.

## PTSD
```{r, eval = TRUE, fig-sub-ptsd, fig.cap = "Recall curves for the PTSD dataset. ", fig.subcap=c('All seven models',  "Models adopting TF-IDF feature extraction", "Models adopting D2V feature extraction",'D2V vs TF-IDF for LR classifier', 'D2V vs TF-IDF for RF classifier', "D2V vs TF-IDF for SVM classifier") }
# path
f <- c("inclusion_all_", "inclusion_tfidf_", "inclusion_d2v_", "inclusion_d2v_vs_tfidf_logistic_", "inclusion_d2v_vs_tfidf_rf_", "inclusion_d2v_vs_tfidf_svm_")

plots <- paste0("../../results/one_seed/plots/ptsd/", f, "ptsd.pdf")
include_graphics(plots)
```

## Software
```{r, fig-sub-software, fig.cap = "Recall curves for the Software dataset. ", fig.subcap=c('All seven models',  "Models adopting TF-IDF feature extraction", "Models adopting D2V feature extraction",'D2V vs TF-IDF for LR classifier', 'D2V vs TF-IDF for RF classifier', "D2V vs TF-IDF for SVM classifier") }
# path
plots <- paste0("../../results/one_seed/plots/software/", f, "software.pdf")
include_graphics(plots)
```

## ACE
```{r, fig-sub-ace, fig.cap = "Recall curves for the ACE dataset. ", fig.subcap=c('All seven models',  "Models adopting TF-IDF feature extraction", "Models adopting D2V feature extraction",'D2V vs TF-IDF for LR classifier', 'D2V vs TF-IDF for RF classifier', "D2V vs TF-IDF for SVM classifier") }
# path
plots <- paste0("../../results/one_seed/plots/ace/", f, "ace.pdf")
include_graphics(plots)
```

## Virus
```{r, fig-sub-virus, fig.cap = "Recall curves for the Virus dataset. ", fig.subcap=c('All seven models',  "Models adopting TF-IDF feature extraction", "Models adopting D2V feature extraction",'D2V vs TF-IDF for LR classifier', 'D2V vs TF-IDF for RF classifier', "D2V vs TF-IDF for SVM classifier") }
# path
plots <- paste0("../../results/one_seed/plots/virus/", f, "virus.pdf")
include_graphics(plots)

```

## Wilson
```{r, eval = TRUE, fig-sub-wilson, fig.cap = "Recall curves for the Wilson dataset. ", fig.subcap=c('All seven models',  "Models adopting TF-IDF feature extraction", "Models adopting D2V feature extraction",'D2V vs TF-IDF for LR classifier', 'D2V vs TF-IDF for RF classifier', "D2V vs TF-IDF for SVM classifier") }
# path
plots <- paste0("../../results/one_seed/plots/wilson/", f, "wilson.pdf")
include_graphics(plots)
```